---
title: "R Notebook"
output: html_notebook
---

### Data anlysis to be performed before using NFDS model
# 1) assemble genomes if necessary (e.g, using shovill 1.1.0 using default parameters (Torsten Seemann. Tseemann/Shovill. 2024)) 
# 2) Compute Global pneumococcal sequence clusters (GPSCs) using PopPUNK 2.6.0 (Lees et al. (2019), doi: 10.1101/gr.241455.118.) with the PopPUNK database for S. pneumoniae GPS_v4 via poppunk_assign, applying quality control (--run-qc), allowing maximally one distance of zero (--max-zero-dist 1), and merging a maximum of three clusters based on one single isolate (--max-merge 3). 
# 3) Construct a pangenome: 
# 3.1) ggCaller 1.3.4 \cite{Horsfield et al. (2023), doi: 10.1101/gr.277733.123} can be used for gene calling using the SPARC dataset for annotation (Croucher et al. (2015), doi: 10.1038/sdata.2015.58 and Croucher et al. (2016), doi: 10.5061/DRYAD.T55GQ) and with the options --gene-finding-only --annotation sensitive. 
# 3.2) bakta (Schwengers et al. (2021), doi: 10.1099/mgen.0.000685) with default parameters can be used for functional annotation (if wanted)
# 3.3) panaroo v1.5.2 ( Tonkin-Hill et al. (2020), doi: 10.1186/s13059-020-02090-4) can be run with the options --refind-mode off --clean-mode moderate --family_threshold 0.7 --min_trailing_support 2 --trailing_recursive 0 --min_edge_support_sv 2 --edge_support_threshold 0.0 --aligner none

### Format data in the way the NFDS model expects it to be

# Read in meta data and PopPUNK GPSCs and add information to PopPUNK data frame
```{r}
MetaData <- read.csv("MetaData.csv") # maybe change to readr

PopPUNK_clusters <- read.csv("poppunk_clusters_external_clusters.csv")
# Add Serotype Information
Serotype_dict <- MetaData$Serotype
names(Serotype_dict) <- MetaData$filename
PopPUNK_clusters$Serotype <- Serotype_dict[PopPUNK_clusters$Taxon]
# Add sequencing year information
SeqYear_dict <- MetaData$`Year.of.Isolation`
names(SeqYear_dict) <- MetaData$filename
PopPUNK_clusters$SeqYear <- SeqYear_dict[PopPUNK_clusters$Taxon]

# calculate how many GPSCs are in the dataset
GPSC_no <- length(unique(PopPUNK_clusters$Cluster))
# Add VT information
VT_dict <- MetaData$`Vaccine.Type`
names(VT_dict) <- MetaData$filename
PopPUNK_clusters$VT <- VT_dict[PopPUNK_clusters$Taxon]

# if the data set contains two vaccine introductions (e.g. PCV7 and PCV13), create two separate dictionaries and add two columns to PopPUNK_clusters

# create dictionary between GPSC and index
PPclustToIndex_dict <- 1:length(unique(PopPUNK_clusters$Cluster))
names(PPclustToIndex_dict) <- (unique(PopPUNK_clusters$Cluster))
# create dictionary between serotype and index
SerotypeToIndex_dict <- 1:length(unique(PopPUNK_clusters$Serotype))
names(SerotypeToIndex_dict) <- (unique(PopPUNK_clusters$Serotype))
# create dictionary between GPSC and taxon
GPSC_dict <- PopPUNK_clusters$Cluster
names(GPSC_dict) <- PopPUNK_clusters$Taxon
```


# Create immigration matrix and start population
```{r}
# Create PP clusters, split by serotype
# calculate the frequency of the gene clusters and year and serotype
# 3 time points --> 3 matrices
# rows = GPSCs, columns = serotypes
PP_mass_cluster_freq_1_sero <- matrix(0,nrow = GPSC_no, ncol = length(unique(PopPUNK_clusters$Serotype)))
PP_mass_cluster_freq_2_sero <- matrix(0,nrow = GPSC_no, ncol = length(unique(PopPUNK_clusters$Serotype)))
PP_mass_cluster_freq_3_sero <- matrix(0,nrow = GPSC_no, ncol = length(unique(PopPUNK_clusters$Serotype)))
PP_mass_cluster_freq_sero_names <- unique(PopPUNK_clusters$Serotype)
for (i in 1:GPSC_no){
  for (k in unique(PopPUNK_clusters$Serotype)){
  PP_mass_cluster_freq_1_sero[i,which(unique(PopPUNK_clusters$Serotype)==k)] <- length(which(PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),][PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),]$Serotype ==k, ]$"SeqYear"==2001))
  PP_mass_cluster_freq_2_sero[i,which(unique(PopPUNK_clusters$Serotype)==k)] <- length(which(PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),][PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),]$Serotype ==k, ]$"SeqYear"==2004))
  PP_mass_cluster_freq_3_sero[i,which(unique(PopPUNK_clusters$Serotype)==k)] <- length(which(PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),][PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),]$Serotype ==k, ]$"SeqYear"==2007))
  }
}

# matrix shows information whether serotypes and GPSCs co-occur
PPxSero_matrix <- matrix(0, nrow = length(unique(PopPUNK_clusters$Cluster)), ncol = length(unique(PopPUNK_clusters$Serotype)))
for (i in 1:nrow(PopPUNK_clusters)) {
  PPxSero_matrix[PPclustToIndex_dict[as.character(PopPUNK_clusters$Cluster[i])],SerotypeToIndex_dict[PopPUNK_clusters$Serotype[i]]] <- 1
}
# matrix shows information whether PP and serotype co-occur in the dataset (1=TRUE)
# this can be used as the migration matrix:
PPxSero_matrix_prob <- PPxSero_matrix / sum(PPxSero_matrix)
PPsero_mig <- matrix(sapply(PPxSero_matrix_prob,as.double), nrow = nrow(PPxSero_matrix), ncol = ncol(PPxSero_matrix))
saveRDS(PPsero_mig, "PPsero_mig.rds") # immigration matrix


# create start population
# based on data, includes some noise
# and is expanded to a population size of 15,000
PPsero_startpop <- matrix(0, nrow = length(unique(PopPUNK_clusters$Cluster)), ncol = length(unique(PopPUNK_clusters$Serotype)))
PP_expand_factor <- 15000 / sum(PP_mass_cluster_freq_1_sero)
exp_noise <- 1000
for (i in 1:nrow(PPsero_startpop6)) {
    PPsero_startpop[i,] <- (sapply((PP_mass_cluster_freq_1_sero[i,] + rexp(n = length(PP_mass_cluster_freq_1_sero[i,]), rate = exp_noise)) * PP_expand_factor, rpois, n=1)) + ceiling(PPxSero_matrix_prob[i,])
}
saveRDS(PPsero_startpop, "PPsero_startpop.rds") # start population
```

# Create vector that saves which serotypes are affected by the vaccine
```{r}
# serotype VT vector
SeroVT <- rep(0, length(unique(PopPUNK_clusters$Serotype)))
for (i in 1:nrow(PopPUNK_clusters)) {
  SeroVT[SerotypeToIndex_dict[PopPUNK_clusters$Serotype[i]]] <- as.integer(PopPUNK_clusters$VT[i]=="VT")
}
saveRDS(SeroVT, "SeroVT.rds") # serotype vector

# if the data set contains two vaccine introductions (e.g. PCV7 and PCV13), create two separate SeroVT vectors 
# these cann then be passed to the model version called "NFDS_Model_2vacc.R"
```

# Start computing genotype matrix

# First, read in panaroo output
```{r}
gene_presence_absence_mtx <- read.csv("gene_presence_absence.csv")


# convert presence absence into boolean (0-1-based) matrix
convert_to_bool <- function(x){
  if (x=="") 0 else 1
}
gene_presence_absence_mtx_bool <- gene_presence_absence_mtx[,c(-2,-3)]
gene_presence_absence_mtx_bool[-1,-1] <- apply(gene_presence_absence_mtx_bool[-1,-1],c(1,2), convert_to_bool)
colnames(gene_presence_absence_mtx_bool) <- gene_presence_absence_mtx_bool[1,]
# filter for mobile genetic elements here if wanted
```

# Create time point specific presence absence matrices
```{r}
gene_presence_absence_mtx_bool_2001 <- data.frame(matrix(0, nrow = nrow(gene_presence_absence_mtx_bool), ncol = length(MetaData[MetaData$Year.of.Isolation==2001,]$filename)+1))
gene_presence_absence_mtx_bool_2001[1,-1] <- gene_presence_absence_mtx_bool[1,c(FALSE,SeqYear_dict[unlist(gene_presence_absence_mtx_bool[1,-1])]==2001)]
gene_presence_absence_mtx_bool_2001[-1,1] <- gene_presence_absence_mtx_bool[-1,1]
gene_presence_absence_mtx_bool_2001[-1,-1] <- gene_presence_absence_mtx_bool[-1,c(FALSE,SeqYear_dict[unlist(gene_presence_absence_mtx_bool[1,-1])]==2001)]

gene_presence_absence_mtx_bool_2004 <- data.frame(matrix(0, nrow = nrow(gene_presence_absence_mtx_bool), ncol = length(MetaData[MetaData$Year.of.Isolation==2004,]$filename)+1))
gene_presence_absence_mtx_bool_2004[1,-1] <- gene_presence_absence_mtx_bool[1,c(FALSE,SeqYear_dict[unlist(gene_presence_absence_mtx_bool[1,-1])]==2004)]
gene_presence_absence_mtx_bool_2004[-1,1] <- gene_presence_absence_mtx_bool[-1,1]
gene_presence_absence_mtx_bool_2004[-1,-1] <- gene_presence_absence_mtx_bool[-1,c(FALSE,SeqYear_dict[unlist(gene_presence_absence_mtx_bool[1,-1])]==2004)]

gene_presence_absence_mtx_bool_2007 <- data.frame(matrix(0, nrow = nrow(gene_presence_absence_mtx_bool), ncol = length(MetaData[MetaData$Year.of.Isolation==2007,]$filename)+1))
gene_presence_absence_mtx_bool_2007[1,-1] <- gene_presence_absence_mtx_bool[1,c(FALSE,SeqYear_dict[unlist(gene_presence_absence_mtx_bool[1,-1])]==2007)]
gene_presence_absence_mtx_bool_2007[-1,1] <- gene_presence_absence_mtx_bool[-1,1]
gene_presence_absence_mtx_bool_2007[-1,-1] <- gene_presence_absence_mtx_bool[-1,c(FALSE,SeqYear_dict[unlist(gene_presence_absence_mtx_bool[1,-1])]==2007)]
```

# Find intermediate genes (5-95%) in first sampling time point (here 2001)
```{r}
sum_as_int <- function(x){
  sum(as.integer(x))
}

gene_freq_2001 <- rep(0, nrow(gene_presence_absence_mtx_bool_2001)-1)
gene_freq_2001 <- apply(gene_presence_absence_mtx_bool_2001[-1,-1],1, sum_as_int)
gene_freq_2001 <- gene_freq_2001 / (ncol(gene_presence_absence_mtx_bool_2001)-1)

gene_filter <- as.integer(gene_freq_2001<=0.95 & gene_freq_2001>=0.05) # these are the genes that are at intermediate frequency 
```

# Filter gene presence absence matrix for intermediate genes only
```{r}
# intermediate gene presence absence matrices filtered by genes that are present at 5-95% frequency in 2001
intermed_gene_presence_absence_mtx_bool_2001 <- data.frame(matrix(0, nrow = sum(gene_filter)+1, ncol = length(MetaData[MetaData$Year.of.Isolation==2001,]$filename)+1))
intermed_gene_presence_absence_mtx_bool_2001 <- gene_presence_absence_mtx_bool_2001[c(1,which(gene_filter==1)+1),]

intermed_gene_presence_absence_mtx_bool_2004 <- data.frame(matrix(0, nrow = sum(gene_filter)+1, ncol = length(MetaData[MetaData$Year.of.Isolation==2004,]$filename)+1))
intermed_gene_presence_absence_mtx_bool_2004 <- gene_presence_absence_mtx_bool_2004[c(1,which(gene_filter==1)+1),]

intermed_gene_presence_absence_mtx_bool_2007 <- data.frame(matrix(0, nrow = sum(gene_filter)+1, ncol = length(MetaData[MetaData$Year.of.Isolation==2007,]$filename)+1))
intermed_gene_presence_absence_mtx_bool_2007 <- gene_presence_absence_mtx_bool_2007[c(1,which(gene_filter==1)+1),]

intermed_gene_presence_absence_mtx_bool <- data.frame(matrix(0, nrow = sum(gene_filter)+1, ncol = length(MetaData$filename)+1))
intermed_gene_presence_absence_mtx_bool <- gene_presence_absence_mtx_bool[c(1,which(gene_filter==1)+1),]
```

# Compute consensus for GPSCs (columns in gene presence absence matrix will now not be isolates but GPSCs)
```{r}
consensus_intermed_gene_presence_absence_mtx_bool <- data.frame(matrix(0, nrow = sum(gene_filter)+1, ncol = length(unique(PopPUNK_clusters$Cluster))+1))
consensus_intermed_gene_presence_absence_mtx_bool[1,-1] <- paste("SeqCl_",unique(PopPUNK_clusters$Cluster),sep = "")

consensus_intermed_gene_presence_absence_mtx_bool_2001 <- data.frame(matrix(0, nrow = sum(gene_filter)+1, ncol = length(unique(PopPUNK_clusters$Cluster))+1))
consensus_intermed_gene_presence_absence_mtx_bool_2001[1,-1] <- paste("SeqCl_",unique(PopPUNK_clusters$Cluster),sep = "")

consensus_intermed_gene_presence_absence_mtx_bool_2004 <- data.frame(matrix(0, nrow = sum(gene_filter)+1, ncol = length(unique(PopPUNK_clusters$Cluster))+1))
consensus_intermed_gene_presence_absence_mtx_bool_2004[1,-1] <- paste("SeqCl_",unique(PopPUNK_clusters$Cluster),sep = "")

consensus_intermed_gene_presence_absence_mtx_bool_2007 <- data.frame(matrix(0, nrow = sum(gene_filter)+1, ncol = length(unique(PopPUNK_clusters$Cluster))+1))
consensus_intermed_gene_presence_absence_mtx_bool_2007[1,-1] <- paste("SeqCl_",unique(PopPUNK_clusters$Cluster),sep = "")


cons_genomes <- function(x){
  as.double(ceiling(median(as.integer(x))))
}

for (i in unique(PopPUNK_clusters$Cluster)) {
  consensus_intermed_gene_presence_absence_mtx_bool[-1,(which(unique(PopPUNK_clusters$Cluster) == i))+1] <- apply(as.matrix(intermed_gene_presence_absence_mtx_bool[-1,c(FALSE,GPSC_dict[unlist(intermed_gene_presence_absence_mtx_bool[1,-1])]==i)]), 1, cons_genomes)
  
  consensus_intermed_gene_presence_absence_mtx_bool_2001[-1,(which(unique(PopPUNK_clusters$Cluster) == i))+1] <- apply(as.matrix(intermed_gene_presence_absence_mtx_bool_2001[-1,c(FALSE,GPSC_dict[unlist(intermed_gene_presence_absence_mtx_bool_2001[1,-1])]==i)]), 1, cons_genomes)
  
    consensus_intermed_gene_presence_absence_mtx_bool_2004[-1,(which(unique(PopPUNK_clusters$Cluster) == i))+1] <- apply(as.matrix(intermed_gene_presence_absence_mtx_bool_2004[-1,c(FALSE,GPSC_dict[unlist(intermed_gene_presence_absence_mtx_bool_2004[1,-1])]==i)]), 1, cons_genomes)
    
    consensus_intermed_gene_presence_absence_mtx_bool_2007[-1,(which(unique(PopPUNK_clusters$Cluster) == i))+1] <- apply(as.matrix(intermed_gene_presence_absence_mtx_bool_2007[-1,c(FALSE,GPSC_dict[unlist(intermed_gene_presence_absence_mtx_bool_2007[1,-1])]==i)]), 1, cons_genomes)
}

consensus_intermed_gene_presence_absence_mtx_bool[-1,1] <- intermed_gene_presence_absence_mtx_bool[-1,1]
consensus_intermed_gene_presence_absence_mtx_bool_2001[-1,1] <- intermed_gene_presence_absence_mtx_bool[-1,1]
consensus_intermed_gene_presence_absence_mtx_bool_2004[-1,1] <- intermed_gene_presence_absence_mtx_bool[-1,1]
consensus_intermed_gene_presence_absence_mtx_bool_2007[-1,1] <- intermed_gene_presence_absence_mtx_bool[-1,1]

saveRDS(apply(as.matrix(consensus_intermed_gene_presence_absence_mtx_bool, ncol = GPSC_no)[-1,-1], c(1,2), as.double), file = "Genotypes_matrix.rds") # Genotypes matrix for model input
```

# Calculate delta statistic for ordering genes according to frequency changes (the less they change the more likely they are to be under NFDS)
```{r}
# calculate gene frequencies first, separate for pre and post vaccine (first and last time point in dataset)
gene_freq_pre <- apply(intermed_gene_presence_absence_mtx_bool_2001[-1,-1], 1, sum_as_int)
gene_freq_post <- apply(intermed_gene_presence_absence_mtx_bool_2007[-1,-1], 1, sum_as_int)

# first, calculate pre and post vacc frequencies of genes:
ggC_pre_vacc_gene_freq <- (gene_freq_pre) / (ncol(intermed_gene_presence_absence_mtx_bool_2001)-1)
ggC_post_vacc_gene_freq <- gene_freq_post / (ncol(intermed_gene_presence_absence_mtx_bool_2007)-1)

# calculate delta statistic (refer to Corander et al. for more info)
gene_delta_data <- (ggC_post_vacc_gene_freq - ggC_pre_vacc_gene_freq) ^ 2 / (1 - ggC_pre_vacc_gene_freq * (1 - ggC_pre_vacc_gene_freq))
gene_delta_ranking <- rank(gene_delta_data)

saveRDS(gene_delta_ranking, file = "gene_delta_ranking.rds") # ranking describes the over from very little frequency changes to large frequency changes of intermediate-frequency genes
```

# Calculate GPSC counts for each data time point (this is for fitting the model to data)
```{r}
#calculate the frequency of the gene clusters and year
GPSC_count_1 <- rep(0,GPSC_no)
GPSC_count_2 <- rep(0,GPSC_no)
GPSC_count_3 <- rep(0,GPSC_no)
for (i in 1:GPSC_no){
  GPSC_count_1[i] <- length(which(PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),]$"SeqYear"==2001))
  GPSC_count_2[i] <- length(which(PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),]$"SeqYear"==2004))
  GPSC_count_3[i] <- length(which(PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),]$"SeqYear"==2007))
}

saveRDS(GPSC_count_1, "GPSC_count_1.rds")
saveRDS(GPSC_count_2, "GPSC_count_2.rds")
saveRDS(GPSC_count_3, "GPSC_count_3.rds")

# create a dataframe that contains all these counts but also accounts for missing years (in this case we have data for 2001, 2004, and 2007, so we want to account for missing data for 2002, 2003, 2005, 2006, and we won't use the first time point for fitting, we've already used that in the definition of the start population)
empty_sample <- rep(NA, length(GPSC_count_2)) # accounts for missing years

joint_GPSC_counts <- cbind(empty_sample, empty_sample, GPSC_count_2, empty_sample, empty_sample, GPSC_count_3)
saveRDS(joint_GPSC_counts, "joint_GPSC_counts.rds")
```
