---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
### Loading packages
# install.packages("drat") # -- if you don't have drat installed
# drat:::add("ncov-ic")
# install.packages("odin.dust")
library(odin.dust)
#install.packages("mcstate")
library(mcstate)
library(coda)
library(ggplot2)
library(viridis)

# read in model
WF_PPxSero <- odin.dust::odin_dust("NFDS_Model.R")
```

### Data anlysis to be performed before using NFDS model
# 1) assemble genomes if necessary (e.g, using shovill 1.1.0 using default parameters (Torsten Seemann. Tseemann/Shovill. 2024)) 
# 2) Compute Global pneumococcal sequence clusters (GPSCs) using PopPUNK 2.6.0 (Lees et al. (2019), doi: 10.1101/gr.241455.118.) with the PopPUNK database for S. pneumoniae GPS_v4 via poppunk_assign, applying quality control (--run-qc), allowing maximally one distance of zero (--max-zero-dist 1), and merging a maximum of three clusters based on one single isolate (--max-merge 3). 
# 3) Construct a pangenome: 
# 3.1) ggCaller 1.3.4 \cite{Horsfield et al. (2023), doi: 10.1101/gr.277733.123} can be used for gene calling using the SPARC dataset for annotation (Croucher et al. (2015), doi: 10.1038/sdata.2015.58 and Croucher et al. (2016), doi: 10.5061/DRYAD.T55GQ) and with the options --gene-finding-only --annotation sensitive. 
# 3.2) bakta (Schwengers et al. (2021), doi: 10.1099/mgen.0.000685) with default parameters can be used for functional annotation (if wanted)
# 3.3) panaroo v1.5.2 ( Tonkin-Hill et al. (2020), doi: 10.1186/s13059-020-02090-4) can be run with the options --refind-mode off --clean-mode moderate --family_threshold 0.7 --min_trailing_support 2 --trailing_recursive 0 --min_edge_support_sv 2 --edge_support_threshold 0.0 --aligner none


```{r}
# Set parameters for model
vacc_time <- 4 # time of vaccination, in years after start of dataset

PPsero_startpop <- readRDS("PPsero_startpop.rds") # start population for model (GPSC x serotypes)
# rows are GPSCs, columns are serotypes
# based on real data (counts) from first time point of dataset
# created by Poisson sampling from real data and scaling it up to model population size (15,000)
# GPSC-serotype combinations that appear in the dataset - but not in the first time point - are set to 1

# freq_clust
no_GPSC <- nrow(PPsero_startpop) # number of GPSCs
#freq_sero 
no_sero <- ncol(PPsero_startpop) # number of serotypes
  
Genotypes_matrix <- readRDS("Genotypes_matrix.rds")
# genotype matrix is constant over time
# prepare genotype matrix by:
# 1) binarise panaroo's pangenome matrix (replace all empty spaces by 0 (=not present), all gene calls by 1 (= present))
# 2) filter for genes that are present at 5-95% frequency (intermediate-frequency genes)
# 3) summarise information by GPSC: take all columns that correspond to one GPSC, take the median and then round the value (to 0 or 1), to determine which genes are present in which GPSC

delta_test <- readRDS("gene_delta_ranking.rds") # delta statistic (computed as in Corander et al.)
# calculates the changes in gene frequencies of the first time point (which are assumed to be at equilibrium) to the last time point
# Genes that change the least in frequency have a low delta statistic value, genes that change more have a higher value.
# then apply rank() function to determine order of genes
# (NFDS model will determine a cut-off for genes to be under NFDS or not based on this order)
gene_no_test <- length(delta_test) # number of intermediate-frequency genes
migMatr_test_mtx <- readRDS("PPsero_mig.rds") # immigration matrix (rows are GPSCs, columns are serotypes)
# compute immigration matrix by assigning all GPSC-serotype combinations that appear at least once in the data set equal probabilities of immigration (i.e. one over the number of GPSC-serotype combinations)
vaccTypes_test <- readRDS(file = "SeroVT.rds") # information on which types are affected by vaccine (1 = serotype affected by vaccine, 0 = serotypes not affected)
 
dt_test <- 1/12 # determines how many generations there are between two data points (here: 12 generations per year)
```

```{r}
### simulate model 
PPxSero_testparams <- list(dt = dt_test, species_no = no_GPSC, sero_no = no_sero, gene_no = gene_no_test, Pop_ini = PPsero_startpop, Pop_eq = rowSums(PPsero_startpop), capacity = sum(PPsero_startpop), Genotypes = Genotypes_matrix, sigma_f = log(0.0345), prop_f = 0.3, delta = (delta_test), m = log(0.0133), migVec = migMatr_test_mtx, vaccTypes = vaccTypes_test, v = 0.0814, vacc_time = 4)

WFmodel_ppxSero <- WF_PPxSero$new(pars = PPxSero_testparams,
                         time = 0,
                         n_particles = 10L,
                         n_threads = 4L,
                         seed = 1L, deterministic = TRUE) # can also run this in stochastic mode with deterministic = FALSE

simulated_data_long <- list()
simulated_sum_data_long <- list()
list_of_particle_list <- list()
for (i in 0:240) {
  simMean_all <- rowMeans(WFmodel_ppxSero$run(i)[-(1:(no_GPSC+1)),])
  simMean_matrix <- matrix(simMean_all, ncol = no_sero, nrow = no_GPSC, byrow = FALSE)
  simulated_data_long[[i+1]] <- simMean_matrix
  simSum_all <- rowSums(WFmodel_ppxSero$run(i)[-(1:(no_GPSC+1)),])
  simSum_matrix <- matrix(simSum_all, ncol = no_sero, nrow = no_GPSC, byrow = FALSE)
  simulated_sum_data_long[[i+1]] <- simSum_matrix
  list_of_particle_list[[i+1]] <-  WFmodel_ppxSero$run(i)[(2:(no_GPSC+1)),]
}
simulated_data_over_time_clust <- sapply(simulated_data_long, rowSums)
simulated_data_over_time_clust_df <- as.data.frame(simulated_data_over_time_clust)

# only save one "sample" per year
simulated_data_over_time_filtered <- simulated_data_over_time_clust[,((1:20)*12)+1]
simulated_data_over_time_filtered_df <- as.data.frame(rbind(1:20,simulated_data_over_time_filtered))
saveRDS(simulated_data_over_time_filtered, "simulated_data_over_time_filtered.rds") # we will use this for fitting below

rnd_colours <- c("#a7622a", "#e54828", "#864291", "#5d977b", "#20e1b1", "#f64b13", "#43237b","#9a2753","#327a78","#913b94","#380b82","#a9ab59","#858202","#177e4d","#1e6b7a","#a57771","#955e9b","#54a291","#574116","#1786a7","black","red","yellow")
plot(1:20, simulated_data_over_time_filtered_df[2,], type = "l", col = rnd_colours[1], ylim = c(0, 1500))
for (i in 3:23) {
  lines(1:20, simulated_data_over_time_filtered_df[i,], col = rnd_colours[i])
}
```

```{r}
### fit model to simulated data (23 GPSCs for 20 years)
simulated_data <- readRDS("simulated_data_over_time_filtered.rds")

# or if you want to use the data that you processed in the ProcessData.Rmd script
simulated_data <- readRDS("joint_GPSC_counts.rds")

fitting_test_data <- data.frame("year" = 1:ncol(simulated_data), t(simulated_data))
names(fitting_test_data) <- c("year", as.character(1:(no_GPSC)))
# prepare data for mcstate:
fitting_test_data_2 <- mcstate::particle_filter_data(data = fitting_test_data,
                                                     time = "year",
                                                     rate = 1 / dt_test,
                                                     initial_time = 0)
```

```{r}
# define likelihood (multinomial likelihood)
# compares GPSC counts from (simulated) data to GPSC frequencies in model
combined_compare <- function(state, observed, pars = NULL) {
  result <- 0
  data_size <- sum(unlist(observed[as.character(1:(length(unlist(observed))-4))]))
  model_size = sum(unlist(state[-1, , drop = TRUE]))
  exp_noise <- 1e6
  data_vals <- unlist(observed[as.character(1:(length(unlist(observed))-4))])
  model_vals <- rep(0, length(unlist(observed))-4)
  data_missing <- FALSE
  for (i in 1:(length(unlist(observed))-4)){ 
    state_name <- paste("sum_clust", i, sep = "")
    model_vals[i] <- state[state_name, , drop = TRUE]
    if (is.na(observed[[as.character(i)]])) {
      data_missing <- TRUE
    } 
  }
  models_vals_err <- model_vals + rexp(n = length(model_vals), rate = exp_noise)
  if(data_missing){
    ll_obs <- 0
  }
  else{
    ll_obs <- dmultinom(x = (data_vals), prob = models_vals_err/model_size, log = TRUE)   
  }
  result <- ll_obs
  result
}
```

```{r}
# define fitting function using two mcmc runs

Fit_model_to_sim_data <- function(fitting_data, fitting_params, output_loc){
  print(paste(output_loc, "/det_mcmc2.rds", sep = ""))
  WF <- odin.dust::odin_dust("NFDS_Model.R")

  det_filter <- particle_deterministic$new(data = fitting_data,
                                           model = WF,
                                           compare = combined_compare)
  
  complex_params = fitting_params

  
  make_transform <- function(m) {
    function(theta) {
      as_double_mtx <- function(x){
        sapply(x,as.double)
      }
      c(lapply(m, as_double_mtx), as.list(theta))
    }
  }
  
  # read this: https://mrc-ide.github.io/mcstate/reference/pmcmc_parameters.html
  # it explains how to not fit all parameters but just the ones I want
  # non-scalar parameters have to be transformed for this.
  
  index <- function(info) {
    list(run = c(sum_clust = info$index$Pop_tot),
         state = c(Pop = info$index$Pop))
  }

  proposal_matrix <- diag(c(exp(1), 0.1, exp(1), 0.1))
  
  mcmc_pars <- mcstate::pmcmc_parameters$new(list(mcstate::pmcmc_parameter("sigma_f", runif(n=1, min=-10, max=0), min = -1000, max = 0, prior = function(a) 1/a), mcstate::pmcmc_parameter("prop_f", runif(n=1, min=0, max=1), min = 0, max = 1, prior = function(a) a), mcstate::pmcmc_parameter("m", runif(n=1, min=-10, max=0), min = -1000, max = 0, prior = function(a) 1/a), mcstate::pmcmc_parameter("v", runif(n=1, min=0, max=1), min = 0, max = 1, prior = function(a) a)), proposal_matrix, make_transform(complex_params))
  
  mcmc_pars$initial()
  
  det_filter <- particle_deterministic$new(data = fitting_data,
                                           model = WF,
                                           index = index,
                                           compare = combined_compare)
  
  n_steps <- 5 # this is a work-around (do a non-parallel run for five steps, which sets up the model and ensures that the parallel mcmc runs)
  n_burnin <- 0
  
  
  control <- mcstate::pmcmc_control(
    n_steps,
    save_state = TRUE, 
    save_trajectories = TRUE,
    progress = TRUE,
    adaptive_proposal = TRUE,
    n_chains = 1)
  det_pmcmc_run <- mcstate::pmcmc(mcmc_pars, det_filter, control = control)
  
  #n_steps <- 1000
  n_steps <- 10
  n_burnin <- 0
  
  
  control <- mcstate::pmcmc_control(
    n_steps,
    save_state = TRUE, 
    save_trajectories = TRUE,
    progress = TRUE,
    adaptive_proposal = TRUE,
    n_chains =4, n_workers = 4,
    n_threads_total = 4)
  
  
  det_pmcmc_run <- mcstate::pmcmc(mcmc_pars, det_filter, control = control)
  processed_chains <- mcstate::pmcmc_thin(det_pmcmc_run, burnin = 20, thin = 1)
  parameter_mean_hpd <- apply(processed_chains$pars, 2, mean)
  print(parameter_mean_hpd)
  
  det_mcmc1 <- coda::as.mcmc(cbind(det_pmcmc_run$probabilities, det_pmcmc_run$pars))
  print("det_mcmc_1 final log likelihood")
  processed_chains$probabilities[nrow(processed_chains$probabilities),2]
  print("det_mcmc_1 mean log likelihood")
  mean(processed_chains$probabilities[,2])
  det_proposal_matrix <- cov(processed_chains$pars)
  
  det_mcmc_pars <- mcstate::pmcmc_parameters$new(list(mcstate::pmcmc_parameter("sigma_f",  parameter_mean_hpd[1], min = -1000, max = 0, prior = function(a) 1/a), mcstate::pmcmc_parameter("prop_f",  parameter_mean_hpd[2], min = 0, max = 1, prior = function(a) a), mcstate::pmcmc_parameter("m",  parameter_mean_hpd[3], min = -1000, max = 0, prior = function(a) 1/a), mcstate::pmcmc_parameter("v",  parameter_mean_hpd[4], min = 0, max = 1, prior = function(a) a)), proposal_matrix, make_transform(complex_params))
  

  det_filter <- particle_deterministic$new(data = fitting_data,
                                           model = WF,
                                           index = index,
                                           compare = combined_compare)
  
  n_steps <- 5
  n_burnin <- 0
  
  
  control <- mcstate::pmcmc_control(
    n_steps,
    save_state = TRUE, 
    save_trajectories = TRUE,
    progress = TRUE,
    adaptive_proposal = TRUE,
    n_chains = 1)
  det_pmcmc_run <- mcstate::pmcmc(det_mcmc_pars, det_filter, control = control)
  
  # n_steps <- 20000 # good amount of steps but should be done on a cluster
  n_steps <- 50
  n_burnin <- 0
  
  
  control <- mcstate::pmcmc_control(
    n_steps,
    save_state = TRUE, 
    save_trajectories = TRUE,
    progress = TRUE,
    adaptive_proposal = TRUE,
    n_chains = 4, n_workers = 4, n_threads_total = 4)
  det_pmcmc_run2 <- mcstate::pmcmc(det_mcmc_pars, det_filter, control = control)
  processed_chains <- mcstate::pmcmc_thin(det_pmcmc_run2, burnin = 100, thin = 1)
  parameter_mean_hpd <- apply(processed_chains$pars, 2, mean)
  print(parameter_mean_hpd)
  par(mfrow = c(1,1))
  
  det_mcmc2 <- coda::as.mcmc(cbind(processed_chains$probabilities, processed_chains$pars))
  percentile95_low <- apply(det_mcmc2,2,quantile,.025)
  percentile95_up <- apply(det_mcmc2,2,quantile,.975)
  print("lower 95 percentile")
  print(percentile95_low)
  print("upper 95 percentile")
  print(percentile95_up)

  print("det_mcmc_2 final log likelihood")
  print(processed_chains$probabilities[nrow(processed_chains$probabilities),2])
  print("det_mcmc_2 mean log likelihood")
  print(mean(processed_chains$probabilities[,2]))
  
  # save MCMC values in rds file
  if (file.exists(output_loc)){
    saveRDS(det_mcmc2, paste(output_loc, "/det_mcmc2.rds", sep = ""))
    pdf(file = paste(output_loc, "/det_mcmc2.pdf", sep = ""),
      width = 6, # The width of the plot in inches
      height = 12)
    plot(det_mcmc2)
    dev.off()
  }
  else{
    print("Warning: output location appears not to exist. Cannot write output files.")
  }
  return(det_pmcmc_run2)
}
```

```{r}
# run mcmc 
fitting_test_params <- list(species_no = no_GPSC, Pop_ini = data.frame(PPsero_startpop), Pop_eq = rowSums(PPsero_startpop), Genotypes = as.data.frame(Genotypes_matrix), capacity = sum(PPsero_startpop), delta = delta_test, vaccTypes = vaccTypes_test, gene_no = gene_no_test, vacc_time = 4, dt = dt_test, pmcmc_sigma_w = -1000, migVec = data.frame(migMatr_test_mtx), sero_no = no_sero)

MCMC_chain_sample1year_run <- Fit_model_to_sim_data(fitting_data = fitting_test_data_2, fitting_params = fitting_test_params, output_loc = "./MCMC_output")
# this will take some time (15-20 mins) 
# for quicker runs, reduce n_steps 
# to get good fits, this should be run for much longer

processed_chains_1year <- mcstate::pmcmc_thin(MCMC_chain_sample1year_run, burnin = 200, thin = 2)
MCMC_chain_sample1year <- coda::as.mcmc(cbind(processed_chains_1year$probabilities, processed_chains_1year$pars))

plot(MCMC_chain_sample1year)
```

