---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
# read in pre-processed fits for simulation (create three independent fits, then calculated parameter mean and 95-percentiles)

sim_fit_m_1 <- read.csv("joined_mcmc_param_fits_sampleSize3_m_2.csv")[,-1]
sim_fit_m_2 <- read.csv("joined_mcmc_param_fits_sampleSize3_m_2_v2.csv")[,-1]
sim_fit_m_3 <- read.csv("joined_mcmc_param_fits_sampleSize3_m_2_v3.csv")[,-1]

sim_fit_v_1 <- read.csv("joined_mcmc_param_fits_sampleSize3_v_2.csv")[,-1]
sim_fit_v_2 <- read.csv("joined_mcmc_param_fits_sampleSize3_v_2_v2.csv")[,-1]
sim_fit_v_3 <- read.csv("joined_mcmc_param_fits_sampleSize3_v_2_v3.csv")[,-1]

sim_fit_sigmaf_1 <- read.csv("joined_mcmc_param_fits_sampleSize3_sigmaf_2.csv")[,-1]
sim_fit_sigmaf_2 <- read.csv("joined_mcmc_param_fits_sampleSize3_sigmaf_2_v2.csv")[,-1]
sim_fit_sigmaf_3 <- read.csv("joined_mcmc_param_fits_sampleSize3_sigmaf_2_v3.csv")[,-1]

sim_fit_propf_1 <- read.csv("joined_mcmc_param_fits_sampleSize3_propf_2.csv")[,-1]
sim_fit_propf_2 <- read.csv("joined_mcmc_param_fits_sampleSize3_propf_2_v2.csv")[,-1]
sim_fit_propf_3 <- read.csv("joined_mcmc_param_fits_sampleSize3_propf_2_v3.csv")[,-1]
```

```{r}
# save all three replicates in one data.frame per parameter
sim_fit_m_mean <- cbind(sim_fit_m_1$params, sim_fit_m_1$population, c(sim_fit_m_1$mean, sim_fit_m_2$mean, sim_fit_m_3$mean), c(sim_fit_m_1$lower, sim_fit_m_2$lower, sim_fit_m_3$lower), c(sim_fit_m_1$upper, sim_fit_m_2$upper, sim_fit_m_3$upper), sim_fit_m_1$budget)
sim_fit_v_mean <- cbind(sim_fit_v_1$params, sim_fit_v_1$population, c(sim_fit_v_1$mean, sim_fit_v_2$mean, sim_fit_v_3$mean), c(sim_fit_v_1$lower, sim_fit_v_2$lower, sim_fit_v_3$lower), c(sim_fit_v_1$upper, sim_fit_v_2$upper, sim_fit_v_3$upper), sim_fit_v_3$budget)
sim_fit_sigmaf_mean <- cbind(sim_fit_sigmaf_1$params, sim_fit_sigmaf_1$population, c(sim_fit_sigmaf_1$mean, sim_fit_sigmaf_2$mean, sim_fit_sigmaf_3$mean), c(sim_fit_sigmaf_1$lower, sim_fit_sigmaf_2$lower, sim_fit_sigmaf_3$lower), c(sim_fit_sigmaf_1$upper, sim_fit_sigmaf_2$upper, sim_fit_sigmaf_3$upper), sim_fit_sigmaf_1$budget)
sim_fit_propf_mean <- cbind(sim_fit_propf_1$params, sim_fit_propf_1$population, c(sim_fit_propf_1$mean, sim_fit_propf_2$mean, sim_fit_propf_3$mean), c(sim_fit_propf_1$lower, sim_fit_propf_2$lower, sim_fit_propf_3$lower), c(sim_fit_propf_1$upper, sim_fit_propf_2$upper, sim_fit_propf_3$upper), sim_fit_propf_1$budget)

colnames(sim_fit_m_mean) <- colnames(sim_fit_m_1)[1:6]
colnames(sim_fit_v_mean) <- colnames(sim_fit_m_1)[1:6]
colnames(sim_fit_sigmaf_mean) <- colnames(sim_fit_m_1)[1:6]
colnames(sim_fit_propf_mean) <- colnames(sim_fit_m_1)[1:6]
```

# next steps: make new plot with means
# calculate mean of of means and credible intervals
```{r}
sim_fit_m_mean_df <- data.frame(sim_fit_m_mean)
sim_fit_m_mean_df <- aggregate(. ~ params + population, data = sim_fit_m_mean_df, FUN = mean)
#sim_fit_m_mean_df[,3:5] <- exp(sim_fit_m_mean_df[,3:5])
sim_fit_m_mean_df$budget <- factor(sim_fit_m_mean_df$budget)
sim_fit_m_mean_df$population <- factor(sim_fit_m_mean_df$population)
sim_fit_m_mean_df$params <- factor(sim_fit_m_mean_df$params)

sim_fit_v_mean_df <- data.frame(sim_fit_v_mean)
sim_fit_v_mean_df <- aggregate(. ~ params + population, data = sim_fit_v_mean_df, FUN = mean)
sim_fit_v_mean_df$budget <- factor(sim_fit_v_mean_df$budget)
sim_fit_v_mean_df$population <- factor(sim_fit_v_mean_df$population)
sim_fit_v_mean_df$params <- factor(sim_fit_v_mean_df$params)

sim_fit_sigmaf_mean_df <- data.frame(sim_fit_sigmaf_mean)
sim_fit_sigmaf_mean_df <- aggregate(. ~ params + population, data = sim_fit_sigmaf_mean_df, FUN = mean)
sim_fit_sigmaf_mean_df[,3:5] <- exp(sim_fit_sigmaf_mean_df[,3:5])
sim_fit_sigmaf_mean_df$budget <- factor(sim_fit_sigmaf_mean_df$budget)
sim_fit_sigmaf_mean_df$population <- factor(sim_fit_sigmaf_mean_df$population)
sim_fit_sigmaf_mean_df$params <- factor(sim_fit_sigmaf_mean_df$params)

sim_fit_propf_mean_df <- data.frame(sim_fit_propf_mean)
sim_fit_propf_mean_df <- aggregate(. ~ params + population, data = sim_fit_propf_mean_df, FUN = mean)
sim_fit_propf_mean_df$budget <- factor(sim_fit_propf_mean_df$budget)
sim_fit_propf_mean_df$population <- factor(sim_fit_propf_mean_df$population)
sim_fit_propf_mean_df$params <- factor(sim_fit_propf_mean_df$params)
```


```{r}
library(viridis)
library(ggplot2)

output_location <- "/Users/llorenz/Documents/PhD_Project/Papers/NFDS_Model/Figures/Data_for_figures/"

# vaccine parameter
real_val_v <- 0.0814

ggplot(sim_fit_v_mean_df,aes(x=params,y=mean, colour = budget, group = interaction(population, params))) + 
    geom_hline(aes(yintercept = real_val_v), colour = "red", linetype = "dashed", linewidth = 1.5) +
  geom_errorbar(aes(ymin=lower,ymax=upper,color=budget,width=0.2), linewidth = 1) +
  geom_point(size = 4) + scale_color_viridis_d(option = "turbo",direction = -1) +
  facet_wrap(~population, ncol = 4) + coord_cartesian(ylim = c(0,0.2)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Parameter estimate", x = "Population")

v_plot <- ggplot(sim_fit_v_mean_df,aes(x=population,y=mean, colour = population, group = interaction(population, params))) + 
  geom_hline(aes(yintercept = real_val_v), colour = "red", linetype = "dashed", linewidth = 1.5) +
  geom_errorbar(aes(ymin=lower,ymax=upper,color=population,width=0.4), linewidth = 1) +
  geom_point(size = 3) + scale_color_viridis_d(option = "turbo",direction = -1) +
  facet_wrap(~params, ncol = 8) + coord_cartesian(ylim = c(0,0.2)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Parameter estimate", x = "Population")

ggsave(v_plot, 
    filename = paste(output_location,"Fig4_byBudget_v_v1.pdf",sep = ""),   # The directory and file name you want to save the file in
    device = "pdf",
    width = 12, # The width of the plot in inches
    height = 4)

# immigration parameter
real_val_m <- 0.0133

m_plot <- ggplot(sim_fit_m_mean_df,aes(x=population,y=mean, colour = population, group = interaction(population, params))) + 
  geom_hline(aes(yintercept = real_val_m), colour = "red", linetype = "dashed", linewidth = 1.5) +
  geom_errorbar(aes(ymin=lower,ymax=upper,color=population,width=0.4), linewidth = 1) +
  geom_point(size = 3) + scale_color_viridis_d(option = "turbo",direction = -1) +
  facet_wrap(~params, ncol = 8) + coord_cartesian(ylim = c(0,0.1)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Parameter estimate", x = "Population")

ggsave(m_plot, 
    filename = paste(output_location,"Fig4_byBudget_m_v1.pdf",sep = ""),   # The directory and file name you want to save the file in
    device = "pdf",
    width = 12, # The width of the plot in inches
    height = 4)

# sigma_f
real_val_sigmaf <- 0.0345

sigmaf_plot <- ggplot(sim_fit_sigmaf_mean_df,aes(x=population,y=mean, colour = population, group = interaction(population, params))) + 
  geom_hline(aes(yintercept = real_val_sigmaf), colour = "red", linetype = "dashed", linewidth = 1.5) +
  geom_errorbar(aes(ymin=lower,ymax=upper,color=population,width=0.4), linewidth = 1) +
  geom_point(size = 3) + scale_color_viridis_d(option = "turbo",direction = -1) +
  facet_wrap(~params, ncol = 8) + coord_cartesian(ylim = c(0,0.2)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Parameter estimate", x = "Population")

ggsave(sigmaf_plot, 
    filename = paste(output_location,"Fig4_byBudget_sigmaf_v1.pdf",sep = ""),   # The directory and file name you want to save the file in
    device = "pdf",
    width = 12, # The width of the plot in inches
    height = 4)

# prop_f
real_val_prop_f <- 0.3
  
propf_plot <- ggplot(sim_fit_propf_mean_df,aes(x=population,y=mean, colour = population, group = interaction(population, params))) + 
  geom_hline(aes(yintercept = real_val_prop_f), colour = "red", linetype = "dashed", linewidth = 1.5) +
  geom_errorbar(aes(ymin=lower,ymax=upper,color=population,width=0.4), linewidth = 1) +
  geom_point(size = 3) + scale_color_viridis_d(option = "turbo",direction = -1) +
  facet_wrap(~params, ncol = 8) + coord_cartesian(ylim = c(0,0.5)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Parameter estimate", x = "Population")

ggsave(propf_plot, 
    filename = paste(output_location,"Fig4_byBudget_propf_v1.pdf",sep = ""),   # The directory and file name you want to save the file in
    device = "pdf",
    width = 12, # The width of the plot in inches
    height = 4)
```


# make joint df with means of all params
# use this df for running simulation analysis on cluster (simulate model forward with fitted parameters and then compare to original trajectory)

```{r}
sim_fit_param_means <- cbind(sim_fit_m_mean[,1:3], sim_fit_v_mean[,3], sim_fit_sigmaf_mean[,3], sim_fit_propf_mean[,3])
colnames(sim_fit_param_means) <- c(colnames(sim_fit_m_mean)[1:2], "m", "v", "sigmaf", "propf")
#write.csv(sim_fit_param_means, "parameter_means_fits_to_simulation.csv", row.names = TRUE)
write.table(sim_fit_param_means, "parameter_means_fits_to_simulation.tsv", row.names = TRUE, sep='\t', col.names = FALSE, quote=FALSE)

# make version that only stores budget and sample freq and run number
# for new version of forward simulation "SimulateModelForSimFits_ParamDistribution.R"
sim_fit_budget <- read.table("parameter_means_fits_to_simulation.tsv")[,2:3]
sim_fit_budget[,3] <- sort(rep(1:3, 8))
colnames(sim_fit_budget) <- c("params", "population", "replicate")
write.table(sim_fit_budget, "sim_fit_budget_fits_to_simulation.tsv", row.names = TRUE, sep='\t', col.names = FALSE, quote=FALSE)
```


### analyse forward simulation here
```{r}
path_to_files <- "/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_10_21/SimModelFromSimFits/"

SimFromSimFits_list <- vector(mode = "list", length = 96)

for (i in 1:96) {
  file_name <- paste(as.character(i), "_params_4_model_data_rel.rds", sep = "")
  file_content <- readRDS(paste(path_to_files, file_name, sep = ""))
  SimFromSimFits_list[[i]] <- file_content # rows=years (20), columns=GPSCs (23)
}
```

# now, compare these results to the original simulated data using Mean squared error (mse)
```{r}
orig_simulated_data <- t(readRDS("simulated_data_over_time_filtered.rds"))
orig_simulated_data_normalised <- orig_simulated_data/rowSums(orig_simulated_data)

#mean((orig_simulated_data[1,]/sum(orig_simulated_data[1,]) - SimFromSimFits_list[[1]][1,])^2)
#mean((orig_simulated_data[2,] - SimFromSimFits_list[[1]][2,])^2)

mse_SimFromSimFits <- rep(NA, 96)
for (i in 1:96) {
  mse_SimFromSimFits[i] <- sum(rowMeans((orig_simulated_data_normalised - SimFromSimFits_list[[i]])^2))
  #mse_SimFromSimFits[i] <- sum(rowMeans((orig_simulated_data_normalised[1:10,] - SimFromSimFits_list[[i]][1:10,])^2))
}

# create df
sampling_strat_order = read.table("parameter_means_fits_to_simulation.tsv", header=FALSE)[,(1:3)]
colnames(sampling_strat_order) <- c("index", "budget","sample_freq")

mse_SimFromSimFits_df <- cbind(sampling_strat_order, mse_SimFromSimFits)
colnames(mse_SimFromSimFits_df) <- c("index", "budget","sample_freq", "MSE")
mse_SimFromSimFits_df$budget <- as.factor(mse_SimFromSimFits_df$budget)
mse_SimFromSimFits_df$sample_freq <- as.factor(mse_SimFromSimFits_df$sample_freq)

plot(mse_SimFromSimFits[which(sampling_strat_order$sample_freq == 1)], ylim = c(0, 0.015))
plot(mse_SimFromSimFits[which(sampling_strat_order$sample_freq == 2)], ylim = c(0, 0.015))
plot(mse_SimFromSimFits[which(sampling_strat_order$sample_freq == 3)], ylim = c(0, 0.015))
plot(mse_SimFromSimFits[which(sampling_strat_order$sample_freq == 4)], ylim = c(0, 0.015))
```

```{r}
ggplot(mse_SimFromSimFits_df,aes(x=budget,y=MSE, colour = budget, group = interaction(sample_freq, budget))) + 
  geom_point(size = 4) + scale_color_viridis_d(option = "turbo",direction = -1) +
  facet_wrap(~sample_freq, ncol = 4) + coord_cartesian(ylim = c(0,0.02)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Mean squared error", x = "Sampling budget") 

ggplot(mse_SimFromSimFits_df,aes(x=budget,y=MSE, colour = budget, group = interaction(sample_freq, budget))) + 
  geom_point(size = 4) + scale_color_viridis_d(option = "turbo",direction = -1) +
  facet_wrap(~sample_freq, ncol = 4) + coord_cartesian(ylim = c(0,0.01)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Mean squared error", x = "Sampling budget") 

ggplot(mse_SimFromSimFits_df,aes(x=budget,y=MSE, colour = sample_freq, group = interaction(sample_freq))) + 
  geom_point(size = 4) + scale_color_viridis_d(option = "turbo",direction = -1) + coord_cartesian(ylim = c(0,0.02)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Mean squared error", x = "Sampling budget") 
```

```{r}
plot(orig_simulated_data_normalised[1,], ylim = c(0, 0.2))
for (i in 2:20) {
  points((1:23) + (i-1)/25, orig_simulated_data_normalised[i,])
}
for (i in 1:20) {
  points((1:23) + (i)/25, SimFromSimFits_list[[8]][i,], col = "red")
}

# plot simulated data and simulation from fits
for (j in 1:23) {
  plot(orig_simulated_data_normalised[,j], ylim = c(0,.2), lty = 1, type = "l", lwd = 3, main = paste("cluster", as.character(j)))
for (i in 1:24) {
  points(SimFromSimFits_list[[i]][,j], col = colors()[i*10], pch = 19)
}
}

```

# try cumulative mse?
```{r}
mse_SimFromSimFits_cumulative <- matrix(NA, nrow = 96, ncol = 20)
for (i in 1:96) {
  mse_SimFromSimFits_cumulative[i,] <- cumsum(rowMeans((orig_simulated_data_normalised - SimFromSimFits_list[[i]])^2))
  #mse_SimFromSimFits[i] <- sum(rowMeans((orig_simulated_data_normalised[1:10,] - SimFromSimFits_list[[i]][1:10,])^2))
}

plot(mse_SimFromSimFits_cumulative[1,])
for (i in 2:8) {
  points(mse_SimFromSimFits_cumulative[i,])
}

mse_SimFromSimFits_df_cumu <- data.frame(cbind(rep(sampling_strat_order$index,20), rep(sampling_strat_order$budget,20), rep(sampling_strat_order$sample_freq,20),as.vector(mse_SimFromSimFits_cumulative)), sort(rep(1:20, 96)))
colnames(mse_SimFromSimFits_df_cumu) <- c("index", "budget","sample_freq", "MSE", "Year")
mse_SimFromSimFits_df_cumu$budget <- as.factor(mse_SimFromSimFits_df_cumu$budget)
mse_SimFromSimFits_df_cumu$sample_freq <- as.factor(mse_SimFromSimFits_df_cumu$sample_freq)

ggplot(mse_SimFromSimFits_df_cumu,aes(x=Year,y=MSE, colour = budget, group = interaction(sample_freq, budget))) + 
  geom_point(size = 4) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  facet_wrap(~sample_freq, ncol = 4) + coord_cartesian(ylim = c(0,0.02)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Mean squared error", x = "Sampling budget") 

# maybe take mean then and not three replicates?
mse_SimFromSimFits_df_cumu_mean <- aggregate(MSE~budget+sample_freq+Year,mse_SimFromSimFits_df_cumu, mean)

ggplot(mse_SimFromSimFits_df_cumu_mean,aes(x=Year,y=MSE, colour = budget, group = interaction(sample_freq, budget))) + 
  geom_point(size = 4) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  facet_wrap(~sample_freq, ncol = 4) + coord_cartesian(ylim = c(0,0.02)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Sampling budget") 

ggplot(mse_SimFromSimFits_df_cumu_mean,aes(x=Year,y=MSE, colour = budget, group = interaction(sample_freq, budget))) + 
  geom_point(size = 3) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  facet_wrap(~sample_freq, ncol = 4) + coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")

ggplot(mse_SimFromSimFits_df_cumu_mean,aes(x=Year,y=MSE, colour = budget, group = interaction(sample_freq, budget))) + 
  geom_line(linewidth = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  facet_wrap(~sample_freq, ncol = 4) + coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")


# make plot that includes all years
ggplot(mse_SimFromSimFits_df_cumu_mean,aes(x=Year,y=MSE, colour = budget, linetype=sample_freq, group = interaction(sample_freq, budget))) + 
  geom_line(linewidth = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")

ggplot(mse_SimFromSimFits_df_cumu_mean,aes(x=Year,y=MSE, colour = budget, shape=sample_freq, group = interaction(sample_freq, budget))) + 
  geom_jitter(size = 2, width = 0.25) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  coord_cartesian(ylim = c(0,0.004)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")

ggplot(mse_SimFromSimFits_df_cumu_mean,aes(x=Year,y=MSE, colour = sample_freq, group = interaction(sample_freq, budget))) + 
  geom_line(linewidth = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  facet_wrap(~budget, ncol = 4) + coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")

# these plots suggest that there is no major difference between sampling frequencies
# but they only take the mean parameter estimate into account, not the credible intervals
```

# repeat analysis for simulation with posterior parameter distribution
```{r}
path_to_files_paramdist <- "/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_10_22/SimModelFromSimFits_ParamDistribution/"

SimFromSimFitsParamDist_list <- vector(mode = "list", length = 96) # 96 files, representing 8 budgets, 4 sampling frequencies and 3 replicates each

for (i in 1:96) {
  file_name <- paste(as.character(i), "_params_4_model_data_rel_paramdist.rds", sep = "")
  file_content <- readRDS(paste(path_to_files_paramdist, file_name, sep = ""))
  SimFromSimFitsParamDist_list[[i]] <- file_content # rows=years (20) * random draws (200), columns=GPSCs (23)
}
```

# cumulative mse
```{r}
mse_SimFromSimFits_cumulative_paramDist <- matrix(NA, nrow = 96 * 200, ncol = 20) # rows = 96 runs, 200 samples per run, cols = 20 years
for (i in 1:96) {
  for (j in 1:200) {
    mse_SimFromSimFits_cumulative_paramDist[(i-1) * 200 + j,] <- cumsum(rowMeans((orig_simulated_data_normalised - SimFromSimFitsParamDist_list[[i]][(1:20)+20*(j-1),])^2))
  }
}

# now take mean of simulation replicates
mse_SimFromSimFits_cumulative_paramDist_mean <- matrix(NA, nrow = 96, ncol = 20) # rows = 96 runs, cols = 20 years
for (i in 1:96) {
    mse_SimFromSimFits_cumulative_paramDist_mean[i,] <- colMeans(mse_SimFromSimFits_cumulative_paramDist[(i-1)*200 + (1:200),])
}


mse_SimFromSimFits_paramDist_df_cumu <- data.frame(cbind(rep(sampling_strat_order$index,20), rep(sampling_strat_order$budget,20), rep(sampling_strat_order$sample_freq,20),as.vector(mse_SimFromSimFits_cumulative_paramDist_mean)), sort(rep(1:20, 96)), as.factor(sort(rep(1:3, 32))))
colnames(mse_SimFromSimFits_paramDist_df_cumu) <- c("index", "budget","sample_freq", "MSE", "Year", "Replicate")
mse_SimFromSimFits_paramDist_df_cumu$budget <- as.factor(mse_SimFromSimFits_paramDist_df_cumu$budget)
mse_SimFromSimFits_paramDist_df_cumu$sample_freq <- as.factor(mse_SimFromSimFits_paramDist_df_cumu$sample_freq)


ggplot(mse_SimFromSimFits_paramDist_df_cumu,aes(x=Year,y=MSE, colour = budget, group = interaction(sample_freq, budget, Replicate))) + 
  geom_line(linewidth = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  facet_wrap(~sample_freq, ncol = 4) + coord_cartesian(ylim = c(0,0.02)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Mean squared error", x = "Sampling budget") 
# plot shows nicely that there is not a lot of variation between three fits --> will move forward with mean over 3 reps

# take mean over fitting replicates
mse_SimFromSimFits_paramDist_df_cumu_mean <- aggregate(MSE~budget+sample_freq+Year,mse_SimFromSimFits_paramDist_df_cumu, mean)

ggplot(mse_SimFromSimFits_paramDist_df_cumu_mean,aes(x=Year,y=MSE, colour = budget, group = interaction(sample_freq, budget))) + 
  geom_line(linewidth = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  facet_wrap(~sample_freq, ncol = 4) + coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")


ggplot(mse_SimFromSimFits_paramDist_df_cumu_mean,aes(x=Year,y=MSE, colour = sample_freq, group = interaction(sample_freq, budget))) + 
  geom_line(linewidth = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  facet_wrap(~budget, ncol = 4) + coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")
```

# other option would be not to take the mean of the estimates but try and show mean + 95% credible interval
```{r}
# take mean & 95-percentiles of simulation replicates (I already have the mean)
#mse_SimFromSimFits_cumulative_paramDist_mean <- matrix(NA, nrow = 96 *3, ncol = 20) # rows = 96 runs, cols = 20 years
mse_SimFromSimFits_cumulative_paramDist_low <- matrix(NA, nrow = 96, ncol = 20) # rows = 96 runs, cols = 20 years
mse_SimFromSimFits_cumulative_paramDist_up <- matrix(NA, nrow = 96, ncol = 20) # rows = 96 runs, cols = 20 years
for (i in 1:96) {
    #mse_SimFromSimFits_cumulative_paramDist_mean95[(i-1)*3 + 1,] <- colMeans(mse_SimFromSimFits_cumulative_paramDist[(i-1)*200 + (1:200),])
    mse_SimFromSimFits_cumulative_paramDist_low[i,] <- apply(mse_SimFromSimFits_cumulative_paramDist[(i-1)*200 + (1:200),],2,quantile,.025)
    mse_SimFromSimFits_cumulative_paramDist_up[i,] <- apply(mse_SimFromSimFits_cumulative_paramDist[(i-1)*200 + (1:200),],2,quantile,.975)
}

mse_SimFromSimFits_paramDistMean95_df_cumu <- data.frame(cbind(rep(sampling_strat_order$index,20), rep(sampling_strat_order$budget,20), rep(sampling_strat_order$sample_freq,20),as.vector(mse_SimFromSimFits_cumulative_paramDist_mean),as.vector(mse_SimFromSimFits_cumulative_paramDist_low),as.vector(mse_SimFromSimFits_cumulative_paramDist_up)), sort(rep(1:20, 96)), as.factor(sort(rep(1:3, 32))))
colnames(mse_SimFromSimFits_paramDistMean95_df_cumu) <- c("index", "budget","sample_freq", "MSEmean", "MSElow","MSEup", "Year", "Replicate")
mse_SimFromSimFits_paramDistMean95_df_cumu$budget <- as.factor(mse_SimFromSimFits_paramDistMean95_df_cumu$budget)
mse_SimFromSimFits_paramDistMean95_df_cumu$sample_freq <- as.factor(mse_SimFromSimFits_paramDistMean95_df_cumu$sample_freq)

ggplot(mse_SimFromSimFits_paramDistMean95_df_cumu,aes(x=Year,y=MSEmean, colour = budget, group = interaction(sample_freq, budget, Replicate))) + 
  geom_line(linewidth = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  geom_ribbon(aes(ymin=MSElow, ymax=MSEup), alpha=0.2, fill = budget) +
  facet_wrap(~sample_freq, ncol = 4) + coord_cartesian(ylim = c(0,0.02)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Mean squared error", x = "Sampling budget") 

# take mean over fitting replicates
mse_SimFromSimFits_paramDistMean95_df_cumu_mean <- aggregate(.~budget+sample_freq+Year,mse_SimFromSimFits_paramDistMean95_df_cumu[,-8], mean)

ggplot(mse_SimFromSimFits_paramDistMean95_df_cumu_mean,aes(x=Year,y=MSEmean, colour = budget, group = budget, fill = budget)) + 
  geom_ribbon(aes(ymin=MSElow, ymax=MSEup),alpha=0.3, linetype=0) +
  geom_line(linewidth = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  facet_wrap(~sample_freq, ncol = 4) + coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")

ggplot(mse_SimFromSimFits_paramDistMean95_df_cumu_mean,aes(x=Year,y=MSEmean, colour = budget, group = budget, fill = budget)) + 
  geom_errorbar(aes(ymin=MSElow, ymax=MSEup),alpha=0.5) +
  geom_point(size = 1) + 
  geom_line(linewidth = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  facet_wrap(~sample_freq, ncol = 4) + coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")


ggplot(mse_SimFromSimFits_paramDistMean95_df_cumu_mean,aes(x=Year,y=MSEmean, colour = sample_freq, group = interaction(sample_freq, budget), fill = sample_freq)) + 
  geom_ribbon(aes(ymin=MSElow, ymax=MSEup),alpha=0.3, linetype=0) +
  geom_line(linewidth = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  facet_wrap(~budget, ncol = 4) + coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")

ggplot(mse_SimFromSimFits_paramDistMean95_df_cumu_mean,aes(x=Year,y=MSEmean, colour = sample_freq, group = interaction(sample_freq, budget), fill = sample_freq)) + 
  geom_errorbar(aes(ymin=MSElow, ymax=MSEup),alpha=0.5) +
  geom_point(size = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  geom_line(linewidth = 1) + 
  facet_wrap(~budget, ncol = 4) + coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")

ggplot(mse_SimFromSimFits_paramDistMean95_df_cumu_mean,aes(x=Year,y=MSEmean, colour = sample_freq, group = interaction(sample_freq, budget), fill = sample_freq)) + 
  geom_errorbar(aes(ymin=MSElow, ymax=MSEup),alpha=0.5) +
  geom_point(size = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  geom_line(linewidth = 1) + 
  facet_wrap(~budget, ncol = 2) + coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")

ggplot(mse_SimFromSimFits_paramDistMean95_df_cumu_mean,aes(x=Year,y=MSEmean, colour = sample_freq, group = interaction(sample_freq, budget), fill = sample_freq)) + 
  geom_errorbar(aes(ymin=MSElow, ymax=MSEup),alpha=0.5) +
  geom_point(size = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  geom_line(linewidth = 1) + 
  facet_wrap(~budget, ncol = 2) + coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines"),legend.position="bottom") +
  labs(y= "Cumulative mean squared error", x = "Years")

# or make 32 plots? :D
ggplot(mse_SimFromSimFits_paramDistMean95_df_cumu_mean,aes(x=Year,y=MSEmean, colour = sample_freq, fill = sample_freq)) + 
  geom_errorbar(aes(ymin=MSElow, ymax=MSEup),alpha=0.5) +
  geom_point(size = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  geom_line(linewidth = 1) + 
  facet_grid(budget~sample_freq, axes = "all", axis.labels = "all_x") + coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")

ggplot(mse_SimFromSimFits_paramDistMean95_df_cumu_mean,aes(x=Year,y=MSEmean, colour = sample_freq, fill = sample_freq)) + 
  geom_errorbar(aes(ymin=MSElow, ymax=MSEup),alpha=0.5) +
  geom_point(size = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  geom_line(linewidth = 1) + 
  facet_grid(sample_freq~budget, axes = "all", axis.labels = "all_x") + coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")
```

```{r}
# horizontal version of the plot (4 columns, 2 rows)
ggplot(mse_SimFromSimFits_paramDistMean95_df_cumu_mean,aes(x=Year,y=MSEmean, colour = sample_freq, group = interaction(sample_freq, budget), fill = sample_freq)) + 
  geom_errorbar(aes(ymin=MSElow, ymax=MSEup),alpha=0.5) +
  geom_point(size = 1) + scale_color_viridis_d(option = "turbo",direction = -1) + 
  geom_line(linewidth = 1) + 
  facet_wrap(~budget, ncol = 4) + coord_cartesian(ylim = c(0,0.0075)) +
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20),panel.spacing = unit(2, "lines")) +
  labs(y= "Cumulative mean squared error", x = "Years")
```

