---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
path_to_data <- "~/Documents/PhD_Project/Data/"
library(readxl)
```

#### read in meta data
```{r}
joined_metadata_list <- vector(mode = "list", length = 3) # meta data list for UK, US, Nepal
names(joined_metadata_list) <- c("UK", "US", "Nepal")

joined_PopPUNK_list <- vector(mode = "list", length = 3) # PopPUNK list for UK, US, Nepal
names(joined_metadata_list) <- c("UK", "US", "Nepal")
```

### Nepal
```{r}
# GPS1
Nepal_meta_data_gps1 <- as.data.frame(readxl::read_excel("~/Documents/PhD_Project/Data/StrepPneumo_Nepal_gps_v2/GPS1_and_GPS2_Nepal_dataset_3Dec2024.xlsx", sheet = "GPS1"))
Nepal_meta_data_gps1_selection <- Nepal_meta_data_gps1[,c(2:6,8:13, 18, 25, 75)]
Nepal_meta_data_gps1_selection$GPS <- 1

Nepal_meta_data_gps1_selection_v2 <- Nepal_meta_data_gps1[,c(2:4,8,9, 12, 13, 18, 25, 75)]
Nepal_meta_data_gps1_selection_v2$GPS <- 1

#GPS2 serotypes
Nepal_gps2_serotyping <- read.csv("~/Documents/PhD_Project/Data/StrepPneumo_Nepal_gps_v2/SerotypeSummary.csv", header = FALSE)
colnames(Nepal_gps2_serotyping) <- c("Sample","Serotype","Genetic_Variant","Contamination_Status")

spl_serotype_names <- function(serotype_str){
  split1 <- strsplit(serotype_str, "/data/SerotypingResults/", fixed = TRUE)[[1]][2]
  split2 <- strsplit(split1, "_RESULT", fixed = TRUE)[[1]][1]
  split2
}
Nepal_gps2_serotyping$Sample <- as.vector(sapply(Nepal_gps2_serotyping$Sample,spl_serotype_names))

gps2_serotype_dict <- Nepal_gps2_serotyping$Serotype
names(gps2_serotype_dict) <- Nepal_gps2_serotyping$Sample

# GPS2
Nepal_meta_data_gps2 <- as.data.frame(readxl::read_excel("~/Documents/PhD_Project/Data/StrepPneumo_Nepal_gps_v2/GPS1_and_GPS2_Nepal_dataset_3Dec2024.xlsx", sheet = "GPS2")) 
# length(which(Nepal_meta_data_gps2$QC == "FAIL")) 90 --> 90 of these sequences did not pass quality control--> I will filter these out now (Oct 2025)
Nepal_meta_data_gps2_selection <- Nepal_meta_data_gps2[which(Nepal_meta_data_gps2$QC != "FAIL"),c(1,2,4:11, 16,21,22)]
Nepal_meta_data_gps2_selection$Clinical_manifestation[which(Nepal_meta_data_gps2_selection$Clinical_manifestation == "OTHER")] <- "CARRIAGE"
Nepal_meta_data_gps2_selection$GPS <- 2
colnames(Nepal_meta_data_gps2_selection)[2] <- "Public_name"
Nepal_meta_data_gps2_selection$"In_silico_serotype" <- rep(NA, nrow(Nepal_meta_data_gps2_selection))
Nepal_meta_data_gps2_selection$"In_silico_serotype" <- gps2_serotype_dict[Nepal_meta_data_gps2_selection$"Lane_id"]
#note: there are 66 entries in here that do not have genome data available 

comm_carriage_dict <- (Nepal_meta_data_gps2$Study_name == "COMMUNITY CARRIAGE")
names(comm_carriage_dict) <- Nepal_meta_data_gps2$Lane_id

# new version of GPSC metadata
library(readr)
Nepal_meta_data_gps2_v2 <- read_csv("~/Documents/PhD_Project/Data/StrepPneumo_Nepal_gps_v2/GPS_Nepal_data_19Sep2025.csv")
filter_for_carriage <- (Nepal_meta_data_gps2_v2$Manifestation == "CARRIAGE")
names(filter_for_carriage) <- Nepal_meta_data_gps2_v2$Lane_id
for (i in names(filter_for_carriage)) {
  if(!is.na(comm_carriage_dict[i]) & comm_carriage_dict[i]){
    filter_for_carriage[i] <- TRUE
  }
}
Nepal_meta_data_gps2_v2_selection <- Nepal_meta_data_gps2_v2[filter_for_carriage,c(2:4, 7:9, 11, 23)]
#Nepal_meta_data_gps2_v2_selection$Month <- rep(NA, nrow(Nepal_meta_data_gps2_v2_selection))
#Nepal_meta_data_gps2_v2_selection$Phenotypic_serotype <- rep(NA, nrow(Nepal_meta_data_gps2_v2_selection))
#Nepal_meta_data_gps2_v2_selection$Clinical_manifestation[which(Nepal_meta_data_gps2_selection$Clinical_manifestation == "OTHER")] <- "CARRIAGE"
#Nepal_meta_data_gps2_v2_selection$GPS <- 2
for (i in unique(Nepal_meta_data_gps2_v2_selection$Year)) {
  print(paste(i, length(which(Nepal_meta_data_gps2_v2_selection$Year == i)), sep = " "))
}
# total 1881 nrow(Nepal_meta_data_gps2_v2_selection) 
# will use this going forward (Oct 2025)

# use this updated metadata sheet for serotyping
#gps2_serotype_dict_v2 <- Nepal_meta_data_gps2_v2$In_silico_serotype
#names(gps2_serotype_dict_v2) <- Nepal_meta_data_gps2_v2$Lane_id

#Nepal_meta_data_gps2_v2_selection <- Nepal_meta_data_gps2_selection
#Nepal_meta_data_gps2_v2_selection$"In_silico_serotype" <- rep(NA, nrow(Nepal_meta_data_gps2_v2_selection))
#Nepal_meta_data_gps2_v2_selection$"In_silico_serotype" <- gps2_serotype_dict_v2[Nepal_meta_data_gps2_v2_selection$"Lane_id"]

# check how much these two serotyping methods really differ
same <- 0
diff <- 0
diff_names <- c()
#diff_sero1 <- rep(NA, 449)
not_in_set <- 0
not_in_set_names <- c()
for (sero_i in names(gps2_serotype_dict)) {
  if(!is.na(gps2_serotype_dict_v2[sero_i]) & gps2_serotype_dict[sero_i] == gps2_serotype_dict_v2[sero_i]){
    same <- same + 1
  } else if(!is.na(gps2_serotype_dict_v2[sero_i])){
    diff <- diff + 1
    diff_names <- c(diff_names, sero_i)
    print(paste(gps2_serotype_dict[sero_i], " and ", gps2_serotype_dict_v2[sero_i], sep = ""))
  } else{
    not_in_set <- not_in_set + 1
    not_in_set_names <- c(not_in_set_names, sero_i)
  }
}
#actually, not really any differences. Just slightly different names, e.g. 06A and 6A.
# not_in_set_names seem to be the ones that did not pass quality control and do not appear in new Nepal dataset

Nepal_meta_data_gps_joined_old <- rbind(Nepal_meta_data_gps1_selection, Nepal_meta_data_gps2_selection)
#Nepal_meta_data_gps_joined_v2 <- rbind(Nepal_meta_data_gps1_selection, Nepal_meta_data_gps2_v2_selection)

# 2610 samples
# now filter by those that are carriage/"other"

Nepal_meta_data_gps_joined_old <- Nepal_meta_data_gps_joined_old[which(Nepal_meta_data_gps_joined_old$Clinical_manifestation == "CARRIAGE"),]
#Nepal_meta_data_gps_joined_v2 <- Nepal_meta_data_gps_joined_v2[which(Nepal_meta_data_gps_joined_v2$Clinical_manifestation == "CARRIAGE"),]
# 2001 samples

# add phenotypic serotype to dataset
pheno_sero_dict <- Nepal_meta_data_gps_joined_old$Phenotypic_serotype
names(pheno_sero_dict) <- Nepal_meta_data_gps_joined_old$Lane_id
Nepal_meta_data_gps2_v2_selection$Phenotypic_serotype <- pheno_sero_dict[Nepal_meta_data_gps2_v2_selection$Lane_id]

Nepal_meta_data_gps_joined <- Nepal_meta_data_gps2_v2_selection

Nepal_time_points <- sort(unique(Nepal_meta_data_gps_joined$Year))

Nepal_SeqYear_dict <- Nepal_meta_data_gps_joined$Year
names(Nepal_SeqYear_dict) <- Nepal_meta_data_gps_joined$Lane_id

Nepal_meta_IDtoInd <- 1:nrow(Nepal_meta_data_gps_joined)
names(Nepal_meta_IDtoInd) <- Nepal_meta_data_gps_joined$Lane_id
```

```{r}
#make ggplot version of this plot
# Nepal
library(RColorBrewer)
myCol <- brewer.pal(3, "Pastel2")
myCol <- c("#D55E00", "#56B4E9", "black")

Nepal_numbers_year_df <- cbind(names(Nepal_numbers_year)[-c(1,2,3,4,16,17)], as.integer(Nepal_numbers_year[-c(1,2,3,4,16,17)]))
colnames(Nepal_numbers_year_df) <- c("Years","Isolates")
ggplot(Nepal_numbers_year_df, aes(x=Years, y=as.integer(Isolates))) + 
  geom_bar(stat = "identity", color="black", fill=myCol[3]) + theme_bw()

# UK
UK_PP <- readRDS("UK_PP.rds")
UK_metadata_proc <- readRDS("UK_metadata_processed.rds")
UKPP_time_points <- unique(sort(UK_metadata_proc$Winter))
# [1] "2006" "2007" "2008" "2009" "2010" "2011" "2012"

UK_seq_no_per_year <- rep(0, length(UKPP_time_points))
names(UK_seq_no_per_year) <- UKPP_time_points
for (i in 1:length(UK_time_points)) {
  UK_seq_no_per_year[i] <- length(which(UK_metadata_proc$Winter==UKPP_time_points[i]))
}
UK_numbers_year_df <- cbind(names(UK_seq_no_per_year), UK_seq_no_per_year)
colnames(UK_numbers_year_df) <- c("Winters","Isolates")
ggplot(UK_numbers_year_df, aes(x=Winters, y=as.integer(Isolates))) + 
  geom_bar(stat = "identity", color="black", fill=myCol[2]) + theme_bw() + ylim(c(0,125))

# US
Mass_Samples_accCodes_processed <- readRDS("Mass_Samples_accCodes_processed.rds")
Mass_time_points <- c(2001, 2004, 2007)
Mass_seq_no_per_year <- rep(0, length(2001:2007))
names(Mass_seq_no_per_year) <- 2001:2007
for (i in 1:length(Mass_time_points)) {
   Mass_seq_no_per_year[as.character(Mass_time_points[i])] <- length(which(Mass_Samples_accCodes_processed$`Year of Isolation`==Mass_time_points[i]))
}
Mass_numbers_year_df <- cbind(names(Mass_seq_no_per_year), Mass_seq_no_per_year)
colnames(Mass_numbers_year_df) <- c("Years","Isolates")
ggplot(Mass_numbers_year_df, aes(x=Years, y=as.integer(Isolates))) + 
  geom_bar(stat = "identity", color="black", fill=myCol[1]) + theme_bw() + ylim(c(0,300))
```

# create serotype dictionary
```{r}
# unique(Nepal_meta_data_gps_joined$In_silico_serotype)
# collate "NCC1_PSPK_NON_ENCAPSULATED", "ALTERNATIVE_ALIB_NT", "SWISS_NT", "UNTYPABLE"  to "NT"
# try to substitute "COVERAGE TOO LOW" with phenotypic serotype

Nepal_Serotype_dict <- rep(NA, nrow(Nepal_meta_data_gps_joined))
names(Nepal_Serotype_dict) <- Nepal_meta_data_gps_joined$Lane_id

Nepal_Serotype_dict[Nepal_meta_data_gps_joined$Lane_id[which(Nepal_meta_data_gps_joined$In_silico_serotype != "COVERAGE TOO LOW")]] <- Nepal_meta_data_gps_joined$In_silico_serotype[which(Nepal_meta_data_gps_joined$In_silico_serotype != "COVERAGE TOO LOW")]

Nepal_Serotype_dict[Nepal_meta_data_gps_joined$Lane_id[which(Nepal_meta_data_gps_joined$In_silico_serotype == "COVERAGE TOO LOW")]] <- Nepal_meta_data_gps_joined$Phenotypic_serotype[which(Nepal_meta_data_gps_joined$In_silico_serotype == "COVERAGE TOO LOW")]

Nepal_Serotype_dict[which(Nepal_Serotype_dict == "UNTYPABLE" | Nepal_Serotype_dict == "SWISS_NT" | Nepal_Serotype_dict == "ALTERNATIVE_ALIB_NT" | Nepal_Serotype_dict == "untypable" | Nepal_Serotype_dict == "NCC1_PSPK_NON_ENCAPSULATED" | Nepal_Serotype_dict == "NCC3_aliD_non_encapsulated" | Nepal_Serotype_dict == "NCC2_S_mitis_aliC_aliD_non_encapsulated" | Nepal_Serotype_dict == "NCC2_aliC_aliD_non_encapsulated")] <- "NT"

# unique(Nepal_Serotype_dict)
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "6E(6B)")] <- "6B"
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "6A(6A-III)")] <- "6A"
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "15B")] <- "15B/15C"
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "15C")] <- "15B/15C"
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "19A(19A-I/19A-II)")] <- "19A"
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "23B1")] <- "23B"
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "24B/24C/24F")] <- "SEROGROUP 24"
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "24F")] <- "SEROGROUP 24"

# length(unique(Nepal_Serotype_dict)) 68

Nepal_has_serotype <- rep(TRUE, nrow(Nepal_meta_data_gps_joined))
names(Nepal_has_serotype) <- Nepal_meta_data_gps_joined$Lane_id
```

# read in PopPUNK results
```{r}
PP_all <- read.csv(paste(path_to_data, "StrepPneumo_Nepal_gps_v2/poppunk_clusters_GPSCs_v9/poppunk_clusters_external_clusters.csv",sep = ""))
# since Nepal was in the database, I had to run PopPUNK with --write-references
# so now I need to filter this dataset to only have the real Nepal sequences

# remove appendix "_contigs_velvet"
#remove_contigs_velvet <- function(str_with_app){
#  str_without_app <- strsplit(str_with_app, "_contigs_velvet", fixed = TRUE)[[1]][1]
#  str_without_app
#}

#PP_all$Taxon <- sapply(PP_all$Taxon, remove_contigs_velvet)
PP_all <- PP_all[!duplicated(PP_all$sample),]

PP_taxon_ind_dict <- 1:nrow(PP_all)
names(PP_taxon_ind_dict) <- PP_all$sample

PP_is_Nepal_dict <- rep(FALSE,nrow(PP_all))
names(PP_is_Nepal_dict) <- PP_all$sample

for (i in 1:length(Nepal_meta_data_gps_joined$Lane_id)) {
  if(Nepal_has_serotype[i]){ # checks if we have serotype for this sequence and only includes it in PP dataset if yes
    PP_is_Nepal_dict[Nepal_meta_data_gps_joined$Lane_id[i]] <- TRUE
  }
}

Nepal_PP <- PP_all[which(PP_is_Nepal_dict[PP_all$sample]),]
Nepal_PP <- Nepal_PP[which(!(is.na(Nepal_PP$GPSC))),]
no_Nepal_PP <- length(unique(Nepal_PP$GPSC)) 
# now: 170 GPSCs

Nepal_seq_clusters_dict <- Nepal_PP$GPSC
names(Nepal_seq_clusters_dict) <- Nepal_PP$sample
```

# filter meta data by PopPUNK sequences (PopPUNK qc removed some of the sequences)
```{r}
#Nepal_meta_dataPP <- Nepal_meta_data_gps_joined[sort(Nepal_meta_IDtoInd[Nepal_PP$sample]),c("Lane_id", "Year")]
Nepal_meta_dataPP <- Nepal_meta_data_gps_joined[sort(Nepal_meta_IDtoInd[Nepal_PP$sample]),c("Lane_id", "Year")]
Nepal_meta_dataPP$Serotype <- Nepal_Serotype_dict[Nepal_meta_dataPP$Lane_id]
colnames(Nepal_meta_dataPP) <- c("Lane_id", "Year", "Serotype")

#rownames(Nepal_meta_dataPP) <- 1:nrow(Nepal_meta_dataPP)

Nepal_meta_IDtoIndPP <- 1:nrow(Nepal_meta_dataPP)
names(Nepal_meta_IDtoIndPP) <- Nepal_meta_dataPP$Lane_id
```

```{r}
# save Nepal metadata in metadata list
joined_metadata_list[["Nepal"]] <- Nepal_meta_dataPP

saveRDS(Nepal_meta_dataPP, "Nepal_meta_dataPP.rds")
```

```{r}
# Add Serotype Information to PP
Nepal_PP$Serotype <- Nepal_Serotype_dict[Nepal_PP$sample]

# create dictionary for VTs of the PCV10
# 1, 4, 5, 6b, 7f, 9V, 14, 18c, 19f and 23f serotypes (source: NCBI)
PCV10_VTsPP <- rep("NVT",length(unique(Nepal_PP$Serotype)))
names(PCV10_VTsPP) <- unique(Nepal_PP$Serotype)
PCV10_VTsPP[c("1","4","5","6A","6B", "7F", "9V", "14", "18C", "19F", "23F")] <- "VT" # also added 6A

Nepal_PP$VT <- PCV10_VTsPP[Nepal_PP$Serotype]


joined_PopPUNK_list[["Nepal"]] <- Nepal_PP

saveRDS(Nepal_PP, "Nepal_PP.rds")
```

### UK
```{r}
### Reading in the Accession Codes, Population and the Sequence Clusters
UK_metadata <- read.csv("~/Documents/PhD_Project/Data/StrepPneumo_UK/metadata.csv")

UK_ids <- UK_metadata$id
UK_ids_dict <- 1:length(UK_ids)
names(UK_ids_dict) <- UK_ids

UK_id_acc_dict <- UK_ids
names(UK_id_acc_dict) <- UK_metadata$Accession

#Isolates <- Samples_accCodes$`Isolate Name`
#Isolates_dict <- 1:length(Isolates)
#names(Isolates_dict) <- Isolates

#Isolate_from_Mass_dict <- rep(0, length(Isolates))
#names(Isolate_from_Mass_dict) <- Isolates
#Isolate_from_Mass_dict[Mass_Isolates] <- 1

# make plot to show year-and-month
UK_year_and_month <- rep(0, )

UK_years <- 1:8
names(UK_years) <- unique(UK_metadata$Year)
UK_numbers_year_month <- data.frame(cbind(sort(rep(unique(UK_metadata$Year),12)), rep(1:12,length(unique(UK_metadata$Year))), rep(0,length(unique(UK_metadata$Year)*12))))
colnames(UK_numbers_year_month) <- c("Year","Month","Count")
for (i in 1:length(UK_metadata$id)) {
  UK_numbers_year_month[12 * (UK_years[as.character(UK_metadata$Year[i])]-1) + UK_metadata$Month[i] ,3] <- UK_numbers_year_month[12 * (UK_years[as.character(UK_metadata$Year[i])]-1) + UK_metadata$Month[i] ,3] + 1
}

par(mfrow=c(2,1))
plot(UK_numbers_year_month[,1] + 1/12 * (UK_numbers_year_month[,2]-1), UK_numbers_year_month[,3])
abline(v=unique(UK_metadata$Year)-1/24, col="blue")
plot(UK_numbers_year_month[,1] + 1/12 * (UK_numbers_year_month[,2]-1), UK_numbers_year_month[,3])
abline(v=c(unique(UK_metadata$Year),2014)-13/24, col="blue")

# yes, winter assignment seems to make sense for UK

split_winter <- function(winter){
  strsplit(winter,"\\/")[[1]][1]
}

UK_SeqYear_dict <- sapply(UK_metadata$Winter, split_winter)
#names(UK_SeqYear_dict) <- UK_ids
names(UK_SeqYear_dict) <- UK_metadata$Accession

plot(UK_SeqYear_dict, UK_metadata$Year) # there seem to be some mis-assignments - I'll assume that the date is correct and the winter is incorrectly assigned?
# so I will re-do the SeqYear dict
UK_SeqYear_dict <- c()
UK_winters <- c()
for (i in 1:length(UK_metadata$Year)){
   if(UK_metadata$Month[i] <= 6){
     UK_SeqYear_dict[i] <- UK_metadata$Year[i] - 1
     UK_winters[i] <- paste(as.character(UK_SeqYear_dict[i]), as.character(UK_metadata$Year[i]), sep= "/")
  }
  else{
    UK_SeqYear_dict[i] <- UK_metadata$Year[i]
    UK_winters[i] <- paste(as.character(UK_SeqYear_dict[i]), as.character(UK_metadata$Year[i]+1), sep="/")
  } 
  
}
names(UK_SeqYear_dict) <- UK_metadata$Accession
plot(UK_SeqYear_dict, UK_metadata$Year) # looks better now

UK_meta_IDtoInd <- 1:nrow(UK_metadata)
names(UK_meta_IDtoInd) <- UK_metadata$Accession

UK_time_points <- sort(unique(UK_SeqYear_dict))

#UK_winters <- sort(unique(UK_metadata$Winter)) # now also defined based on month-year
UK_metadata$Winter <- UK_winters # correct winters in df as well
```

```{r}
# save UK metadata in metadata list
joined_metadata_list[["UK"]] <- UK_metadata

saveRDS(UK_metadata, "UK_metadata_processed.rds")
```

# PopPUNK UK
```{r}
#PP_all <- read.csv(paste(path_to_data, "StrepPneumoUKjoined/PopPUNK/poppunk_clusters_clusters.csv",sep = ""))
PP_all <- read.csv(paste(path_to_data, "StrepPneumoUKjoined/poppunk_clusters_GPSCs_v9/poppunk_clusters_external_clusters.csv",sep = ""))
colnames(PP_all) <- c("Taxon","Cluster")
no_PP_all <- length(unique(PP_all$Cluster)) 
```

```{r}
PP_taxon_ind_dict <- 1:nrow(PP_all)
names(PP_taxon_ind_dict) <- PP_all$Taxon

PP_is_UK_dict <- rep(FALSE,nrow(PP_all))
names(PP_is_UK_dict) <- PP_all$Taxon

for (i in 1:length(UK_metadata$Accession)) {
  PP_is_UK_dict[UK_metadata$Accession[i]] <- TRUE
}
# seq ERR657323 is only in metadata, not in PP info. Need to remove that one
UK_PP <- PP_all[PP_is_UK_dict,]
UK_PP <- UK_PP[-nrow(UK_PP),]
no_UK_PP <- length(unique(UK_PP$Cluster)) 
# 54 GPSCs
# 59 clusters (old)

UK_seq_clusters_dict <- UK_PP$Cluster
names(UK_seq_clusters_dict) <- UK_PP$Taxon
```

```{r}
# Add Serotype Information
UK_Serotype_dict <- UK_metadata$Serotype
names(UK_Serotype_dict) <- UK_metadata$Accession

UK_PP$Serotype <- UK_Serotype_dict[UK_PP$Taxon]

UK_PP$Time <- UK_SeqYear_dict[UK_PP$Taxon]

# create dictionary for VTs of the PCV10
# 1, 4, 5, 6b, 7f, 9V, 14, 18c, 19f and 23f serotypes (source: NCBI)
#PCV7: 4, 6B, 9V, 14, 18C, 19F and 23F. (source: https://journals.plos.org/plospathogens/article?id=10.1371/journal.ppat.1006966#sec018)
# PCV13: (1, 3, 4, 5, 6A and 6B, 7F, 9V, 14, 19A, and 19F, 18C, and 23F) (source: https://www.ncbi.nlm.nih.gov/books/NBK507794/)
PCV13_VTs <- rep("NVT",length(unique(UK_metadata$Serotype)))
names(PCV13_VTs) <- unique(UK_metadata$Serotype)
PCV13_VTs[c("1", "3", "4","5","6A","6B", "7F", "9V", "14", "18C", "19A", "19F", "23F")] <- "VT"

# add 6A to PCV7 because there is strong cross-immunity btw PVC7 and 6A (4.Croucher, N. J. et al. Population genomics of post-vaccine changes in pneumococcal epidemiology. Nat. Genet. 45, 656–663 (2013).)
PCV7_VTs <- rep("NVT",length(unique(UK_metadata$Serotype)))
names(PCV7_VTs) <- unique(UK_metadata$Serotype)
PCV7_VTs[c("4","6A", "6B", "9V", "14", "18C", "19F", "23F")] <- "VT"

UK_PP$VT_PCV7 <- PCV7_VTs[UK_PP$Serotype]
UK_PP$VT_PCV13 <- PCV13_VTs[UK_PP$Serotype]

saveRDS(UK_PP,"UK_PP.rds")

joined_PopPUNK_list[["UK"]] <- UK_PP
```


### US
```{r}
### Reading in the Accession Codes, Population and the Sequence Clusters
Samples_accCodes <- read_excel(paste(path_to_data, "Massachusetts_data_NickCroucher/SupplementaryDataPaper/Samples_accCodes.xlsx", sep = ""))
Mass_Samples_accCodes <- Samples_accCodes[Samples_accCodes$Population=="Massachusetts",]
Mass_Isolates <- Mass_Samples_accCodes$`Isolate Name`
Mass_Isolates_dict <- 1:length(Mass_Isolates)
names(Mass_Isolates_dict) <- Mass_Isolates

Isolates <- Samples_accCodes$`Isolate Name`
Isolates_dict <- 1:length(Isolates)
names(Isolates_dict) <- Isolates

Isolate_from_Mass_dict <- rep(0, length(Isolates))
names(Isolate_from_Mass_dict) <- Isolates
Isolate_from_Mass_dict[Mass_Isolates] <- 1
```

# Read in PopPUNK clusters:

```{r}
#PopPUNK_all <- read.csv("/Users/llorenz/Documents/PhD_Project/Data/Massachusetts_SeqClusters/PopPUNKwithStandardDB/poppunk_clusters_clusters.csv")
PopPUNK_all <- read.csv("/Users/llorenz/Documents/PhD_Project/Data/Massachusetts_SeqClusters/poppunk_clusters_GPSCs_v9/poppunk_clusters_external_clusters.csv")
colnames(PopPUNK_all) <- c("Taxon","Cluster")
no_PopPUNK_all <- length(unique(PopPUNK_all$Cluster)) # number of clusters in dataset
```

```{r}
# Add Isolate Name to PopPUNK data frame
accNo_to_filename <- read.delim("~/Documents/PhD_Project/Code/1st_project/odin-dust-examples/Data/filereport_read_run_PRJEB2632_tsv.txt")
accNo_to_filename <- accNo_to_filename[,c(1,8)]
accNo_to_filename[,3] <- matrix(unlist(strsplit(accNo_to_filename[,2],"/")), ncol=6, byrow = TRUE)[,6]
accNo_to_filename[,3] <- matrix(unlist(strsplit(accNo_to_filename[,3],"[.]")), ncol=2, byrow = TRUE)[,1]
accNo_to_filename <- accNo_to_filename[,c(1,3)]
colnames(accNo_to_filename) <- c(colnames(accNo_to_filename)[1], "filenames")

filenameToAccNo_dict <- accNo_to_filename$run_accession
names(filenameToAccNo_dict) <- accNo_to_filename$filenames

AccNoToIsolate_dict <- Mass_Samples_accCodes$`Isolate Name`
names(AccNoToIsolate_dict) <- Mass_Samples_accCodes$`Accession Code`

#PopPUNK_clusters$IsolateName <- AccNoToIsolate_dict[filenameToAccNo_dict[PopPUNK_clusters$Taxon]]
```

```{r}
PopPUNK_clusters <- data.frame(matrix(NA, nrow = nrow(Mass_Samples_accCodes), ncol = 3))
colnames(PopPUNK_clusters) <- c(colnames(PopPUNK_all), "IsolateName")

counter <- 0
for (i in 1:nrow(PopPUNK_all)) {
  if(is.element(PopPUNK_all$Taxon[i], names(filenameToAccNo_dict))){
    if(is.element(filenameToAccNo_dict[PopPUNK_all$Taxon[i]], names(AccNoToIsolate_dict))){
    counter <- counter + 1
    PopPUNK_clusters[counter,] <- c(PopPUNK_all[i,],AccNoToIsolate_dict[filenameToAccNo_dict[PopPUNK_all[i,1]]])
    }
  }
}

no_PopPUNK_clusters <- length(unique(PopPUNK_clusters$Cluster)) 

# 52 GPSCs
```

```{r}
# Add VT information
VT_dict <- Mass_Samples_accCodes$`Vaccine Type`
names(VT_dict) <- Mass_Samples_accCodes$`Isolate Name`

PopPUNK_clusters$VT <- VT_dict[PopPUNK_clusters$IsolateName]
```

```{r}
# Add sequencing year information
SeqYear_dict <- Mass_Samples_accCodes$`Year of Isolation`
names(SeqYear_dict) <- Mass_Samples_accCodes$`Isolate Name`

PopPUNK_clusters$SeqYear <- SeqYear_dict[PopPUNK_clusters$IsolateName]

SeqYear_dict_Taxon <- PopPUNK_clusters$SeqYear
names(SeqYear_dict_Taxon) <- PopPUNK_clusters$Taxon
```

```{r}
# Add Serotype Information
Serotype_dict <- Mass_Samples_accCodes$Serotype
names(Serotype_dict) <- Mass_Samples_accCodes$`Isolate Name`

PopPUNK_clusters$Serotype <- Serotype_dict[PopPUNK_clusters$IsolateName]
```

```{r}
# Create PP clusters, split by serotype
#calculate the frequency of the gene clusters and year and serotype
PP_mass_cluster_freq_1_sero <- matrix(0,nrow = no_PopPUNK_clusters, ncol = length(unique(PopPUNK_clusters$Serotype)))
PP_mass_cluster_freq_2_sero <- matrix(0,nrow = no_PopPUNK_clusters, ncol = length(unique(PopPUNK_clusters$Serotype)))
PP_mass_cluster_freq_3_sero <- matrix(0,nrow = no_PopPUNK_clusters, ncol = length(unique(PopPUNK_clusters$Serotype)))
PP_mass_cluster_freq_sero_names <- unique(PopPUNK_clusters$Serotype)
for (i in 1:no_PopPUNK_clusters){
  for (k in unique(PopPUNK_clusters$Serotype)){
  PP_mass_cluster_freq_1_sero[i,which(unique(PopPUNK_clusters$Serotype)==k)] <- length(which(PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),][PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),]$Serotype ==k, ]$"SeqYear"==2001))
  PP_mass_cluster_freq_2_sero[i,which(unique(PopPUNK_clusters$Serotype)==k)] <- length(which(PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),][PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),]$Serotype ==k, ]$"SeqYear"==2004))
  PP_mass_cluster_freq_3_sero[i,which(unique(PopPUNK_clusters$Serotype)==k)] <- length(which(PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),][PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),]$Serotype ==k, ]$"SeqYear"==2007))
  }
}

#save sequence cluster frequencies
saveRDS(PP_mass_cluster_freq_1_sero, file = "PP_mass_cluster_freq_1_seroMod.rds")
saveRDS(PP_mass_cluster_freq_2_sero, file = "PP_mass_cluster_freq_2_seroMod.rds")
saveRDS(PP_mass_cluster_freq_3_sero, file = "PP_mass_cluster_freq_3_seroMod.rds")
```

# create matrix with PP clusters and serotypes
```{r}
PPxSero_matrix <- matrix(0, nrow = length(unique(PopPUNK_clusters$Cluster)), ncol = length(unique(PopPUNK_clusters$Serotype)))
# sort(unique(PopPUNK_clusters$Serotype))
# [1] "10A"   "11A"   "14"    "15A"   "15B/C" "15F"   "16F"   "17F"   "18C"   "19A"   "19F"   "21"    "22F"   "23A"
#[15] "23B"   "23F"   "3"     "31"    "33F"   "34"    "35B"   "35F"   "37"    "38"    "6A"    "6B"    "6C"    "7C" 
#[29] "7F"    "9N"    "9V"    "NT"

SerotypeToIndex_dict <- 1:length(unique(PopPUNK_clusters$Serotype))
names(SerotypeToIndex_dict) <- (unique(PopPUNK_clusters$Serotype))

#IndexToSerotype_dict <- (unique(PopPUNK_clusters$Serotype))
#names(IndexToSerotype_dict) <- 1:length(unique(PopPUNK_clusters$Serotype))

PPclustToIndex_dict <- 1:length(unique(PopPUNK_clusters$Cluster))
names(PPclustToIndex_dict) <- (unique(PopPUNK_clusters$Cluster))

for (i in 1:nrow(PopPUNK_clusters)) {
  PPxSero_matrix[PPclustToIndex_dict[as.character(PopPUNK_clusters$Cluster[i])],SerotypeToIndex_dict[PopPUNK_clusters$Serotype[i]]] <- 1
}
# matrix shows information whether PP and serotype co-occur in the dataset (1=TRUE)
# this can be used as the migration matrix:
PPxSero_matrix_prob <- PPxSero_matrix / sum(PPxSero_matrix)
PPxSero_matrix_prob <- matrix(sapply(PPxSero_matrix_prob,as.double), nrow = nrow(PPxSero_matrix), ncol = ncol(PPxSero_matrix))
saveRDS(PPxSero_matrix_prob, "PPsero_mig.rds")
saveRDS(PPxSero_matrix_prob, "PPsero_mig2.rds")

# same with new sorting 
PPsero_startpop6 <- matrix(0, nrow = length(unique(PopPUNK_clusters$Cluster)), ncol = length(unique(PopPUNK_clusters$Serotype)))
PP_expand_factor <- 15000 / sum(PP_mass_cluster_freq_1_sero)
exp_noise <- 1000
for (i in 1:nrow(PPsero_startpop6)) {
    PPsero_startpop6[i,] <- (sapply((PP_mass_cluster_freq_1_sero[i,] + rexp(n = length(PP_mass_cluster_freq_1_sero[i,]), rate = exp_noise)) * PP_expand_factor, rpois, n=1)) + ceiling(PPxSero_matrix_prob[i,])
  #PPsero_startpop5[i,] <- PP_mass_cluster_freq_1_sero[i,] + ceiling(PPxSero_matrix_prob[i,])
}
saveRDS(PPsero_startpop6, "PPsero_startpop6.rds")

plot(colSums(PP_mass_cluster_freq_1_sero)/sum(colSums(PP_mass_cluster_freq_1_sero)))
points(colSums(PPsero_startpop6)/sum(colSums(PPsero_startpop6)), col = "red")

plot(rowSums(PP_mass_cluster_freq_1_sero)/sum((PP_mass_cluster_freq_1_sero)))
points(rowSums(PPsero_startpop6)/sum(rowSums(PPsero_startpop6)), col = "red")
#much better

which(PPsero_startpop6[(PPclustToIndex_dict[4]),]>0)
# 7 11 27 28 29 32
# serotypes: 15C, 23B, 4, 1, 38, 34
which(PPsero_startpop6[(PPclustToIndex_dict[7]),]>0)
# 8 10 24 27
# serotypes: 11A, 8, 35F, 4
which(PPsero_startpop6[(PPclustToIndex_dict[9]),]>0)
# 2 11 13 14 16 24 30
# serotypes: 19F, 23B, 23A, 6A, 6C, 35F, 16F

# serotype VT vector
SeroVT <- rep(0, length(unique(PopPUNK_clusters$Serotype)))
for (i in 1:nrow(PopPUNK_clusters)) {
  SeroVT[SerotypeToIndex_dict[PopPUNK_clusters$Serotype[i]]] <- as.integer(PopPUNK_clusters$VT[i]=="VT")
}
saveRDS(SeroVT, "SeroVT.rds")
saveRDS(SeroVT, "SeroVT2.rds")

```

```{r}
# save US metadata in metadata list
joined_metadata_list[["US"]] <- Mass_Samples_accCodes

saveRDS(Mass_Samples_accCodes, "Mass_Samples_accCodes_processed.rds")

# save US PopPUNK clusters
joined_PopPUNK_list[["US"]] <- PopPUNK_clusters

saveRDS(PopPUNK_clusters, "PopPUNK_clusters.rds")
```

### ggCaller (this is what was newly analysed here - running ggCaller and Panaroo in one go on UK, US and Nepal)

```{r}
joined_ggCaller <- read.csv("/Users/llorenz/Documents/PhD_Project/Data/StrepPneumo_UKUSNepal/gene_presence_absence.csv", header=FALSE)

# converting the gene presence absence matrix into a boolean df (0 = gene not present, 1 = gene present)
convert_to_bool <- function(x){
  if (x=="") 0 else 1
}

joined_ggCaller_bool <- joined_ggCaller[,c(-2,-3)]
joined_ggCaller_bool[-1,-1] <- apply(joined_ggCaller_bool[-1,-1],c(1,2), convert_to_bool)

joined_ggCaller_bool[1,-1] <- sapply(joined_ggCaller_bool[1,-1], strsplit, split = ".contigs_velvet") # remove contigs appendix
colnames(joined_ggCaller_bool) <- joined_ggCaller_bool[1,]
```

# try filtering mobile genetic elements
```{r}
bakta_annotation_panaroo <- read.delim("~/Documents/PhD_Project/Data/StrepPneumo_UKUSNepal/bakta_annotation/create_tsv_annotation/bakta_annotation_and_before.tsv")
library(stringr)
bakta_annotation_mobile <- rep(FALSE, nrow(bakta_annotation_panaroo))
names(bakta_annotation_mobile) <- bakta_annotation_panaroo$Header2
bakta_annotation_mobile <- (str_detect(bakta_annotation_panaroo$Header1, coll("transposon", ignore_case = TRUE)) | str_detect(bakta_annotation_panaroo$Header1, coll("transposase", ignore_case = TRUE)) | str_detect(bakta_annotation_panaroo$Header1, coll("insertion sequence", ignore_case = TRUE)) | str_detect(bakta_annotation_panaroo$Header1, coll("mobile", ignore_case = TRUE)))
# 692/24451 =  0.0283015 match at least one of these 
names(bakta_annotation_mobile) <- bakta_annotation_panaroo$Header2

# filter ggCaller matrix by mobile genetic elements
mobile_genes_joined_ggCaller <- rep(FALSE, length(joined_ggCaller_bool[-1,1]))
names(mobile_genes_joined_ggCaller) <- joined_ggCaller_bool[-1,1]
mobile_genes_joined_ggCaller[names(which(bakta_annotation_mobile == TRUE))] <- TRUE

joined_ggCaller_bool_non_mobile <- joined_ggCaller_bool[c(1, 1 + (which(mobile_genes_joined_ggCaller == FALSE))),]

joined_ggCaller_bool <- joined_ggCaller_bool_non_mobile

saveRDS(joined_ggCaller_bool, "joined_ggCaller_bool.rds")
```


# split ggCaller matrix into UK, US, Nepal
```{r}
# filter ggCaller results by Nepal 
Nepal_ggCallerPP_bool <- data.frame(matrix(0, nrow = nrow(joined_ggCaller_bool), ncol = length(Nepal_PP$sample)+1))
Nepal_ggCallerPP_bool[-1,-1] <- joined_ggCaller_bool[-1,c(Nepal_PP$sample)]
colnames(Nepal_ggCallerPP_bool) <- c("Gene",Nepal_PP$sample)
Nepal_ggCallerPP_bool[1,] <- c("Gene",Nepal_PP$sample)
Nepal_ggCallerPP_bool[,1] <- joined_ggCaller_bool[,1]

# filter ggCaller results by the ones I have metadata for from UK
UK_ggCallerPP_bool <- data.frame(matrix(0, nrow = nrow(joined_ggCaller_bool), ncol = length(UK_PP$Taxon)+1))
for (i in 1:length(UK_PP$Taxon)) {
  UK_ggCallerPP_bool[-1,(i+1)] <- joined_ggCaller_bool[-1,UK_PP$Taxon[i]]
}

UK_ggCallerPP_bool[-1,-1] <- joined_ggCaller_bool[-1,c(UK_PP$Taxon)]
colnames(UK_ggCallerPP_bool) <- c("Gene",UK_PP$Taxon)
UK_ggCallerPP_bool[1,] <- c("Gene",UK_PP$Taxon)
UK_ggCallerPP_bool[,1] <- joined_ggCaller_bool[,1]

# filter ggCaller results by US 
US_ggCallerPP_bool <- data.frame(matrix(0, nrow = nrow(joined_ggCaller_bool), ncol = length(PopPUNK_clusters$Taxon)+1))
US_ggCallerPP_bool[-1,-1] <- joined_ggCaller_bool[-1,c(PopPUNK_clusters$Taxon)]
colnames(US_ggCallerPP_bool) <- c("Gene",PopPUNK_clusters$Taxon)
US_ggCallerPP_bool[1,] <- c("Gene",PopPUNK_clusters$Taxon)
US_ggCallerPP_bool[,1] <- joined_ggCaller_bool[,1]

# save ggCaller bool matrices in list
joined_ggCaller_list <- vector(mode = "list", length = 3) # ggCaller list for UK, US, Nepal
names(joined_ggCaller_list) <- c("UK", "US", "Nepal")

joined_ggCaller_list[["Nepal"]] <- Nepal_ggCallerPP_bool
joined_ggCaller_list[["UK"]] <- UK_ggCallerPP_bool
joined_ggCaller_list[["US"]] <- US_ggCallerPP_bool

saveRDS(joined_ggCaller_list, "joined_ggCaller_list.rds")
```

# process Nepal ggCaller
# split ggCaller gene presence absence matrix into different time points
```{r}
### continue working here
Nepal_time_points_selected <- min(Nepal_time_points):max(Nepal_time_points)
Nepal_ggCaller_byYear <- vector(mode = "list", length = length(Nepal_time_points_selected))

for (i in 1:length(Nepal_time_points_selected)) {
  Nepal_ggCaller_year <- data.frame(matrix(0, nrow = nrow(Nepal_ggCallerPP_bool), ncol = length(which(Nepal_meta_dataPP$Year==Nepal_time_points_selected[i]))+1)) # initialize data frame
  Nepal_ggCaller_year[1,-1] <- Nepal_ggCallerPP_bool[1,c(FALSE,Nepal_SeqYear_dict[unlist(Nepal_ggCallerPP_bool[1,-1])]==Nepal_time_points_selected[i])] # fill in first row (seq names)
  Nepal_ggCaller_year[-1,1] <- Nepal_ggCallerPP_bool[-1,1] # fill in first column (gene cluster names)
  Nepal_ggCaller_year[-1,-1] <- Nepal_ggCallerPP_bool[-1,c(FALSE,Nepal_SeqYear_dict[unlist(Nepal_ggCallerPP_bool[1,-1])]==Nepal_time_points_selected[i])] # fill in presence absence information
  Nepal_ggCaller_byYear[[i]] <- Nepal_ggCaller_year
}
```

# find intermediate frequency genes
# I will combine all pre-vaccine data points for this
```{r}
sum_as_int <- function(x){
  sum(as.integer(x))
}

Nepal_gene_freq_preVac <- rep(0, nrow(Nepal_ggCaller_byYear[[1]])-1)
#Nepal_time_points_preVac <- c(5,8,10) # updated this now too (July 2025) to use all pre-vac samples
Nepal_time_points_preVac <- c(1,4,6) # Oct 2025, correction after just using carriage data
preVac_datapoints <- 0
for (i in Nepal_time_points_preVac) {
  Nepal_gene_freq_preVac <- Nepal_gene_freq_preVac + apply(Nepal_ggCaller_byYear[[i]][-1,-1],1, sum_as_int)
  preVac_datapoints <- preVac_datapoints + length(Nepal_ggCaller_byYear[[i]][1,-1])
}
Nepal_gene_freq_preVac <- Nepal_gene_freq_preVac / preVac_datapoints

Nepal_gene_filter <- as.integer(Nepal_gene_freq_preVac<=0.95 & Nepal_gene_freq_preVac>=0.05)
# now 2340 (after filtering for carriage only)
```

# create intermediate frequency dataframes
```{r}
Nepal_ggCaller_byYear_intermed <- vector(mode = "list", length = length(Nepal_time_points_selected))

for (i in 1:length(Nepal_time_points_selected)) {
  Nepal_ggCaller_year <- data.frame(matrix(0, nrow = sum(Nepal_gene_filter)+1,ncol = length(which(Nepal_meta_dataPP$Year==Nepal_time_points_selected[i]))+1))
  if(ncol(Nepal_ggCaller_byYear[[i]])>1){
    Nepal_ggCaller_year <-  Nepal_ggCaller_byYear[[i]][c(1,which(Nepal_gene_filter==1)+1),]
  }
  else{
    Nepal_ggCaller_year <-  data.frame(matrix(Nepal_ggCaller_byYear[[2]][c(1,which(Nepal_gene_filter==1)+1),1], nrow = sum(Nepal_gene_filter)+1,ncol = 1))
  }
  Nepal_ggCaller_byYear_intermed[[i]] <- Nepal_ggCaller_year 
}

Nepal_ggCaller_intermed <- data.frame(matrix(0, nrow = sum(Nepal_gene_filter)+1,ncol = ncol(Nepal_ggCallerPP_bool)))
Nepal_ggCaller_intermed <-  Nepal_ggCallerPP_bool[c(1,which(Nepal_gene_filter==1)+1),]

# use 2009 as preVac reference
#Nepal_ggCaller_intermed_preVac <- Nepal_ggCaller_byYear_intermed[[5]]
Nepal_ggCaller_intermed_preVac <- Nepal_ggCaller_byYear_intermed[[1]]
# no, use all pre-vac years
#Nepal_ggCaller_intermed_preVac <- cbind(Nepal_ggCaller_byYear_intermed[[5]],Nepal_ggCaller_byYear_intermed[[8]][,-1],Nepal_ggCaller_byYear_intermed[[10]][,-1])
Nepal_ggCaller_intermed_preVac <- cbind(Nepal_ggCaller_byYear_intermed[[1]],Nepal_ggCaller_byYear_intermed[[4]][,-1],Nepal_ggCaller_byYear_intermed[[6]][,-1])

```

# create consensus gene presence absence matrices for the clusters
```{r}
cons_genomes <- function(x){
  as.double(ceiling(median(as.integer(x))))
}

# overall
Nepal_ggCaller_intermed_consensus <- data.frame(matrix(0, nrow = sum(Nepal_gene_filter)+1, ncol = no_Nepal_PP+1))
Nepal_ggCaller_intermed_consensus[1,-1] <- paste("SeqCl_",unique(Nepal_PP$GPSC),sep = "")

for (i in unique(Nepal_PP$GPSC)) {
  Nepal_ggCaller_intermed_consensus[-1,(which(unique(Nepal_PP$GPSC) == i))+1] <- apply(as.matrix(Nepal_ggCaller_intermed[-1,c(FALSE,Nepal_seq_clusters_dict[unlist(Nepal_ggCaller_intermed[1,-1])]==i)]), 1, cons_genomes)

}
Nepal_ggCaller_intermed_consensus[,1] <- Nepal_ggCaller_intermed[,1]

#by year
Nepal_ggCaller_byYear_intermed_consensus <- vector(mode = "list", length = length(Nepal_time_points_selected))

for (i in 1:length(Nepal_time_points_selected)) {
  Nepal_ggCaller_year <- data.frame(matrix(0, nrow = sum(Nepal_gene_filter)+1,ncol = no_Nepal_PP+1))
  Nepal_ggCaller_year[1,-1] <- paste("SeqCl_",unique(Nepal_PP$GPSC),sep = "")
  for (j in unique(Nepal_PP$GPSC)) {
    Nepal_ggCaller_year[-1,(which(unique(Nepal_PP$GPSC) == j))+1] <- apply(as.matrix(Nepal_ggCaller_byYear_intermed[[i]][-1,c(FALSE,Nepal_seq_clusters_dict[unlist(Nepal_ggCaller_byYear_intermed[[i]][1,-1])]==j)]), 1, cons_genomes)
  }
  Nepal_ggCaller_year[,1] <- Nepal_ggCaller_byYear_intermed[[i]][,1]
  Nepal_ggCaller_byYear_intermed_consensus[[i]] <- Nepal_ggCaller_year
}

saveRDS(Nepal_ggCaller_intermed_consensus,"Nepal_ggCaller_intermed_consensus.rds")
```

# calculate cluster frequencies for different years
```{r}
Nepal_cluster_freqs <- vector(mode = "list", length = length(Nepal_time_points_selected))

for(j in 1:length(Nepal_time_points_selected)){
  Nepal_cluster_freqs[[j]] <- rep(0, no_Nepal_PP)
  for (i in unique(Nepal_PP$GPSC)){
    Nepal_cluster_freqs[[j]][which(unique(Nepal_PP$GPSC) == i)] <- length(which(Nepal_meta_dataPP[Nepal_meta_IDtoIndPP[Nepal_PP[Nepal_PP$GPSC == i,"sample"]],]$Year==Nepal_time_points_selected[j]))
  }
}

# separate VT and NVT parts of cluster for GPSC-VT likelihood
Nepal_cluster_freqs_VT <- vector(mode = "list", length = length(Nepal_time_points_selected))
Nepal_cluster_freqs_NVT <- vector(mode = "list", length = length(Nepal_time_points_selected))

for(j in 1:length(Nepal_time_points_selected)){
  Nepal_cluster_freqs_VT[[j]] <- rep(0, no_Nepal_PP)
  for (i in unique(Nepal_PP$GPSC)){
    Nepal_cluster_freqs_VT[[j]][which(unique(Nepal_PP$GPSC) == i)] <- length(which(Nepal_meta_dataPP[Nepal_meta_IDtoIndPP[Nepal_PP[Nepal_PP$GPSC == i & Nepal_PP$VT == "VT","sample"]],]$Year==Nepal_time_points_selected[j]))
    Nepal_cluster_freqs_NVT[[j]][which(unique(Nepal_PP$GPSC) == i)] <- length(which(Nepal_meta_dataPP[Nepal_meta_IDtoIndPP[Nepal_PP[Nepal_PP$GPSC == i & Nepal_PP$VT == "NVT","sample"]],]$Year==Nepal_time_points_selected[j]))
  }
}

#Nepal_cluster_freqs[[6]] <- rep(NA, no_Nepal_PP)

# I should use a combination of 2009, 2012 and 2014 (see next code bit)
#Nepal_time_points_preVac <- c(5,8,10)
Nepal_time_points_preVac <- c(1,4,6)


Nepal_cluster_freqs_preVac <- rep(0, no_Nepal_PP)
for (i in Nepal_time_points_preVac) {
  Nepal_cluster_freqs_preVac <- Nepal_cluster_freqs_preVac + replace(Nepal_cluster_freqs[[i]],which(is.na(Nepal_cluster_freqs[[i]])),0)
}

for (i in 1:length(Nepal_time_points_selected)) {
  file_name <- paste("Nepal_cluster_freqs_", as.character(i),".rds", sep = "")
  file_name_VT <- paste("Nepal_clusterVT_freqs_", as.character(i),".rds", sep = "")
  file_name_NVT <- paste("Nepal_clusterNVT_freqs_", as.character(i),".rds", sep = "")
  #print(file_name_VT)
  saveRDS(Nepal_cluster_freqs[[i]], file_name)
  saveRDS(Nepal_cluster_freqs_VT[[i]], file_name_VT)
  saveRDS(Nepal_cluster_freqs_NVT[[i]], file_name_NVT)
}


# I want to have a start and final population GPSC-x-serotype that just stores the data for preVac and postVac 2009 and 2019
Nepal_cluster_sero_freqs_pre_post_df <- data.frame(matrix(NA, nrow = 2 * length(unique(Nepal_PP$GPSC))*length(unique(Nepal_PP$Serotype)), ncol = 4))
colnames(Nepal_cluster_sero_freqs_pre_post_df) <- c("Serotype", "GPSC", "relCount", "Time")
Nepal_cluster_sero_freqs_pre_post_df[,1] <- rep(sort(rep(unique(Nepal_PP$Serotype), length(unique(Nepal_PP$GPSC)))),2)
Nepal_cluster_sero_freqs_pre_post_df[,2] <- rep(rep(unique(Nepal_PP$GPSC), length(unique(Nepal_PP$Serotype))),2)
Nepal_cluster_sero_freqs_pre_post_df[,4] <- factor(c(rep("preVac", length(unique(Nepal_PP$GPSC))*length(unique(Nepal_PP$Serotype))), rep("postVac", length(unique(Nepal_PP$GPSC))*length(unique(Nepal_PP$Serotype)))), levels = c("preVac","postVac"))

iter1 <- 1
iter2 <- 1
for (i in sort(unique(Nepal_PP$Serotype))){
  for (j in unique(Nepal_PP$GPSC)) {
    Nepal_cluster_sero_freqs_pre_post_df[(iter1+iter2-1),3] <- length(which(Nepal_meta_dataPP[Nepal_meta_IDtoIndPP[Nepal_PP[Nepal_PP$GPSC == j,][Nepal_PP[Nepal_PP$GPSC == j,]$Serotype==i,"sample"]],]$Year==Nepal_time_points_selected[1]))/length(which(Nepal_meta_dataPP$Year == "2009"))
    #print(length(which(Nepal_meta_dataPP[Nepal_meta_IDtoIndPP[Nepal_PP[Nepal_PP$GPSC == j,][Nepal_PP[Nepal_PP$GPSC == j,]$Serotype==i,"sample"]],]$Year==Nepal_time_points_selected[1]))/length(which(Nepal_meta_dataPP$Year == "2009")))
    #print(iter1+iter2-1)
    iter1 <- iter1 + 1
  }
  iter2 <- iter2 + iter1 - 1
  iter1 <- 1
}
#iter1 <- 1
#iter2 <- 1
for (i in sort(unique(Nepal_PP$Serotype))){
  for (j in unique(Nepal_PP$GPSC)) {
    Nepal_cluster_sero_freqs_pre_post_df[(iter1+iter2-1),3] <- length(which(Nepal_meta_dataPP[Nepal_meta_IDtoIndPP[Nepal_PP[Nepal_PP$GPSC == j,][Nepal_PP[Nepal_PP$GPSC == j,]$Serotype==i,"sample"]],]$Year==Nepal_time_points_selected[11]))/length(which(Nepal_meta_dataPP$Year == "2019"))
    #print(iter1+iter2-1)
    iter1 <- iter1 + 1
  }
  iter2 <- iter2 + iter1 - 1
  iter1 <- 1
}

saveRDS(Nepal_cluster_sero_freqs_pre_post_df, "Nepal_cluster_sero_freqs_pre_post_df.rds")

# GPSC counts and serotype counts
Nepal_serotype_counts <- rep(0, length(unique(Nepal_PP$Serotype)))
names(Nepal_serotype_counts) <- sort(unique(Nepal_PP$Serotype))
for (i in sort(unique(Nepal_PP$Serotype))){
  Nepal_serotype_counts[i] <- length(which(Nepal_PP$Serotype == i))
}

Nepal_GPSC_counts <- rep(0, length(unique(Nepal_PP$GPSC)))
names(Nepal_GPSC_counts) <- unique(Nepal_PP$GPSC)
for (i in (unique(Nepal_PP$GPSC))){
  Nepal_GPSC_counts[i] <- length(which(Nepal_PP$GPSC == i))
}

saveRDS(Nepal_serotype_counts, "Nepal_serotype_counts.rds")
saveRDS(Nepal_GPSC_counts, "Nepal_GPSC_counts.rds")

Nepal_serotype_counts_selected <- Nepal_serotype_counts[which(Nepal_serotype_counts > 18)]
Nepal_GPSC_counts_selected <- Nepal_GPSC_counts[which(Nepal_GPSC_counts > 18)]

Nepal_cluster_sero_freqs_pre_post_df_selected <- data.frame(matrix(NA, nrow = 2 * length(Nepal_serotype_counts_selected)*length(Nepal_GPSC_counts_selected), ncol = 4))
colnames(Nepal_cluster_sero_freqs_pre_post_df_selected) <- c("Serotype", "GPSC", "relCount", "Time")
Nepal_cluster_sero_freqs_pre_post_df_selected[,1] <- rep(sort(rep(names(Nepal_serotype_counts_selected), length(Nepal_GPSC_counts_selected))),2)
Nepal_cluster_sero_freqs_pre_post_df_selected[,2] <- rep(rep(names(Nepal_GPSC_counts_selected), length(Nepal_serotype_counts_selected)),2)
Nepal_cluster_sero_freqs_pre_post_df_selected[,4] <- factor(c(rep("preVac", length(Nepal_GPSC_counts_selected)*length(Nepal_serotype_counts_selected)), rep("postVac", length(Nepal_GPSC_counts_selected)*length(Nepal_serotype_counts_selected))), levels = c("preVac","postVac"))

iter1 <- 1
iter2 <- 1
for (i in sort(names(Nepal_serotype_counts_selected))){
  for (j in names(Nepal_GPSC_counts_selected)) {
    Nepal_cluster_sero_freqs_pre_post_df_selected[(iter1+iter2-1),3] <- length(which(Nepal_meta_dataPP[Nepal_meta_IDtoIndPP[Nepal_PP[Nepal_PP$GPSC == j,][Nepal_PP[Nepal_PP$GPSC == j,]$Serotype==i,"sample"]],]$Year==Nepal_time_points_selected[1]))/length(which(Nepal_meta_dataPP$Year == "2009"))
    #print(iter1+iter2-1)
    iter1 <- iter1 + 1
  }
  iter2 <- iter2 + iter1 - 1
  iter1 <- 1
}
#iter1 <- 1
#iter2 <- 1
for (i in sort(names(Nepal_serotype_counts_selected))){
  for (j in names(Nepal_GPSC_counts_selected)) {
    Nepal_cluster_sero_freqs_pre_post_df_selected[(iter1+iter2-1),3] <- length(which(Nepal_meta_dataPP[Nepal_meta_IDtoIndPP[Nepal_PP[Nepal_PP$GPSC == j,][Nepal_PP[Nepal_PP$GPSC == j,]$Serotype==i,"sample"]],]$Year==Nepal_time_points_selected[11]))/length(which(Nepal_meta_dataPP$Year == "2019"))
    #print(iter1+iter2-1)
    iter1 <- iter1 + 1
  }
  iter2 <- iter2 + iter1 - 1
  iter1 <- 1
}

saveRDS(Nepal_cluster_sero_freqs_pre_post_df_selected, "Nepal_cluster_sero_freqs_pre_post_df_selected.rds")
```

```{r}
# investigating whether I am missing a lot when just using 2009 as the start population
par(mfrow = c(3,1))
barplot(Nepal_cluster_freqs[[1]][which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 0)]/sum(Nepal_cluster_freqs[[1]]), ylim = c(0, 0.1))
barplot(Nepal_cluster_freqs[[4]][which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 0)]/sum(Nepal_cluster_freqs[[4]]), col = "red")
barplot(Nepal_cluster_freqs[[6]][which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 0)]/sum(Nepal_cluster_freqs[[6]]), col = "blue")

Nepal_pre_vac_df <- data.frame(rbind(Nepal_cluster_freqs[[1]][which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)]/sum(Nepal_cluster_freqs[[1]]), Nepal_cluster_freqs[[4]][which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)]/sum(Nepal_cluster_freqs[[4]]), Nepal_cluster_freqs[[6]][which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)]/sum(Nepal_cluster_freqs[[6]])))
colnames(Nepal_pre_vac_df) <- unique(Nepal_PP$GPSC)[which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)]
rownames(Nepal_pre_vac_df) <- c("2009", "2012", "2014")

par(mfrow = c(1,1))
barplot(as.matrix(Nepal_pre_vac_df), beside = TRUE, col = c("black","red","blue"),legend = rownames(as.matrix(Nepal_pre_vac_df)))

# yes, I am clearly missing some GPSCs! 
# Especially bad is 9

Nepal_pre_and_post_vac_df <- data.frame(rbind(Nepal_cluster_freqs[[1]][which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)]/sum(Nepal_cluster_freqs[[1]]), Nepal_cluster_freqs[[4]][which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)]/sum(Nepal_cluster_freqs[[4]]), Nepal_cluster_freqs[[6]][which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)]/sum(Nepal_cluster_freqs[[6]]), Nepal_cluster_freqs[[9]][which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)]/sum(Nepal_cluster_freqs[[9]]), Nepal_cluster_freqs[[10]][which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)]/sum(Nepal_cluster_freqs[[10]]), Nepal_cluster_freqs[[11]][which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)]/sum(Nepal_cluster_freqs[[11]])))
colnames(Nepal_pre_and_post_vac_df) <- unique(Nepal_PP$GPSC)[which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)]
rownames(Nepal_pre_and_post_vac_df) <- c("2009", "2012", "2014", "2017", "2018", "2019")

barplot(as.matrix(Nepal_pre_and_post_vac_df), beside = TRUE, col = c(rep("red",3),rep("blue",3)),legend = rownames(as.matrix(Nepal_pre_and_post_vac_df)), ylim = c(0, 0.15))



# make stacked barplot - with serotypes - or maybe VTs?
Nepal_PP_with_time1 <- Nepal_PP
Nepal_PP_with_time1$Time <- Nepal_SeqYear_dict[Nepal_PP$sample]

Nepal_VT_dict <- rep(0, length(Nepal_PP$sample))
Nepal_VT_dict[which(Nepal_PP$VT == "VT")] <- 1
names(Nepal_VT_dict) <- Nepal_PP$sample

Nepal_GPSC_dict <- Nepal_PP$GPSC
names(Nepal_GPSC_dict) <- Nepal_PP$sample

Nepal_GPSC_ind <- 1:length(which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1))
names(Nepal_GPSC_ind) <- unique(Nepal_PP$GPSC)[(which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1))]

Nepal_GPSC_ind_all <- 1:length(unique(Nepal_PP$GPSC))
names(Nepal_GPSC_ind_all) <- unique(Nepal_PP$GPSC)

Nepal_SeqYear_ind <- 1:6
names(Nepal_SeqYear_ind) <- unique(Nepal_SeqYear_dict)

Nepal_PP_with_time <- data.frame(matrix(NA, nrow = length(which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)) * 2 * length(unique(Nepal_SeqYear_dict)), ncol = 4))
Nepal_PP_with_time[,1] <- rep(unique(Nepal_PP$GPSC)[(which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1))], 2 * length(unique(Nepal_SeqYear_dict)))
Nepal_PP_with_time[,2] <- rep(c(rep("NVT", length(which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1))), rep("VT", length(which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)))), length(unique(Nepal_SeqYear_dict)))
Nepal_PP_with_time[,3] <- sort(rep(unique(Nepal_SeqYear_dict), length(which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)) * 2))
Nepal_PP_with_time[,4] <- rep(0, length(which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)) * 2 * length(unique(Nepal_SeqYear_dict)))

for (i in Nepal_PP$sample) {
  #print(i)
  if(is.element(Nepal_GPSC_ind_all[Nepal_GPSC_dict[i]], which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1))){
   pos <- ((Nepal_SeqYear_ind[as.character(Nepal_SeqYear_dict[i])]) -1) * length(which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)) * 2 + (Nepal_VT_dict[i]) * length(which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1)) + ((Nepal_GPSC_ind[as.character(Nepal_GPSC_dict[i])]) -1)
  Nepal_PP_with_time[pos, 4] <- Nepal_PP_with_time[pos, 4] + 1 
  }
}
colnames(Nepal_PP_with_time) <- c("GPSC","VT","Time","value")

ggplot(Nepal_PP_with_time, aes(x=Time, y=value, fill = VT)) + 
  geom_bar(stat = "identity") + facet_grid(.~GPSC) + theme_bw()

# this should really be frequency, not absolute counts
Nepal_PP_with_time_rel <- Nepal_PP_with_time
for (i in unique(Nepal_PP_with_time_rel$Time)) {
  ind <- which(Nepal_PP_with_time_rel$Time == i)
  sum <- sum(Nepal_PP_with_time_rel$value[which(Nepal_PP_with_time_rel$Time == i)])
  Nepal_PP_with_time_rel$value[which(Nepal_PP_with_time_rel$Time == i)] <- Nepal_PP_with_time_rel$value[which(Nepal_PP_with_time_rel$Time == i)] / sum
}
ggplot(Nepal_PP_with_time_rel, aes(x=Time, y=value, fill = VT)) + 
  geom_bar(stat = "identity") + facet_grid(.~GPSC) + theme_bw()

# now pick a couple of interesting GPSCs
Nepal_PP_with_time_rel_selected <- data.frame(matrix(NA, nrow = 5 * 2 * 6, ncol = 4))
colnames(Nepal_PP_with_time_rel_selected) <- colnames(Nepal_PP_with_time_rel)
selected_GPSCs <- (unique(Nepal_PP$GPSC)[(which(Nepal_cluster_freqs[[1]] + Nepal_cluster_freqs[[4]] + Nepal_cluster_freqs[[6]] > 1))])[c(12, 4, 5, 11, 6)]
for (i in 1:length(selected_GPSCs)) {
  Nepal_PP_with_time_rel_selected[((i-1)*2*6+1):((i)*2*6),] <- Nepal_PP_with_time_rel[which(Nepal_PP_with_time_rel$GPSC == selected_GPSCs[i]),]
}

Nepal_PP_with_time_rel_selected$GPSC_order = factor(Nepal_PP_with_time_rel_selected$GPSC, levels=c('101','45','26','12','11'))

#group.colors <- c(VT = "#800000", NVT = "#000080")
group.colors <- c(VT = "#e69f00", NVT = "#009e73")

ggplot(Nepal_PP_with_time_rel_selected, aes(x=Time, y=value, fill = VT)) + 
  geom_bar(stat = "identity") + facet_grid(.~GPSC_order) + theme_bw() +
  scale_fill_manual(values=group.colors)

Nepal_PP_with_time_rel_selected$Time = factor(Nepal_PP_with_time_rel_selected$Time)

ggplot(Nepal_PP_with_time_rel_selected, aes(x=Time, y=value, fill = VT)) + 
  geom_bar(stat = "identity") + facet_grid(.~GPSC_order) + theme_bw() + theme(text = element_text(size = 20)) +
  scale_fill_manual(values=group.colors)

```



#create initial population
```{r}
### create initial population that is based on the first time point data set
# sample from it with an Poisson process
expand_factor <- 15000 / sum(Nepal_cluster_freqs_preVac)
exp_noise <- 10
Nepal_model_start_pop <- (sapply((Nepal_cluster_freqs_preVac + rexp(n = length(Nepal_cluster_freqs_preVac), rate = exp_noise)) * expand_factor, rpois, n=1))

# visual check for the "sampling"
plot(Nepal_cluster_freqs_preVac/sum(Nepal_cluster_freqs_preVac))
points(Nepal_model_start_pop/sum(Nepal_model_start_pop), col = "red")

# aha, I compared the data but 90% of the data are actually coming from 2009 (of the preVac era)
# so it would be much better to only use that time point I think

Nepal_model_start_pop <- (sapply((Nepal_cluster_freqs[[5]] + rexp(n = length(Nepal_cluster_freqs[[5]]), rate = exp_noise)) * expand_factor, rpois, n=1))
# no, do this based on all three pre-vac years
Nepal_model_start_pop <- (sapply((Nepal_cluster_freqs_preVac + rexp(n = length(Nepal_cluster_freqs_preVac), rate = exp_noise)) * expand_factor, rpois, n=1))

# visual check for the "sampling"
plot(Nepal_cluster_freqs_preVac/sum(Nepal_cluster_freqs_preVac))
points(Nepal_model_start_pop/sum(Nepal_model_start_pop), col = "red")

saveRDS(Nepal_model_start_pop, "Nepal_model_start_pop.rds")
```


# calculate delta ranking
# I will do 2019 - 2009
```{r}
# For strong and weak selection in the model, I need to calculate the beta statistics.

# calculate gene frequencies first, separate for three time points
#Nepal_gene_freq_preVac <- apply(Nepal_ggCaller_intermed_preVac[-1,-1], 1, sum_as_int)
#Nepal_gene_freq_postVac <- apply(Nepal_ggCaller_byYear_intermed[[length(Nepal_ggCaller_byYear_intermed)]][-1,-1], 1, sum_as_int)
Nepal_gene_freq_preVac <- apply(Nepal_ggCaller_intermed_preVac[-1,-1], 1, sum_as_int) # this now includes all three pre-vac samples
#Nepal_gene_freq_postVac <- apply(Nepal_ggCaller_byYear_intermed[[15]][-1,-1], 1, sum_as_int)
Nepal_gene_freq_postVac <- apply(Nepal_ggCaller_byYear_intermed[[11]][-1,-1], 1, sum_as_int)


# first, calculate pre/peri and post vacc frequencies of genes:
#Nepal_gene_freq_preVac_rel <- Nepal_gene_freq_preVac/sum(Nepal_gene_freq_preVac)
#Nepal_gene_freq_postVac_rel <- Nepal_gene_freq_postVac/sum(Nepal_gene_freq_postVac)
# no, I don't want to divide by the sum of the number of genes, I want to divide by number of seqs
Nepal_gene_freq_preVac_rel <- Nepal_gene_freq_preVac/(ncol(Nepal_ggCaller_intermed_preVac[-1,-1]))
Nepal_gene_freq_postVac_rel <- Nepal_gene_freq_postVac/(ncol(Nepal_ggCaller_byYear_intermed[[11]][-1,-1]))

# calculate delta statistic (refer to Corander et al. for more info)
Nepal_delta_data <- (Nepal_gene_freq_postVac_rel - Nepal_gene_freq_preVac_rel) ^ 2 / (1 - Nepal_gene_freq_preVac_rel * (1 - Nepal_gene_freq_preVac_rel))
Nepal_delta_ranking <- rank(Nepal_delta_data)

saveRDS(Nepal_delta_ranking,"Nepal_delta_ranking.rds")
saveRDS(Nepal_delta_data,"Nepal_delta_data.rds")
```

### Data processing for PPxSero model
```{r}
Nepal_clusters <- readRDS("Nepal_PP.rds")
Nepal_model_start_pop <- readRDS("Nepal_model_start_pop.rds")
Nepal_clusters_no <- length(unique(Nepal_clusters$GPSC))
Nepal_avg_cluster_freq <- rep(1/Nepal_clusters_no, Nepal_clusters_no)


PPxSero_matrix <- matrix(0, nrow = length(unique(Nepal_clusters$GPSC)), ncol = length(unique(Nepal_clusters$Serotype)))
# sort(unique(PopPUNK_clusters$Serotype))
# [1] "10A"   "11A"   "14"    "15A"   "15B/C" "15F"   "16F"   "17F"   "18C"   "19A"   "19F"   "21"    "22F"   "23A"
#[15] "23B"   "23F"   "3"     "31"    "33F"   "34"    "35B"   "35F"   "37"    "38"    "6A"    "6B"    "6C"    "7C" 
#[29] "7F"    "9N"    "9V"    "NT"
# now, 74 serotypes

SerotypeToIndex_dict <- 1:length(unique(Nepal_clusters$Serotype))
names(SerotypeToIndex_dict) <- sort(unique(Nepal_clusters$Serotype))

IndexToSerotype_dict <- sort(unique(Nepal_clusters$Serotype))
names(IndexToSerotype_dict) <- 1:length(unique(Nepal_clusters$Serotype))

PPclustToIndex_dict <- 1:length(unique(Nepal_clusters$GPSC))
names(PPclustToIndex_dict) <- sort(unique(Nepal_clusters$GPSC))

for (i in 1:nrow(Nepal_clusters)) {
  PPxSero_matrix[PPclustToIndex_dict[as.character(Nepal_clusters$GPSC[i])],SerotypeToIndex_dict[Nepal_clusters$Serotype[i]]] <- 1
}


PPxSero_matrix_prob2 <-  matrix(0, nrow = length(unique(Nepal_clusters$GPSC)), ncol = length(unique(Nepal_clusters$Serotype)))
for (i in 1:length(unique(Nepal_clusters$GPSC))) {
  PPxSero_matrix_prob2[i,] <- ((PPxSero_matrix[i,] * Nepal_avg_cluster_freq[i])/sum(PPxSero_matrix[i,]))
}
saveRDS(PPxSero_matrix_prob2, "Nepal_PPsero_mig.rds")
# new migration matrix that is closer to original one

# start population:
# I could maybe use the existing PP_model_start_pop and distribute it to the different serotypes?
PPsero_startpop2 <-  matrix(0, nrow = length(unique(Nepal_clusters$GPSC)), ncol = length(unique(Nepal_clusters$Serotype)))
for (i in 1:length(unique(Nepal_clusters$GPSC))) {
  PPsero_startpop2[i,] <- round((PPxSero_matrix[i,] * Nepal_model_start_pop[i])/sum(PPxSero_matrix[i,]))
}
saveRDS(PPsero_startpop2, "Nepal_PPsero_startpop.rds")
# this is very close to the original PP_model_start_pop
# maybe I am giving out too much information because the serotype distribution is exactly the real one?


# serotype VT vector
SeroVT <- rep(0, length(unique(Nepal_clusters$Serotype)))
for (i in 1:nrow(Nepal_clusters)) {
  SeroVT[SerotypeToIndex_dict[Nepal_clusters$Serotype[i]]] <- as.integer(Nepal_clusters$VT[i]=="VT")
}
names(SeroVT) <- names(SerotypeToIndex_dict)
saveRDS(SeroVT, "Nepal_SeroVT.rds")

### vt vector that includes 6A (which appears to have cross immunity with 6B and is therefore affected by the vaccine, e.g. https://pmc.ncbi.nlm.nih.gov/articles/PMC5848315/)
SeroVT_6A <- SeroVT
SeroVT_6A["6A"] <- 1
saveRDS(SeroVT_6A, "Nepal_SeroVT_6A.rds")
```

# UK
# split ggCaller gene presence absence matrix into different time points
```{r}
UK_ggCaller_byYear <- vector(mode = "list", length = length(UK_time_points))

for (i in 1:length(UK_time_points)) {
  #UK_ggCaller_year <- data.frame(matrix(0, nrow = nrow(UK_ggCallerPP_bool), ncol = length(which(UK_metadata$Winter==UK_winters[i]))+1)) # initialize data frame
  UK_ggCaller_year <- data.frame(matrix(0, nrow = nrow(UK_ggCallerPP_bool), ncol = length(which(UK_SeqYear_dict==UK_time_points[i]))+1)) # initialize data frame
  UK_ggCaller_year[1,-1] <- UK_ggCallerPP_bool[1,c(FALSE,UK_SeqYear_dict[unlist(UK_ggCallerPP_bool[1,-1])]==UK_time_points[i])] # fill in first row (seq names)
  UK_ggCaller_year[-1,1] <- UK_ggCallerPP_bool[-1,1] # fill in first column (gene cluster names)
  UK_ggCaller_year[-1,-1] <- UK_ggCallerPP_bool[-1,c(FALSE,UK_SeqYear_dict[unlist(UK_ggCallerPP_bool[1,-1])]==UK_time_points[i])] # fill in presence absence information
  #print(i)
  UK_ggCaller_byYear[[i]] <- UK_ggCaller_year
}
```

# find intermediate frequency genes
```{r}
sum_as_int <- function(x){
  sum(as.integer(x))
}
UK_time_points_preVac <- 1
UK_gene_freq_preVac <- rep(0, nrow(UK_ggCaller_byYear[[1]])-1)
preVac_datapoints <- 0
for (i in 1:length(UK_time_points_preVac)) {
  UK_gene_freq_preVac <- UK_gene_freq_preVac + apply(UK_ggCaller_byYear[[i]][-1,-1],1, sum_as_int)
  preVac_datapoints <- preVac_datapoints + length(UK_ggCaller_byYear[[i]][1,-1])
}
UK_gene_freq_preVac <- UK_gene_freq_preVac / preVac_datapoints

UK_gene_filter <- as.integer(UK_gene_freq_preVac<=0.95 & UK_gene_freq_preVac>=0.05)
# sum(UK_gene_filter)
# now (joined analysis) 2048
# 1614 intermediate-frequency genes (of 6526)

UK_ggC_intermed_gene_freqs <- UK_gene_freq_preVac[which(UK_gene_filter==1)]
saveRDS(UK_ggC_intermed_gene_freqs, file = "UK_ggC_intermed_gene_freqs.rds")

UK_ggC_all_gene_freqs <- UK_gene_freq_preVac
saveRDS(UK_ggC_all_gene_freqs, file = "UK_ggC_all_gene_freqs.rds")

UK_ggC_all_gene_names <- UK_ggCaller_byYear[[1]][-1,1]
saveRDS(UK_ggC_all_gene_names, file = "UK_ggC_all_gene_names.rds")
```

# create intermediate frequency dataframes
```{r}
UK_ggCaller_byYear_intermed <- vector(mode = "list", length = length(UKPP_time_points))

for (i in 1:length(UKPP_time_points)) {
  UK_ggCaller_year <- data.frame(matrix(0, nrow = sum(UK_gene_filter)+1,ncol = length(which(UK_metadata$Year==UKPP_time_points[i]))+1))
  if(ncol(UK_ggCaller_byYear[[i]])>1){
    UK_ggCaller_year <-  UK_ggCaller_byYear[[i]][c(1,which(UK_gene_filter==1)+1),]
  }
  else{
    UK_ggCaller_year <-  data.frame(matrix(UK_ggCaller_byYear[[6]][c(1,which(UK_gene_filter==1)+1),1], nrow = sum(UK_gene_filter)+1,ncol = 1))
  }
  rownames(UK_ggCaller_year) <- c("name", which(UK_gene_filter==1))
  UK_ggCaller_byYear_intermed[[i]] <- UK_ggCaller_year 
}

UK_ggCaller_intermed <- data.frame(matrix(0, nrow = sum(UK_gene_filter)+1,ncol = ncol(UK_ggCallerPP_bool)))
UK_ggCaller_intermed <-  UK_ggCallerPP_bool[c(1,which(UK_gene_filter==1)+1),]

# create pre_Vac intermed frequency data frame
UK_ggCaller_intermed_preVac <- data.frame(matrix(0, nrow = sum(UK_gene_filter)+1,ncol = (sum(unlist(lapply(UK_ggCaller_byYear_intermed,ncol))[1:length(UK_time_points_preVac)]) - length(UK_time_points_preVac) + 1)))
col_count <- 2
for (i in 1:length(UK_time_points_preVac)) {
  if(ncol(UK_ggCaller_byYear_intermed[[i]])>1){
    UK_ggCaller_intermed_preVac[,(col_count:(col_count + ncol(UK_ggCaller_byYear_intermed[[i]])-1))] <- UK_ggCaller_byYear_intermed[[i]][,-1]
    col_count <- col_count + ncol(UK_ggCaller_byYear_intermed[[i]])
  }
}
UK_ggCaller_intermed_preVac[,1] <-  UK_ggCaller_byYear_intermed[[1]][,1]
rownames(UK_ggCaller_intermed_preVac) <- c("name", which(UK_gene_filter==1))
```

# create consensus gene presence absence matrices for the clusters
```{r}
cons_genomes <- function(x){
  as.double(ceiling(median(as.integer(x))))
}

# overall
UK_ggCaller_intermed_consensus <- data.frame(matrix(0, nrow = sum(UK_gene_filter)+1, ncol = no_UK_PP+1))
UK_ggCaller_intermed_consensus[1,-1] <- paste("SeqCl_",unique(UK_PP$Cluster),sep = "")

for (i in unique(UK_PP$Cluster)) {
  UK_ggCaller_intermed_consensus[-1,(which(unique(UK_PP$Cluster) == i))+1] <- apply(as.matrix(UK_ggCaller_intermed[-1,c(FALSE,UK_seq_clusters_dict[unlist(UK_ggCaller_intermed[1,-1])]==i)]), 1, cons_genomes)

}
UK_ggCaller_intermed_consensus[,1] <- UK_ggCaller_intermed[,1]

#by year
UK_ggCaller_byYear_intermed_consensus <- vector(mode = "list", length = length(UK_time_points))

for (i in 1:length(UK_time_points)) {
  UK_ggCaller_year <- data.frame(matrix(0, nrow = sum(UK_gene_filter)+1,ncol = no_UK_PP+1))
  UK_ggCaller_year[1,-1] <- paste("SeqCl_",unique(UK_PP$Cluster),sep = "")
  for (j in unique(UK_PP$Cluster)) {
    UK_ggCaller_year[-1,(which(unique(UK_PP$Cluster) == j))+1] <- apply(as.matrix(UK_ggCaller_byYear_intermed[[i]][-1,c(FALSE,UK_seq_clusters_dict[unlist(UK_ggCaller_byYear_intermed[[i]][1,-1])]==j)]), 1, cons_genomes)
  }
  UK_ggCaller_year[,1] <- UK_ggCaller_byYear_intermed[[i]][,1]
  UK_ggCaller_byYear_intermed_consensus[[i]] <- UK_ggCaller_year
}

saveRDS(UK_ggCaller_intermed_consensus,"UK_ggCaller_intermed_consensus.rds")

UK_ggC_intermed_gene_names <- UK_ggCaller_intermed_consensus[-1,1]
saveRDS(UK_ggC_intermed_gene_names, file = "UK_ggC_intermed_gene_names.rds")
```

# Calculate VT/NVT for clusters
```{r}
# introduction of PCV7 in 2006 and PCV13 in 2010

# create dictionary for VTs of the PCV7
# 1, 4, 5, 6b, 7f, 9V, 14, 18c, 19f and 23f serotypes (source: NCBI)
#PCV7: 4, 6B, 9V, 14, 18C, 19F and 23F. (source: https://journals.plos.org/plospathogens/article?id=10.1371/journal.ppat.1006966#sec018)
# PCV13: (1, 3, 4, 5, 6A and 6B, 7F, 9V, 14, 19A, and 19F, 18C, and 23F) (source: https://www.ncbi.nlm.nih.gov/books/NBK507794/)
PCV13_VTs <- rep(0,length(unique(UK_metadata$Serotype)))
names(PCV13_VTs) <- unique(UK_metadata$Serotype)
PCV13_VTs[c("1", "3", "4","5","6A","6B", "7F", "9V", "14", "18C", "19A", "19F", "23F")] <- 1

PCV7_VTs_no6A <- rep(0,length(unique(UK_metadata$Serotype)))
names(PCV7_VTs_no6A) <- unique(UK_metadata$Serotype)
PCV7_VTs_no6A[c("4","6B", "9V", "14", "18C", "19F", "23F")] <- 1

# add 6A to PCV7 because there is strong cross-immunity btw PVC7 and 6A (4.Croucher, N. J. et al. Population genomics of post-vaccine changes in pneumococcal epidemiology. Nat. Genet. 45, 656–663 (2013).)
PCV7_VTs <- rep(0,length(unique(UK_metadata$Serotype)))
names(PCV7_VTs) <- unique(UK_metadata$Serotype)
PCV7_VTs[c("4","6A", "6B", "9V", "14", "18C", "19F", "23F")] <- 1

UK_SeqToSero_dict <- UK_metadata$Serotype
names(UK_SeqToSero_dict) <- UK_metadata$Accession

# calculate Vaccine Type consensus for clusters
UK_VT_no6A <- rep(0, no_UK_PP)

for (i in unique(UK_PP$Cluster)){
  UK_VT_no6A[(which(unique(UK_PP$Cluster) == i))] <- ceiling(median(as.integer(PCV7_VTs_no6A[UK_SeqToSero_dict[UK_PP[UK_PP$Cluster == i,"Taxon"]]])))
}

# calculate Vaccine Type consensus for clusters
UK_VT <- rep(0, no_UK_PP)

for (i in unique(UK_PP$Cluster)){
  UK_VT[(which(unique(UK_PP$Cluster) == i))] <- ceiling(median(as.integer(PCV7_VTs[UK_SeqToSero_dict[UK_PP[UK_PP$Cluster == i,"Taxon"]]])))
}
# 1 means VT
# 0 means NVT
# if cluster is 50/50 it will count as VT

# calculate Vaccine Type consensus for clusters for PCV13
UK_VT2 <- rep(0, no_UK_PP)

for (i in unique(UK_PP$Cluster)){
  UK_VT2[(which(unique(UK_PP$Cluster) == i))] <- ceiling(median(as.integer(PCV13_VTs[UK_SeqToSero_dict[UK_PP[UK_PP$Cluster == i,"Taxon"]]])))
}
```

```{r}
#save VTs
saveRDS(UK_VT_no6A, file = "UK_VT_no6A.rds")
saveRDS(UK_VT, file = "UK_VT.rds")
saveRDS(UK_VT2, file = "UK_VT2.rds")
```

# calculate cluster frequencies for different years
```{r}
UK_cluster_freqs <- vector(mode = "list", length = length(UK_time_points))

for(j in 1:length(UK_time_points)){
  UK_cluster_freqs[[j]] <- rep(NA, no_UK_PP)
  for (i in unique(UK_PP$Cluster)){
    UK_cluster_freqs[[j]][which(unique(UK_PP$Cluster) == i)] <- length(which(UK_metadata[UK_meta_IDtoInd[UK_PP[UK_PP$Cluster == i,"Taxon"]],]$Year==UK_time_points[j]))
  }
}

for(j in 1:length(UK_time_points)){
  if(sum(UK_cluster_freqs[[j]])==0){
    UK_cluster_freqs[[j]] <- rep(NA, no_UK_PP)
  }
}

#for (i in 1:length(UK_time_points)) {
#  file_name <- paste("UK_cluster_freqs_", as.character(i),".rds", sep = "")
#  saveRDS(UK_cluster_freqs[[i]], file_name)
#}

### actually I'd want winters, not years (have not tested this, need to try fitting)
# and double check: I think some of them are incorrectly assigned (yes, fixed that above)
UK_cluster_freqs_winter <- vector(mode = "list", length = length(unique(UK_winters)))

for(j in 1:length(unique(UK_winters))){ # "unique" was missing, creating hundreds of list entries
  UK_cluster_freqs_winter[[j]] <- rep(NA, no_UK_PP)
  for (i in unique(UK_PP$Cluster)){
    UK_cluster_freqs_winter[[j]][which(unique(UK_PP$Cluster) == i)] <- length(which(UK_metadata[UK_meta_IDtoInd[UK_PP[UK_PP$Cluster == i,"Taxon"]],]$Winter==unique(UK_winters)[j]))
  }
}

for(j in 1:length(UK_time_points)){
  if(sum(UK_cluster_freqs_winter[[j]])==0){
    UK_cluster_freqs_winter[[j]] <- rep(NA, no_UK_PP)
  }
}

for (i in 1:length(UK_time_points)) {
  file_name <- paste("UK_cluster_freqs_", as.character(i),".rds", sep = "")
  print(file_name)
  saveRDS(UK_cluster_freqs_winter[[i]], file_name)
}

#UK_cluster_freqs[[6]] <- rep(NA, no_UK_PP)
# 901 data points in total
UK_cluster_freqs_preVac <- rep(0, no_UK_PP)
for (i in 1:length(UK_time_points_preVac)) {
  UK_cluster_freqs_preVac <- UK_cluster_freqs_preVac + replace(UK_cluster_freqs[[i]],which(is.na(UK_cluster_freqs[[i]])),0)
}

# repeat this for winter
UK_cluster_freqs_preVac <- rep(0, no_UK_PP)
for (i in 1:length(UK_time_points_preVac)) {
  UK_cluster_freqs_preVac <- UK_cluster_freqs_preVac + replace(UK_cluster_freqs_winter[[i]],which(is.na(UK_cluster_freqs_winter[[i]])),0)
}

# do the same for Strain x Serotype model
UK_cluster_freqs_sero <- vector(mode = "list", length = length(UK_time_points))

for(j in 1:length(UK_time_points)){
  UK_cluster_freqs_sero[[j]] <- matrix(NA, ncol =  length(unique(UK_PP$Serotype)), nrow = no_UK_PP)
  for (i in unique(UK_PP$Cluster)){
    for (k in unique(UK_PP$Serotype)) {
      UK_cluster_freqs_sero[[j]][which(unique(UK_PP$Cluster) == i),which(unique(UK_PP$Serotype) == k)] <- length(which(UK_metadata[UK_meta_IDtoInd[UK_PP[UK_PP$Cluster == i,"Taxon"]],][UK_metadata[UK_meta_IDtoInd[UK_PP[UK_PP$Cluster == i,"Taxon"]],]$Serotype==k,]$Year==UK_time_points[j]))
    }
  }
}
for(j in 1:length(UK_time_points)){
  if(sum(UK_cluster_freqs_sero[[j]])==0){
    UK_cluster_freqs_sero[[j]] <- matrix(NA, ncol =  length(unique(UK_PP$Serotype)), nrow = no_UK_PP)
  }
}

#for (i in 1:length(UK_time_points)) {
#  file_name <- paste("UK_cluster_freqs_sero_", as.character(i),".rds", sep = "")
#  saveRDS(UK_cluster_freqs_sero[[i]], file_name)
#}

# and repeat for winter
UK_cluster_freqs_sero <- vector(mode = "list", length = length(UK_time_points))

for(j in 1:length(UK_time_points)){
  UK_cluster_freqs_sero[[j]] <- matrix(NA, ncol =  length(unique(UK_PP$Serotype)), nrow = no_UK_PP)
  for (i in unique(UK_PP$Cluster)){
    for (k in unique(UK_PP$Serotype)) {
      UK_cluster_freqs_sero[[j]][which(unique(UK_PP$Cluster) == i),which(unique(UK_PP$Serotype) == k)] <- length(which(UK_metadata[UK_meta_IDtoInd[UK_PP[UK_PP$Cluster == i,"Taxon"]],][UK_metadata[UK_meta_IDtoInd[UK_PP[UK_PP$Cluster == i,"Taxon"]],]$Serotype==k,]$Winter==UK_winters[j]))
    }
  }
}
for(j in 1:length(UK_time_points)){
  if(sum(UK_cluster_freqs_sero[[j]])==0){
    UK_cluster_freqs_sero[[j]] <- matrix(NA, ncol =  length(unique(UK_PP$Serotype)), nrow = no_UK_PP)
  }
}

for (i in 1:length(UK_time_points)) {
  file_name <- paste("UK_cluster_freqs_sero_", as.character(i),".rds", sep = "")
  saveRDS(UK_cluster_freqs_sero[[i]], file_name)
}
```

#create initial population
```{r}
### create initial population that is based on the first time point data set
# sample from it with an Poisson process
expand_factor <- 15000 / sum(UK_cluster_freqs_preVac)
exp_noise <- 10
UK_model_start_pop <- (sapply((UK_cluster_freqs_preVac + rexp(n = length(UK_cluster_freqs_preVac), rate = exp_noise)) * expand_factor, rpois, n=1))

# visual check for the "sampling"
plot(UK_cluster_freqs_preVac/sum(UK_cluster_freqs_preVac))
points(UK_model_start_pop/sum(UK_model_start_pop), col = "red")

expand_factor <- 15000 / sum(UK_cluster_freqs_winter[[1]])
exp_noise <- 10
#UK_model_start_pop <- (sapply((UK_cluster_freqs[[1]] + rexp(n = length(UK_cluster_freqs[[1]]), rate = exp_noise)) * expand_factor, rpois, n=1))
# change to winter
UK_model_start_pop <- (sapply((UK_cluster_freqs_winter[[1]] + rexp(n = length(UK_cluster_freqs_winter[[1]]), rate = exp_noise)) * expand_factor, rpois, n=1))


# visual check for the "sampling"
plot(UK_cluster_freqs_winter[[1]]/sum(UK_cluster_freqs_winter[[1]]))
points(UK_model_start_pop/sum(UK_model_start_pop), col = "red")

saveRDS(UK_model_start_pop, "UK_model_start_pop.rds")
```


# calculate delta ranking
# I will do last data point - all pre_vac data points combined
```{r}
# For strong and weak selection in the model, I need to calculate the beta statistics.

# calculate gene frequencies first, separate for three time points
UK_gene_freq_preVac <- apply(UK_ggCaller_intermed_preVac[-1,-1], 1, sum_as_int)
UK_gene_freq_postVac <- apply(UK_ggCaller_byYear_intermed[[length(UK_ggCaller_byYear_intermed)]][-1,-1], 1, sum_as_int) + apply(UK_ggCaller_byYear_intermed[[length(UK_ggCaller_byYear_intermed)-1]][-1,-1], 1, sum_as_int) + apply(UK_ggCaller_byYear_intermed[[length(UK_ggCaller_byYear_intermed)-2]][-1,-1], 1, sum_as_int)
# why is this the sum of all post vacc values?
# I think I want just the last timepoint
UK_gene_freq_postVac <- apply(UK_ggCaller_byYear_intermed[[length(UK_ggCaller_byYear_intermed)]][-1,-1], 1, sum_as_int)


# first, calculate pre/peri and post vacc frequencies of genes:
#UK_gene_freq_preVac_rel <- UK_gene_freq_preVac/sum(UK_gene_freq_preVac) # no, I don't want to divide by the sum - I want to divide by the number of sequences
# UK_gene_freq_postVac_rel <- UK_gene_freq_postVac/sum(UK_gene_freq_postVac) # same
UK_gene_freq_preVac_rel <- UK_gene_freq_preVac/(ncol(UK_ggCaller_intermed_preVac) -1)
#UK_gene_freq_postVac_rel <- UK_gene_freq_postVac/((ncol(UK_ggCaller_byYear_intermed[[length(UK_ggCaller_byYear_intermed)]]) -1) + (ncol(UK_ggCaller_byYear_intermed[[length(UK_ggCaller_byYear_intermed)-1]]) -1) + (ncol(UK_ggCaller_byYear_intermed[[length(UK_ggCaller_byYear_intermed)-2]]) -1))
UK_gene_freq_postVac_rel <- UK_gene_freq_postVac/((ncol(UK_ggCaller_byYear_intermed[[length(UK_ggCaller_byYear_intermed)]]) -1))

# calculate delta statistic (refer to Corander et al. for more info)
UK_delta_data <- (UK_gene_freq_postVac_rel - UK_gene_freq_preVac_rel) ^ 2 / (1 - UK_gene_freq_preVac_rel * (1 - UK_gene_freq_preVac_rel))
UK_delta_ranking <- rank(UK_delta_data)

saveRDS(UK_delta_ranking,"UK_delta_ranking.rds")
saveRDS(UK_delta_data,"UK_delta_data.rds")

# test UK vacc1 vs vacc2 delta
UK_gene_freq_midVac <- apply(UK_ggCaller_byYear_intermed[[4]][-1,-1], 1, sum_as_int)
UK_gene_freq_postVac2 <- apply(UK_ggCaller_byYear_intermed[[7]][-1,-1], 1, sum_as_int)
UK_gene_freq_midVac_rel <- UK_gene_freq_midVac/(ncol(UK_ggCaller_byYear_intermed[[4]]) -1)
UK_gene_freq_postVac2_rel <- UK_gene_freq_postVac2/(ncol(UK_ggCaller_byYear_intermed[[7]]) -1)

UK_delta_data_vacc1 <- (UK_gene_freq_midVac_rel - UK_gene_freq_preVac_rel) ^ 2 / (1 - UK_gene_freq_preVac_rel * (1 - UK_gene_freq_preVac_rel))
UK_delta_ranking_vacc1 <- rank(UK_delta_data_vacc1)

UK_delta_data_vacc2 <- (UK_gene_freq_postVac2_rel - UK_gene_freq_midVac) ^ 2 / (1 - UK_gene_freq_midVac * (1 - UK_gene_freq_midVac))
UK_delta_ranking_vacc2 <- rank(UK_delta_data_vacc2)

UK_delta_data_vacc02 <- (UK_gene_freq_postVac2_rel - UK_gene_freq_preVac_rel) ^ 2 / (1 - UK_gene_freq_preVac_rel * (1 - UK_gene_freq_preVac_rel))
UK_delta_ranking_vacc02 <- rank(UK_delta_data_vacc02)

plot(UK_delta_ranking_vacc1, UK_delta_ranking_vacc2)
plot(UK_delta_ranking_vacc1, UK_delta_ranking_vacc02)



plot(UK_delta_ranking, UK_delta_ranking_vacc02) # this is the difference btw taking all three post vacc timepoints vs just one...

```

### Data processing for PPxSero model
```{r}
UK_clusters <- readRDS("UK_PP.rds")
UK_model_start_pop <- readRDS("UK_model_start_pop.rds")
UK_clusters_no <- length(unique(UK_clusters$Cluster))
UK_avg_cluster_freq <- rep(1/UK_clusters_no, UK_clusters_no)


PPxSero_matrix <- matrix(0, nrow = length(unique(UK_clusters$Cluster)), ncol = length(unique(UK_clusters$Serotype)))
# sort(unique(UK_clusters$Serotype))
# [1] "10A"   "11A"   "14"    "15A"   "15B/C" "15F"   "16F"   "17F"   "18C"   "19A"   "19F"   "21"    "22F"   "23A"
#[15] "23B"   "23F"   "3"     "31"    "33F"   "34"    "35B"   "35F"   "37"    "38"    "6A"    "6B"    "6C"    "7C" 
#[29] "7F"    "9N"    "9V"    "NT"

SerotypeToIndex_dict <- 1:length(unique(UK_clusters$Serotype))
names(SerotypeToIndex_dict) <- (unique(UK_clusters$Serotype))

IndexToSerotype_dict <- sort(unique(UK_clusters$Serotype))
names(IndexToSerotype_dict) <- 1:length(unique(UK_clusters$Serotype))

PPclustToIndex_dict <- 1:length(unique(UK_clusters$Cluster))
names(PPclustToIndex_dict) <- (unique(UK_clusters$Cluster))

for (i in 1:nrow(UK_clusters)) {
  PPxSero_matrix[PPclustToIndex_dict[as.character(UK_clusters$Cluster[i])],SerotypeToIndex_dict[UK_clusters$Serotype[i]]] <- 1
}
# matrix shows information whether PP and serotype co-occur in the dataset (1=TRUE)
# this can be used as the migration matrix:
#PPxSero_matrix_prob <- PPxSero_matrix / sum(PPxSero_matrix)
#PPxSero_matrix_prob <- matrix(sapply(PPxSero_matrix_prob,as.double), nrow = nrow(PPxSero_matrix), ncol = ncol(PPxSero_matrix))
#saveRDS(PPxSero_matrix_prob, "Navajo_PPsero_mig.rds")


PPxSero_matrix_prob2 <-  matrix(0, nrow = length(unique(UK_clusters$Cluster)), ncol = length(unique(UK_clusters$Serotype)))
for (i in 1:length(unique(UK_clusters$Cluster))) {
  PPxSero_matrix_prob2[i,] <- ((PPxSero_matrix[i,] * UK_avg_cluster_freq[i])/sum(PPxSero_matrix[i,]))
}
saveRDS(PPxSero_matrix_prob2, "UK_PPsero_mig.rds")
# new migration matrix that is closer to original one

# start population:
PPsero_startpop <- matrix(0, nrow = length(unique(UK_clusters$Cluster)), ncol = length(unique(UK_clusters$Serotype)))

for (i in which(UK_clusters$SeqYear=="1998")) {
  PPsero_startpop[PPclustToIndex_dict[UK_clusters$Cluster[i]],SerotypeToIndex_dict[UK_clusters$Serotype[i]]] <- PPsero_startpop[PPclustToIndex_dict[UK_clusters$Cluster[i]],SerotypeToIndex_dict[UK_clusters$Serotype[i]]] + 1
}

PP_expand_factor <- 15000 / sum(PPsero_startpop)
exp_noise <- 10
#PPsero_startpop_corr <- matrix(sapply((PPsero_startpop + matrix(rexp(n = ncol(PPsero_startpop) * nrow(PPsero_startpop), rate = exp_noise), nrow= nrow(PPsero_startpop), ncol = ncol(PPsero_startpop))) * PP_expand_factor, rpois, n=1), nrow= nrow(PPsero_startpop), ncol = ncol(PPsero_startpop))
#PPsero_startpop_corr <- matrix(sapply(PPsero_startpop_corr,as.double), nrow= nrow(PPsero_startpop), ncol = ncol(PPsero_startpop))
#saveRDS(PPsero_startpop_corr, "PPsero_startpop.rds")

# this lead to a start population that is way too evenly distributed across the different clusters
# instead, I could maybe use the existing PP_model_start_pop and distribute it to the different serotypes?
PPsero_startpop2 <-  matrix(0, nrow = length(unique(UK_clusters$Cluster)), ncol = length(unique(UK_clusters$Serotype)))
for (i in 1:length(unique(UK_clusters$Cluster))) {
  PPsero_startpop2[i,] <- round((PPxSero_matrix[i,] * UK_model_start_pop[i])/sum(PPxSero_matrix[i,]))
}
saveRDS(PPsero_startpop2, "UK_PPsero_startpop.rds")
# this is very close to the original PP_model_start_pop
# maybe I am giving out too much information because the serotype distribution is exactly the real one?

plot(rowSums(UK_cluster_freqs_sero[[1]])/sum(rowSums(UK_cluster_freqs_sero[[1]])))
points(rowSums(PPsero_startpop2)/sum(rowSums(PPsero_startpop2)), col = "red")

plot(colSums(UK_cluster_freqs_sero[[1]])/sum(colSums(UK_cluster_freqs_sero[[1]])))
points(colSums(PPsero_startpop2)/sum(colSums(PPsero_startpop2)), col = "red")
# the serotype distribution is quite far off

# same as PPsero_startpop6 for Mass
PPsero_startpop4 <- matrix(0, nrow = length(unique(UK_clusters$Cluster)), ncol = length(unique(UK_clusters$Serotype)))
PP_expand_factor <- 15000 / sum(UK_cluster_freqs_sero[[1]])
exp_noise <- 1000
for (i in 1:nrow(PPsero_startpop4)) {
    PPsero_startpop4[i,] <- (sapply((UK_cluster_freqs_sero[[1]][i,] + rexp(n = length(UK_cluster_freqs_sero[[1]][i,]), rate = exp_noise)) * PP_expand_factor, rpois, n=1)) + ceiling(PPxSero_matrix_prob2[i,])
  #PPsero_startpop5[i,] <- PP_mass_cluster_freq_1_sero[i,] + ceiling(PPxSero_matrix_prob[i,])
}
saveRDS(PPsero_startpop4, "UK_PPsero_startpop.rds")
# see next bit, I had already implemented that

plot(rowSums(UK_cluster_freqs_sero[[1]])/sum(rowSums(UK_cluster_freqs_sero[[1]])))
points(rowSums(PPsero_startpop4)/sum(rowSums(PPsero_startpop4)), col = "red")

plot(colSums(UK_cluster_freqs_sero[[1]])/sum(colSums(UK_cluster_freqs_sero[[1]])))
points(colSums(PPsero_startpop4)/sum(colSums(PPsero_startpop4)), col = "red")
# the serotype distribution is much closter now

# serotype VT vector
SeroVT <- rep(0, length(unique(UK_clusters$Serotype)))
for (i in 1:nrow(UK_clusters)) {
  SeroVT[SerotypeToIndex_dict[UK_clusters$Serotype[i]]] <- as.integer(UK_clusters$VT_PCV7[i]=="VT")
}
saveRDS(SeroVT, "UK_SeroVT.rds")

# serotype VT vector PCV13
SeroVT_PCV13 <- rep(0, length(unique(UK_clusters$Serotype)))
for (i in 1:nrow(UK_clusters)) {
  SeroVT_PCV13[SerotypeToIndex_dict[UK_clusters$Serotype[i]]] <- as.integer(UK_clusters$VT_PCV13[i]=="VT")
}
saveRDS(SeroVT_PCV13, "UK_SeroVT_PCV13.rds")
```

```{r}
# Create PP clusters, split by serotype
#calculate the frequency of the gene clusters and year and serotype
UK_mass_cluster_freq_1_sero <- matrix(0,nrow = UK_clusters_no, ncol = length(unique(UK_clusters$Serotype)))

for (i in 1:UK_clusters_no){
  for (k in unique(UK_clusters$Serotype)){
  UK_mass_cluster_freq_1_sero[i,which(unique(UK_clusters$Serotype)==k)] <- length(which(UK_clusters[which(UK_clusters$Cluster==unique(UK_clusters$Cluster)[i]),][UK_clusters[which(UK_clusters$Cluster==unique(UK_clusters$Cluster)[i]),]$Serotype ==k, ]$"Time"==sort(unique(UK_clusters$Time))[1]))
  }
}

#save sequence cluster frequencies
saveRDS(UK_mass_cluster_freq_1_sero, file = "UK_mass_cluster_freq_1_sero.rds")

PPxSero_matrix_prob2 <- readRDS("UK_PPsero_mig.rds")

UK_PPsero_startpop3 <- matrix(0,nrow = UK_clusters_no, ncol = length(unique(UK_clusters$Serotype)))
PP_expand_factor <- 15000 / sum(UK_mass_cluster_freq_1_sero)
exp_noise <- 1000
for (i in 1:nrow(UK_PPsero_startpop3)) {
    UK_PPsero_startpop3[i,] <- (sapply((UK_mass_cluster_freq_1_sero[i,] + rexp(n = length(UK_mass_cluster_freq_1_sero[i,]), rate = exp_noise)) * PP_expand_factor, rpois, n=1)) + ceiling(PPxSero_matrix_prob2[i,])
}
plot(colSums(UK_mass_cluster_freq_1_sero)/sum(colSums(UK_mass_cluster_freq_1_sero)))
points(colSums(UK_PPsero_startpop3)/sum(colSums(UK_PPsero_startpop3)), col = "red")

plot(rowSums(UK_mass_cluster_freq_1_sero)/sum((UK_mass_cluster_freq_1_sero)))
points(rowSums(UK_PPsero_startpop3)/sum(rowSums(UK_PPsero_startpop3)), col = "red")

saveRDS(UK_PPsero_startpop3, file = "UK_PPsero_startpop3.rds")
```

# US

```{r}
#US_time_points <- c("2001", "2003", "2007")
#US_ggCaller_byYear <- vector(mode = "list", length = length(US_time_points))

#for (i in 1:length(US_time_points)) {
#  US_ggCaller_year <- data.frame(matrix(0, nrow = nrow(US_ggCallerPP_bool), ncol = length(which(Mass_Samples_accCodes$`Year of Isolation`==US_time_points[i]))+1)) # initialize data frame
#  US_ggCaller_year[1,-1] <- US_ggCallerPP_bool[1,c(FALSE,SeqYear_dict_Taxon[unlist(US_ggCallerPP_bool[1,-1])]==US_time_points[i])] # fill in first row (seq names)
#  US_ggCaller_year[-1,1] <- US_ggCallerPP_bool[-1,1] # fill in first column (gene cluster names)
#  US_ggCaller_year[-1,-1] <- US_ggCallerPP_bool[-1,c(FALSE,SeqYear_dict_Taxon[unlist(US_ggCallerPP_bool[1,-1])]==US_time_points[i])] # fill in presence absence information
  #print(i)
#  US_ggCaller_byYear[[i]] <- US_ggCaller_year
#}
```

```{r}
# create time-point specific gene presence absence matrices
# SeqYear_dict has sequence years as values and isolate names as keys
# I need to first translate between ggCaller names (from file names) to isolate names via Accesscion number, using filenameToAccNo_dict and AccNoToIsolate_dict

# empty gene presence absence matrix
ggCaller_gene_presence_absence_2001 <- data.frame(matrix(0, nrow = nrow(US_ggCallerPP_bool), ncol = length(Mass_Samples_accCodes[Mass_Samples_accCodes$`Year of Isolation`==2001,]$`Isolate Name`)+1))
ggCaller_gene_presence_absence_2001[1,-1] <- US_ggCallerPP_bool[1,c(FALSE,SeqYear_dict[AccNoToIsolate_dict[filenameToAccNo_dict[unlist(US_ggCallerPP_bool[1,-1])]]]==2001)]
ggCaller_gene_presence_absence_2001[-1,1] <- US_ggCallerPP_bool[-1,1]
ggCaller_gene_presence_absence_2001[-1,-1] <- US_ggCallerPP_bool[-1,c(FALSE,SeqYear_dict[AccNoToIsolate_dict[filenameToAccNo_dict[unlist(US_ggCallerPP_bool[1,-1])]]]==2001)]

ggCaller_gene_presence_absence_2004 <- data.frame(matrix(0, nrow = nrow(US_ggCallerPP_bool), ncol = length(Mass_Samples_accCodes[Mass_Samples_accCodes$`Year of Isolation`==2004,]$`Isolate Name`)+1))
ggCaller_gene_presence_absence_2004[1,-1] <- US_ggCallerPP_bool[1,c(FALSE,SeqYear_dict[AccNoToIsolate_dict[filenameToAccNo_dict[unlist(US_ggCallerPP_bool[1,-1])]]]==2004)]
ggCaller_gene_presence_absence_2004[-1,1] <- US_ggCallerPP_bool[-1,1]
ggCaller_gene_presence_absence_2004[-1,-1] <- US_ggCallerPP_bool[-1,c(FALSE,SeqYear_dict[AccNoToIsolate_dict[filenameToAccNo_dict[unlist(US_ggCallerPP_bool[1,-1])]]]==2004)]

ggCaller_gene_presence_absence_2007 <- data.frame(matrix(0, nrow = nrow(US_ggCallerPP_bool), ncol = length(Mass_Samples_accCodes[Mass_Samples_accCodes$`Year of Isolation`==2007,]$`Isolate Name`)+1))
ggCaller_gene_presence_absence_2007[1,-1] <- US_ggCallerPP_bool[1,c(FALSE,SeqYear_dict[AccNoToIsolate_dict[filenameToAccNo_dict[unlist(US_ggCallerPP_bool[1,-1])]]]==2007)]
ggCaller_gene_presence_absence_2007[-1,1] <- US_ggCallerPP_bool[-1,1]
ggCaller_gene_presence_absence_2007[-1,-1] <- US_ggCallerPP_bool[-1,c(FALSE,SeqYear_dict[AccNoToIsolate_dict[filenameToAccNo_dict[unlist(US_ggCallerPP_bool[1,-1])]]]==2007)]
```

```{r}
#saveRDS(ggCaller_gene_presence_absence_2001, file = "ggCaller_gene_presence_absence_2001.rds")
#saveRDS(ggCaller_gene_presence_absence_2004, file = "ggCaller_gene_presence_absence_2004.rds")
#saveRDS(ggCaller_gene_presence_absence_2007, file = "ggCaller_gene_presence_absence_2007.rds")
```

```{r}
# Find intermediate frequency genes in 2001 data set
# compute gene frequencies
sum_as_int <- function(x){
  sum(as.integer(x))
}

ggC_gene_freq_2001 <- rep(0, nrow(ggCaller_gene_presence_absence_2001)-1)
ggC_gene_freq_2001 <- apply(ggCaller_gene_presence_absence_2001[-1,-1],1, sum_as_int)
ggC_gene_freq_2001 <- ggC_gene_freq_2001 / (length(ggCaller_gene_presence_absence_2001[1,-1]))

ggC_filter <- as.integer(ggC_gene_freq_2001<=0.95 & ggC_gene_freq_2001>=0.05)

ggC_intermed_gene_names <- ggCaller_gene_presence_absence_2001[which(ggC_filter==1),1]
saveRDS(ggC_intermed_gene_names, file = "Mass_ggC_intermed_gene_names.rds")

ggC_intermed_gene_freqs <- ggC_gene_freq_2001[which(ggC_filter==1)]
saveRDS(ggC_intermed_gene_freqs, file = "Mass_ggC_intermed_gene_freqs.rds")

ggC_all_gene_freqs <- ggC_gene_freq_2001
saveRDS(ggC_all_gene_freqs, file = "Mass_ggC_all_gene_freqs.rds")

ggC_all_gene_names <- ggCaller_gene_presence_absence_2001[-1,1]
saveRDS(ggC_all_gene_names, file = "Mass_ggC_all_gene_names.rds")

# intermediate gene presence absence matrices filtered by genes that are present at 5-95% frequency in 2001
ggC_intermed_gene_presence_absence_2001 <- data.frame(matrix(0, nrow = sum(ggC_filter)+1, ncol = length(Mass_Samples_accCodes[Mass_Samples_accCodes$`Year of Isolation`==2001,]$`Isolate Name`)+1))
ggC_intermed_gene_presence_absence_2001 <- ggCaller_gene_presence_absence_2001[c(1,which(ggC_filter==1)+1),]

ggC_intermed_gene_presence_absence_2004 <- data.frame(matrix(0, nrow = sum(ggC_filter)+1, ncol = length(Mass_Samples_accCodes[Mass_Samples_accCodes$`Year of Isolation`==2004,]$`Isolate Name`)+1))
ggC_intermed_gene_presence_absence_2004 <- ggCaller_gene_presence_absence_2004[c(1,which(ggC_filter==1)+1),]

ggC_intermed_gene_presence_absence_2007 <- data.frame(matrix(0, nrow = sum(ggC_filter)+1, ncol = length(Mass_Samples_accCodes[Mass_Samples_accCodes$`Year of Isolation`==2007,]$`Isolate Name`)+1))
ggC_intermed_gene_presence_absence_2007 <- ggCaller_gene_presence_absence_2007[c(1,which(ggC_filter==1)+1),]

ggC_intermed_gene_presence_absence <- data.frame(matrix(0, nrow = sum(ggC_filter)+1, ncol = length(Mass_Samples_accCodes$`Isolate Name`)+1))
ggC_intermed_gene_presence_absence <- US_ggCallerPP_bool[c(1,which(ggC_filter==1)+1),]
```

```{r}
#Compute consensus genomes for sequence clusters
ggCPP_Mass_seq_clusters_dict <- PopPUNK_clusters$Cluster
names(ggCPP_Mass_seq_clusters_dict) <- PopPUNK_clusters$Taxon

ggCPP_intermed_gene_presence_absence_consensus <- data.frame(matrix(0, nrow = sum(ggC_filter)+1, ncol = length(unique(PopPUNK_clusters$Cluster))+1))
ggCPP_intermed_gene_presence_absence_consensus[1,-1] <- paste("SeqCl_",unique(PopPUNK_clusters$Cluster),sep = "")

ggCPP_intermed_gene_presence_absence_consensus_2001 <- data.frame(matrix(0, nrow = sum(ggC_filter)+1, ncol = length(unique(PopPUNK_clusters$Cluster))+1))
ggCPP_intermed_gene_presence_absence_consensus_2001[1,-1] <- paste("SeqCl_",unique(PopPUNK_clusters$Cluster),sep = "")

ggCPP_intermed_gene_presence_absence_consensus_2004 <- data.frame(matrix(0, nrow = sum(ggC_filter)+1, ncol = length(unique(PopPUNK_clusters$Cluster))+1))
ggCPP_intermed_gene_presence_absence_consensus_2004[1,-1] <- paste("SeqCl_",unique(PopPUNK_clusters$Cluster),sep = "")

ggCPP_intermed_gene_presence_absence_consensus_2007 <- data.frame(matrix(0, nrow = sum(ggC_filter)+1, ncol = length(unique(PopPUNK_clusters$Cluster))+1))
ggCPP_intermed_gene_presence_absence_consensus_2007[1,-1] <- paste("SeqCl_",unique(PopPUNK_clusters$Cluster),sep = "")


cons_genomes <- function(x){
  as.double(ceiling(median(as.integer(x))))
}

for (i in unique(PopPUNK_clusters$Cluster)) {
  ggCPP_intermed_gene_presence_absence_consensus[-1,(which(unique(PopPUNK_clusters$Cluster) == i))+1] <- apply(as.matrix(ggC_intermed_gene_presence_absence[-1,c(FALSE,ggCPP_Mass_seq_clusters_dict[unlist(ggC_intermed_gene_presence_absence[1,-1])]==i)]), 1, cons_genomes)
  ggCPP_intermed_gene_presence_absence_consensus_2001[-1,(which(unique(PopPUNK_clusters$Cluster) == i))+1] <- apply(as.matrix(ggC_intermed_gene_presence_absence_2001[-1,c(FALSE,ggCPP_Mass_seq_clusters_dict[unlist(ggC_intermed_gene_presence_absence_2001[1,-1])]==i)]), 1, cons_genomes)
    ggCPP_intermed_gene_presence_absence_consensus_2004[-1,(which(unique(PopPUNK_clusters$Cluster) == i))+1] <- apply(as.matrix(ggC_intermed_gene_presence_absence_2004[-1,c(FALSE,ggCPP_Mass_seq_clusters_dict[unlist(ggC_intermed_gene_presence_absence_2004[1,-1])]==i)]), 1, cons_genomes)
    ggCPP_intermed_gene_presence_absence_consensus_2007[-1,(which(unique(PopPUNK_clusters$Cluster) == i))+1] <- apply(as.matrix(ggC_intermed_gene_presence_absence_2007[-1,c(FALSE,ggCPP_Mass_seq_clusters_dict[unlist(ggC_intermed_gene_presence_absence_2007[1,-1])]==i)]), 1, cons_genomes)
}

ggCPP_intermed_gene_presence_absence_consensus[-1,1] <- ggC_intermed_gene_presence_absence[-1,1]
ggCPP_intermed_gene_presence_absence_consensus_2001[-1,1] <- ggC_intermed_gene_presence_absence[-1,1]
ggCPP_intermed_gene_presence_absence_consensus_2004[-1,1] <- ggC_intermed_gene_presence_absence[-1,1]
ggCPP_intermed_gene_presence_absence_consensus_2007[-1,1] <- ggC_intermed_gene_presence_absence[-1,1]
```


```{r}
#save intermed gene presence absence matrices
saveRDS(ggCPP_intermed_gene_presence_absence_consensus, file = "ggCPP_intermed_gene_presence_absence_consensus.rds")
#saveRDS(ggCPP_intermed_gene_presence_absence_consensus_2001, file = "ggCPP_intermed_gene_presence_absence_consensus_2001.rds")
#saveRDS(ggCPP_intermed_gene_presence_absence_consensus_2004, file = "ggCPP_intermed_gene_presence_absence_consensus_2004.rds")
#saveRDS(ggCPP_intermed_gene_presence_absence_consensus_2007, file = "ggCPP_intermed_gene_presence_absence_consensus_2007.rds")
```

```{r}
# For strong and weak selection in the model, I need to calculate the delta statistics.
#calculate the frequency of the sequence clusters and year
mass_clusters <- length(unique(Mass_Samples_accCodes$SequenceCluster))
mass_cluster_freq_1 <- rep(0,mass_clusters)
mass_cluster_freq_2 <- rep(0,mass_clusters)
mass_cluster_freq_3 <- rep(0,mass_clusters)
for (i in 1:mass_clusters){
  mass_cluster_freq_1[i] <- length(which(Mass_Samples_accCodes[which(Mass_Samples_accCodes$SequenceCluster==i-1),]$"Year of Isolation"==2001))
  mass_cluster_freq_2[i] <- length(which(Mass_Samples_accCodes[which(Mass_Samples_accCodes$SequenceCluster==i-1),]$"Year of Isolation"==2004))
  mass_cluster_freq_3[i] <- length(which(Mass_Samples_accCodes[which(Mass_Samples_accCodes$SequenceCluster==i-1),]$"Year of Isolation"==2007))
}


# calculate gene frequencies first, separate for three time points
ggC_mass_gene_freq_1 <- apply(ggC_intermed_gene_presence_absence_2001[-1,-1], 1, sum_as_int)
ggC_mass_gene_freq_2 <- apply(ggC_intermed_gene_presence_absence_2004[-1,-1], 1, sum_as_int)
ggC_mass_gene_freq_3 <- apply(ggC_intermed_gene_presence_absence_2007[-1,-1], 1, sum_as_int)

# first, calculate pre/peri and post vacc frequencies of genes:
ggC_pre_peri_vacc_gene_freq <- (ggC_mass_gene_freq_1 + ggC_mass_gene_freq_2) / (sum(mass_cluster_freq_1) + sum(mass_cluster_freq_2))
ggC_pre_vacc_gene_freq <- (ggC_mass_gene_freq_1) / sum(mass_cluster_freq_1)
ggC_post_vacc_gene_freq <- ggC_mass_gene_freq_3 / sum(mass_cluster_freq_3)
ggC_peri_post <- (ggC_mass_gene_freq_2 + ggC_mass_gene_freq_3) / (sum(mass_cluster_freq_2) + sum(mass_cluster_freq_3))
ggC_peri_vacc_gene_freq <- (ggC_mass_gene_freq_2) / sum(mass_cluster_freq_2)

# calculate delta statistic (refer to Corander et al. for more info)

ggC_delta_data2 <- (ggC_post_vacc_gene_freq - ggC_pre_vacc_gene_freq) ^ 2 / (1 - ggC_pre_vacc_gene_freq * (1 - ggC_pre_vacc_gene_freq))
ggC_delta_ranking2 <- rank(ggC_delta_data2)

# rownames to gene names
ggCaller_us_rowname_genename_dict <- ggC_intermed_gene_presence_absence[-1,1]
names(ggCaller_us_rowname_genename_dict) <- rownames(ggC_intermed_gene_presence_absence)[-1]
saveRDS(ggCaller_us_rowname_genename_dict, "ggCaller_us_rowname_genename_dict.rds")
```

```{r}
#save delta ranking
saveRDS(ggC_delta_ranking2, file = "ggC_delta_ranking.rds")
saveRDS(ggC_delta_data2, file = "ggC_delta_data2.rds")
```

```{r}
#calculate the frequency of the gene clusters and year
PP_mass_cluster_freq_1 <- rep(0,no_PopPUNK_clusters)
PP_mass_cluster_freq_2 <- rep(0,no_PopPUNK_clusters)
PP_mass_cluster_freq_3 <- rep(0,no_PopPUNK_clusters)
for (i in 1:no_PopPUNK_clusters){
  PP_mass_cluster_freq_1[i] <- length(which(PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),]$"SeqYear"==2001))
  PP_mass_cluster_freq_2[i] <- length(which(PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),]$"SeqYear"==2004))
  PP_mass_cluster_freq_3[i] <- length(which(PopPUNK_clusters[which(PopPUNK_clusters$Cluster==unique(PopPUNK_clusters$Cluster)[i]),]$"SeqYear"==2007))
}
```

```{r}
#save sequence cluster frequencies
saveRDS(PP_mass_cluster_freq_1, file = "PP_mass_cluster_freq_1.rds")
saveRDS(PP_mass_cluster_freq_2, file = "PP_mass_cluster_freq_2.rds")
saveRDS(PP_mass_cluster_freq_3, file = "PP_mass_cluster_freq_3.rds")
```

# calculate delta across countries instead of time
```{r}
# find overlapping intermediate-freq genes btw UK and US
US_UK_intermed_genes <- intersect(names(ggC_mass_gene_freq_1), names(UK_gene_freq_preVac)) # 1535 genes overlap

# first, calculate pre/peri and post vacc frequencies of genes:
ggC_pre_vacc_gene_freq <- (ggC_mass_gene_freq_1) / sum(mass_cluster_freq_1)
ggC_post_vacc_gene_freq <- ggC_mass_gene_freq_3 / sum(mass_cluster_freq_3)

# calculate delta statistic (refer to Corander et al. for more info)
ggC_delta_data2_sharedUSUK <- (ggC_post_vacc_gene_freq[US_UK_intermed_genes] - ggC_pre_vacc_gene_freq[US_UK_intermed_genes]) ^ 2 / (1 - ggC_pre_vacc_gene_freq[US_UK_intermed_genes] * (1 - ggC_pre_vacc_gene_freq[US_UK_intermed_genes]))
ggC_delta_ranking2_sharedUSUK <- rank(ggC_delta_data2_sharedUSUK)

UK_gene_freq_preVac_rel <- UK_gene_freq_preVac/(ncol(UK_ggCaller_intermed_preVac) -1)
UK_gene_freq_postVac_rel <- UK_gene_freq_postVac/((ncol(UK_ggCaller_byYear_intermed[[length(UK_ggCaller_byYear_intermed)]]) -1) + (ncol(UK_ggCaller_byYear_intermed[[length(UK_ggCaller_byYear_intermed)-1]]) -1) + (ncol(UK_ggCaller_byYear_intermed[[length(UK_ggCaller_byYear_intermed)-2]]) -1))

# calculate delta statistic (refer to Corander et al. for more info)
UK_delta_data_sharedUSUK <- (UK_gene_freq_postVac_rel[US_UK_intermed_genes] - UK_gene_freq_preVac_rel[US_UK_intermed_genes]) ^ 2 / (1 - UK_gene_freq_preVac_rel[US_UK_intermed_genes] * (1 - UK_gene_freq_preVac_rel[US_UK_intermed_genes]))
UK_delta_ranking_sharedUSUK <- rank(UK_delta_data_sharedUSUK)

# save this for US and UK (each) and test fitting
# US for US (only those that are shared)
ggC_delta_ranking2_sharedUSUK_forUS <- rep(max(ggC_delta_ranking2), length(ggC_delta_ranking2))
names(ggC_delta_ranking2_sharedUSUK_forUS) <- names(ggC_delta_ranking2)
ggC_delta_ranking2_sharedUSUK_forUS[US_UK_intermed_genes] <- ggC_delta_ranking2_sharedUSUK[US_UK_intermed_genes]

# UK for US
UK_delta_ranking_sharedUSUK_forUS <- rep(max(ggC_delta_ranking2), length(ggC_delta_ranking2))
names(UK_delta_ranking_sharedUSUK_forUS) <- names(ggC_delta_ranking2)
UK_delta_ranking_sharedUSUK_forUS[US_UK_intermed_genes] <- UK_delta_ranking_sharedUSUK[US_UK_intermed_genes]

# US for UK
ggC_delta_ranking2_sharedUSUK_forUK <- rep(max(UK_delta_ranking),length(UK_delta_ranking))
names(ggC_delta_ranking2_sharedUSUK_forUK) <- names(UK_delta_ranking)
ggC_delta_ranking2_sharedUSUK_forUK[US_UK_intermed_genes] <- ggC_delta_ranking2_sharedUSUK[US_UK_intermed_genes]

# UK for UK (filter by shared intermed genes)
UK_delta_ranking_sharedUSUK_forUK <- rep(max(UK_delta_ranking),length(UK_delta_ranking))
names(UK_delta_ranking_sharedUSUK_forUK) <- names(UK_delta_ranking)
UK_delta_ranking_sharedUSUK_forUK[US_UK_intermed_genes] <- UK_delta_ranking_sharedUSUK[US_UK_intermed_genes]

#saveRDS(ggC_delta_ranking2_sharedUSUK, "ggC_delta_ranking_sharedUSUK.rds")
#saveRDS(UK_delta_ranking_sharedUSUK, "UK_delta_ranking_sharedUSUK.rds")

saveRDS(ggC_delta_ranking2_sharedUSUK_forUS, "ggC_delta_ranking2_sharedUSUK_forUS.rds")
saveRDS(UK_delta_ranking_sharedUSUK_forUS, "UK_delta_ranking_sharedUSUK_forUS.rds")
saveRDS(ggC_delta_ranking2_sharedUSUK_forUK, "ggC_delta_ranking2_sharedUSUK_forUK.rds")
saveRDS(UK_delta_ranking_sharedUSUK_forUK, "UK_delta_ranking_sharedUSUK_forUK.rds")

US_UK_delta_data <- (UK_gene_freq_preVac_rel[US_UK_intermed_genes] - ggC_pre_vacc_gene_freq[US_UK_intermed_genes]) ^ 2 / (1 - ggC_pre_vacc_gene_freq[US_UK_intermed_genes] * (1 - ggC_pre_vacc_gene_freq[US_UK_intermed_genes]))
US_UK_delta_ranking <- rank(US_UK_delta_data)

UK_US_delta_data <- (ggC_pre_vacc_gene_freq[US_UK_intermed_genes] - UK_gene_freq_preVac_rel[US_UK_intermed_genes]) ^ 2 / (1 - UK_gene_freq_preVac_rel[US_UK_intermed_genes] * (1 - UK_gene_freq_preVac_rel[US_UK_intermed_genes]))
UK_US_delta_ranking <- rank(UK_US_delta_data)


plot(US_UK_delta_ranking, UK_US_delta_ranking) # very strong correlation, i.e. does not matter whether I define US as the "post" or UK

# but no correlation with the original delta-rankings...
plot(US_UK_delta_ranking[US_UK_intermed_genes], ggC_delta_ranking2[US_UK_intermed_genes])
plot(US_UK_delta_ranking[US_UK_intermed_genes], UK_delta_ranking[US_UK_intermed_genes])

plot(ggC_delta_ranking2[US_UK_intermed_genes], UK_delta_ranking[US_UK_intermed_genes])
plot(ggC_delta_ranking2_sharedUSUK, UK_delta_ranking_sharedUSUK)

# create ranking for US dataset, based on US-UK ranking, set rest to max
ggC_delta_ranking_USUK <- rep(max(ggC_delta_ranking2), length(ggC_delta_ranking2))
names(ggC_delta_ranking_USUK) <- names(ggC_delta_ranking2)
ggC_delta_ranking_USUK[US_UK_intermed_genes] <- UK_US_delta_ranking[US_UK_intermed_genes]

plot(US_UK_delta_ranking[US_UK_intermed_genes], ggC_delta_ranking_USUK[US_UK_intermed_genes])

# same for UK
UK_delta_ranking_USUK <- rep(max(UK_delta_ranking),length(UK_delta_ranking))
names(UK_delta_ranking_USUK) <- names(UK_delta_ranking)
UK_delta_ranking_USUK[US_UK_intermed_genes] <- UK_US_delta_ranking[US_UK_intermed_genes]

plot(US_UK_delta_ranking[US_UK_intermed_genes], UK_delta_ranking_USUK[US_UK_intermed_genes])

plot(ggC_pre_vacc_gene_freq[US_UK_intermed_genes], UK_gene_freq_preVac_rel[US_UK_intermed_genes], ylim = c(0,1), xlim = c(0,1))
plot(ggC_post_vacc_gene_freq[US_UK_intermed_genes], UK_gene_freq_postVac_rel[US_UK_intermed_genes], ylim = c(0,1), xlim = c(0,1))
# checking the correlation btw gene freqs btw US & UK, and pre-/post-vacc
cor(ggC_pre_vacc_gene_freq[US_UK_intermed_genes], UK_gene_freq_preVac_rel[US_UK_intermed_genes])
# 0.8269251
cor(ggC_post_vacc_gene_freq[US_UK_intermed_genes], UK_gene_freq_postVac_rel[US_UK_intermed_genes])
# 0.8700141
cor(ggC_pre_vacc_gene_freq[US_UK_intermed_genes],ggC_post_vacc_gene_freq[US_UK_intermed_genes])
# 0.963727
cor(UK_gene_freq_preVac_rel[US_UK_intermed_genes], UK_gene_freq_postVac_rel[US_UK_intermed_genes])
# 0.8814543

plot(ggC_pre_vacc_gene_freq[US_UK_intermed_genes], UK_gene_freq_preVac_rel[US_UK_intermed_genes], ylim = c(0,1), xlim = c(0,1))
#points()


# investigate gene correlation
joined_gene_freq_list <- vector(mode = "list", length = 3) # ggCaller gene freq list for UK, US, Nepal
names(joined_gene_freq_list) <- c("UK", "US", "Nepal")

sum_as_int <- function(x){
  sum(as.integer(x))
}

for (country in names(joined_ggCaller_list)) {
  geneSum <- apply(joined_ggCaller_list_UKUSNepal[[country]][-1,-1],1, sum_as_int)
  names(geneSum) <- joined_ggCaller_list_UKUSNepal[[country]][-1,1]
  joined_gene_freq_list[[country]] <- geneSum / (ncol(joined_ggCaller_list_UKUSNepal[[country]]) - 1)
}
plot(joined_gene_freq_list[["UK"]], joined_gene_freq_list[["US"]])
plot(joined_gene_freq_list[["UK"]][paste("group_",US_UK_intermed_genes, sep = "")], joined_gene_freq_list[["US"]][paste("group_",US_UK_intermed_genes, sep = "")]) # these are the genes that are at intermed-freq in the beginning...?!

saveRDS(US_UK_delta_ranking,"US_UK_delta_ranking.rds")
saveRDS(ggC_delta_ranking_USUK, "ggC_delta_ranking_USUK.rds")
saveRDS(UK_delta_ranking_USUK, "UK_delta_ranking_USUK.rds")


# maybe n~=100 is just not enough?
# use whole US and UK dataset?
US_intermed_gene_freq <- (ggC_mass_gene_freq_1 + ggC_mass_gene_freq_2 + ggC_mass_gene_freq_3) / (sum(mass_cluster_freq_1) + sum(mass_cluster_freq_2) + sum(mass_cluster_freq_3))
UK_intermed_gene_freq <- apply(UK_ggCaller_intermed[-1,-1], 1, sum_as_int)/(ncol(UK_ggCaller_intermed)-1)

US_UK_delta_data_all_time <- (UK_intermed_gene_freq[US_UK_intermed_genes] - US_intermed_gene_freq[US_UK_intermed_genes]) ^ 2 / (1 - US_intermed_gene_freq[US_UK_intermed_genes] * (1 - US_intermed_gene_freq[US_UK_intermed_genes]))
US_UK_delta_ranking_all_time <- rank(US_UK_delta_data_all_time)

# but no correlation with the original delta-rankings... (still not!)
plot(US_UK_delta_ranking_all_time[US_UK_intermed_genes], ggC_delta_ranking2[US_UK_intermed_genes])
plot(US_UK_delta_ranking_all_time[US_UK_intermed_genes], UK_delta_ranking[US_UK_intermed_genes])
```

