---
title: "Model comparison 2, 3, 4, 5 parameters"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

```{r}
library(bayestestR)
library(coda)
### Mass

# Null model
Mass_2param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2024_12_05/ModelComparison_Mass/PPxSero_ggCaller_PopPUNK_Null_det_pmcmc_run2.rds")
Mass_2param_mcmc2_probs <- coda::as.mcmc(cbind(Mass_2param_mcmc2$probabilities, Mass_2param_mcmc2$pars))
#coda::effectiveSize(Mass_2param_mcmc2_probs)
#summary(coda::as.mcmc(Mass_2param_mcmc2_probs))
Mass_likelihood_2param <- median(Mass_2param_mcmc2_probs[,"log_likelihood"])
#mean(Mass_2param_mcmc2_probs[,"log_likelihood"])
param_estimates <- map_estimate(Mass_2param_mcmc2$pars)
#-4.54, 0.04
param_mean <- apply(Mass_2param_mcmc2_probs, 2, mean)
percentile95_low <- apply(Mass_2param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(Mass_2param_mcmc2_probs,2,quantile,.975)
#     log_prior log_likelihood  log_posterior              m              v 
#    0.00000000  -283.03864760  -283.03864760    -4.55741935     0.03749215 
#    0.00000000  -285.77115406  -285.77115406    -4.78545282     0.02894042 
#    0.00000000  -282.05070637  -282.05070637    -4.33260021     0.04784521
rm(Mass_2param_mcmc2)
rm(Mass_2param_mcmc2_probs)

# 3-parameter model
Mass_3param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2024_12_05/ModelComparison_Mass/PPxSero_ggCaller_PopPUNK_3param_det_pmcmc_run2.rds")
Mass_3param_mcmc2_probs <- coda::as.mcmc(cbind(Mass_3param_mcmc2$probabilities, Mass_3param_mcmc2$pars))
#coda::effectiveSize(Mass_3param_mcmc2_probs)
#summary(coda::as.mcmc(Mass_3param_mcmc2_probs))
Mass_likelihood_3param <- median(Mass_3param_mcmc2_probs[,"log_likelihood"])
#mean(Mass_3param_mcmc2_probs[,"log_likelihood"])
param_estimates <- map_estimate(Mass_3param_mcmc2$pars)
# -6.49, -4.50, 0.05
param_mean <- apply(Mass_3param_mcmc2_probs, 2, mean)
percentile95_low <- apply(Mass_3param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(Mass_3param_mcmc2_probs,2,quantile,.975)
#     log_prior log_likelihood  log_posterior        sigma_f              m              v 
#    0.00000000  -278.16020683  -278.16020683    -6.74752111    -4.51538531     0.05239445
#    0.00000000  -282.04398459  -282.04398459    -8.28515193    -4.78809296     0.03730393
#    0.00000000  -276.63008552  -276.63008552    -5.92139441    -4.24192984     0.06905461
rm(Mass_3param_mcmc2)
rm(Mass_3param_mcmc2_probs)

# 4-parameter model with GPSCs
Mass_4param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2025_01_13/Mass_GPSCs/PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
Mass_4param_mcmc2_probs <- coda::as.mcmc(cbind(Mass_4param_mcmc2$probabilities, Mass_4param_mcmc2$pars))
#coda::effectiveSize(Mass_4param_mcmc2_probs)
#summary(coda::as.mcmc(Mass_4param_mcmc2_probs))
Mass_likelihood_4param <- median(Mass_4param_mcmc2_probs[,"log_likelihood"])
#mean(Mass_4param_mcmc2_probs[,"log_likelihood"])
param_estimates <- map_estimate(Mass_4param_mcmc2$pars)
# -3.27, 0.24, -4.36, 0.08
# GPSCs:
# -3.24, 0.24, -4.31, 0.08
param_mean <- apply(Mass_4param_mcmc2_probs, 2, mean)
percentile95_low <- apply(Mass_4param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(Mass_4param_mcmc2_probs,2,quantile,.975)
# (old values)
#     log_prior log_likelihood  log_posterior        sigma_f         prop_f              m              v 
#    0.00000000  -236.18187830  -236.18187830    -3.29342091     0.27971311    -4.31740603     0.08169713 
#    0.00000000  -240.56908318  -240.56908318    -4.18698747     0.11780847    -4.80025991     0.06278332
#     0.0000000   -233.6428808   -233.6428808     -2.1707973      0.4555952     -3.8273690      0.1059178
rm(Mass_4param_mcmc2)
rm(Mass_4param_mcmc2_probs)

# 5-parameter model
Mass_5param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2024_12_05/ModelComparison_Mass/PPxSero_ggCaller_PopPUNK_5param_det_pmcmc_run2.rds")
Mass_5param_mcmc2_probs <- coda::as.mcmc(cbind(Mass_5param_mcmc2$probabilities, Mass_5param_mcmc2$pars))
#coda::effectiveSize(Mass_5param_mcmc2_probs)
#summary(coda::as.mcmc(Mass_5param_mcmc2_probs))
Mass_likelihood_5param <- median(Mass_5param_mcmc2_probs[,"log_likelihood"])
#mean(Mass_5param_mcmc2_probs[,"log_likelihood"])
param_estimates <- map_estimate(Mass_5param_mcmc2$pars)
# -3.07, 0.25, -308.77, -4.27, 0.08
param_mean <- apply(Mass_5param_mcmc2_probs, 2, mean)
percentile95_low <- apply(Mass_5param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(Mass_5param_mcmc2_probs,2,quantile,.975)
#      log_prior log_likelihood  log_posterior        sigma_f         prop_f        sigma_w              m              v 
#     0.0000000   -235.8505446   -235.8505446     -3.1303171      0.2529448   -478.3519069     -4.2950099      0.0831402 
#     0.0000000   -239.8943341   -239.8943341     -3.4820120      0.1705453   -970.0844042     -4.7681127      0.0648898 
#     0.0000000   -233.6029196   -233.6029196     -2.5691558      0.3820193    -31.7122708     -3.7922820      0.1061980 
rm(Mass_5param_mcmc2)
rm(Mass_5param_mcmc2_probs)


# 11.08.2025

Mass_4param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_04/US_newPrior/PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
Mass_4param_mcmc2_probs <- coda::as.mcmc(cbind(Mass_4param_mcmc2$probabilities, Mass_4param_mcmc2$pars))
Mass_likelihood_4param <- mean(Mass_4param_mcmc2_probs[,"log_likelihood"])
Mass_param_mean4 <- apply(Mass_4param_mcmc2_probs, 2, mean)
Mass_percentile95_low <- apply(Mass_4param_mcmc2_probs,2,quantile,.025)
Mass_percentile95_up <- apply(Mass_4param_mcmc2_probs,2,quantile,.975)
rm(Mass_4param_mcmc2)

Mass_2param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_04/US_newPrior/PPxSero_ggCaller_PopPUNK_Null_det_pmcmc_run2.rds")
Mass_2param_mcmc2_probs <- coda::as.mcmc(cbind(Mass_2param_mcmc2$probabilities, Mass_2param_mcmc2$pars))
Mass_likelihood_2param <- mean(Mass_2param_mcmc2_probs[,"log_likelihood"])
Mass_param_mean2 <- apply(Mass_2param_mcmc2_probs, 2, mean)
Mass_percentile95_low_2param <- apply(Mass_2param_mcmc2_probs,2,quantile,.025)
Mass_percentile95_up_2param <- apply(Mass_2param_mcmc2_probs,2,quantile,.975)
rm(Mass_2param_mcmc2)

Mass_3param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_04/US_newPrior/PPxSero_ggCaller_PopPUNK_3param_det_pmcmc_run2.rds")
Mass_3param_mcmc2_probs <- coda::as.mcmc(cbind(Mass_3param_mcmc2$probabilities, Mass_3param_mcmc2$pars))
Mass_likelihood_3param <- mean(Mass_3param_mcmc2_probs[,"log_likelihood"])
Mass_param_mean3 <- apply(Mass_3param_mcmc2_probs, 2, mean)
Mass_percentile95_low_3param <- apply(Mass_3param_mcmc2_probs,2,quantile,.025)
Mass_percentile95_up_3param <- apply(Mass_3param_mcmc2_probs,2,quantile,.975)
rm(Mass_3param_mcmc2)

Mass_5param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_04/US_newPrior/PPxSero_ggCaller_PopPUNK_5param_det_pmcmc_run2.rds")
Mass_5param_mcmc2_probs <- coda::as.mcmc(cbind(Mass_5param_mcmc2$probabilities, Mass_5param_mcmc2$pars))
Mass_likelihood_5param <- mean(Mass_5param_mcmc2_probs[,"log_likelihood"])
Mass_param_mean5 <- apply(Mass_5param_mcmc2_probs, 2, mean)
Mass_percentile95_low_5param <- apply(Mass_5param_mcmc2_probs,2,quantile,.025)
Mass_percentile95_up_5param <- apply(Mass_5param_mcmc2_probs,2,quantile,.975)
rm(Mass_5param_mcmc2)


# mean fit 
Mass_4param_mcmc2_mean <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2026_03_23/US_mean/PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
Mass_4param_mcmc2_probs_mean <- coda::as.mcmc(cbind(Mass_4param_mcmc2_mean$probabilities, Mass_4param_mcmc2_mean$pars))
Mass_likelihood_4param_mean <- mean(Mass_4param_mcmc2_probs_mean[,"log_likelihood"])
Mass_param_mean4_mean <- apply(Mass_4param_mcmc2_probs_mean, 2, mean)
Mass_percentile95_low_mean <- apply(Mass_4param_mcmc2_probs_mean,2,quantile,.025)
Mass_percentile95_up_mean <- apply(Mass_4param_mcmc2_probs_mean,2,quantile,.975)
rm(Mass_4param_mcmc2_mean)

# large population fit
Mass_4param_mcmc2_largePop <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2026_03_25/US_LargePop/PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
Mass_4param_mcmc2_probs_largePop <- coda::as.mcmc(cbind(Mass_4param_mcmc2_largePop$probabilities, Mass_4param_mcmc2_largePop$pars))
Mass_likelihood_4param_largePop <- mean(Mass_4param_mcmc2_probs_largePop[,"log_likelihood"])
Mass_param_mean4_largePop <- apply(Mass_4param_mcmc2_probs_largePop, 2, mean)
Mass_percentile95_low_largePop <- apply(Mass_4param_mcmc2_probs_largePop,2,quantile,.025)
Mass_percentile95_up_largePop <- apply(Mass_4param_mcmc2_probs_largePop,2,quantile,.975)
rm(Mass_4param_mcmc2_largePop)
```


# plot mean vs standard together
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

plot_pmcmc_posteriors <- function(fit1, fit2, 
                                  label1 = "Model 1", label2 = "Model 2",
                                  burnin = 0, exp_params = c("m", "sigma_f")) {
  
  # Extract parameter samples as data frames, dropping the log-likelihood column
  extract_samples <- function(fit, label, burnin = 0, n_chains = 4) {
  samples <- as.data.frame(fit$pars)
  
  n_total <- nrow(samples)
  
  if (n_total %% n_chains != 0) warning("Total iterations not evenly divisible by n_chains — check n_chains is correct")
  
  n_iter <- n_total / n_chains
  
  if (burnin >= n_iter) stop("burnin is >= number of iterations per chain")
  
  # Build an index of rows to keep: drop first `burnin` rows from each chain's block
  keep <- unlist(lapply(seq_len(n_chains), function(i) {
    block_start <- (i - 1) * n_iter + 1
    block_end   <- i * n_iter
    seq(block_start + burnin, block_end)
  }))
  
  samples <- samples[keep, ]
  
  if (!is.null(exp_params)) {
    missing_params <- setdiff(exp_params, names(samples))
    if (length(missing_params) > 0) warning("exp_params not found in samples: ", paste(missing_params, collapse = ", "))
    for (p in intersect(exp_params, names(samples))) {
      samples[[p]] <- exp(samples[[p]])
    }
  }
  
  samples$model <- label
  samples
}
  
  df1 <- extract_samples(fit1, label1, burnin = burnin)
  df2 <- extract_samples(fit2, label2, burnin = burnin)
  
  # Only keep parameters present in both objects
  params1 <- setdiff(names(df1), "model")
  params2 <- setdiff(names(df2), "model")
  shared_params <- intersect(params1, params2)
  
  if (length(shared_params) == 0) stop("No shared parameters between the two objects.")
  
  if (!setequal(params1, params2)) {
    message("Note: parameters not shared between objects will be dropped: ",
            paste(union(setdiff(params1, params2), setdiff(params2, params1)), collapse = ", "))
  }
  
  # Combine and pivot to long format
  combined <- bind_rows(
    df1[, c(shared_params, "model")],
    df2[, c(shared_params, "model")]
  ) |>
    pivot_longer(cols = all_of(shared_params),
                 names_to = "parameter",
                 values_to = "value")
  
  ggplot(combined, aes(x = value, colour = model, fill = model)) +
    geom_density(alpha = 0.2, linewidth = 0.8) +
    facet_wrap(~ parameter, scales = "free") +
    scale_colour_brewer(palette = "Set1") +
    scale_fill_brewer(palette = "Set1") +
    labs(x = "Value", y = "Density", colour = NULL, fill = NULL) +
    theme_bw(base_size = 22) +  
    theme(legend.position = "top",
          axis.title  = element_text(size = 22),   
          axis.text   = element_text(size = 22),   
          strip.text  = element_text(size = 22),   
          legend.text = element_text(size = 22))  
}

# Usage
plot_pmcmc_posteriors(Mass_4param_mcmc2, Mass_4param_mcmc2_mean, label1 = "0/1", label2 = "Mean", burnin = 10000)


plot_pmcmc_posteriors(Mass_4param_mcmc2, Mass_4param_mcmc2_largePop, label1 = "15,000", label2 = "200,000", burnin = 10000)
```




# Likelihood ratio test
#  lambda = -2*(x_1-x_2).
# pvalue_testing_null=1-pchisq(lambda,difference_in_degrees_of_freedom) 
```{r}
# second is 11.08.
lambda <- -2 * (Mass_likelihood_2param - Mass_likelihood_3param)
pchisq(lambda, 1,lower.tail = FALSE)
# 0.001665563
# 6.580147e-06

lambda <- -2 * (Mass_likelihood_2param - Mass_likelihood_4param)
pchisq(lambda, 2,lower.tail = FALSE)
# 4.630899e-21
# 4.687568e-25

lambda <- -2 * (Mass_likelihood_3param - Mass_likelihood_4param)
pchisq(lambda, 1,lower.tail = FALSE)
# 5.594179e-20
# 9.945497e-22

lambda <- -2 * (Mass_likelihood_2param - Mass_likelihood_5param)
pchisq(lambda, 3,lower.tail = FALSE)
# 2.381506e-20
# 3.515104e-24

lambda <- -2 * (Mass_likelihood_3param - Mass_likelihood_5param)
pchisq(lambda, 2,lower.tail = FALSE)
# 4.25975e-19
# 1.060714e-20

lambda <- -2 * (Mass_likelihood_4param - Mass_likelihood_5param)
pchisq(lambda, 1,lower.tail = FALSE)
# 0.3586622
# 0.6117305

Mass_2_bic<- -2 * (Mass_likelihood_2param) + log(616) * 2 # 558.1187
Mass_3_bic<- -2 * (Mass_likelihood_3param) + log(616) * 3 # 544.2304
Mass_4_bic<- -2 * (Mass_likelihood_4param) + log(616) * 4 # 458.9257
Mass_5_bic<- -2 * (Mass_likelihood_5param) + log(616) * 5 # 465.0913

Mass_4_bic_mean <- -2 * (Mass_likelihood_4param_mean) + log(616) * 4 # 449.8755
```

```{r}
### UK

# Null model
UK_2param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2024_12_05/ModelComparison_UK/UK_PPxSero_ggCaller_PopPUNK_Null_det_pmcmc_run2.rds")
UK_2param_mcmc2_probs <- coda::as.mcmc(cbind(UK_2param_mcmc2$probabilities, UK_2param_mcmc2$pars))
#coda::effectiveSize(UK_2param_mcmc2_probs)
#summary(coda::as.mcmc(UK_2param_mcmc2_probs))
UK_likelihood_2param <- median(UK_2param_mcmc2_probs[,"log_likelihood"])
#mean(UK_2param_mcmc2_probs[,"log_likelihood"])
param_mean <- apply(UK_2param_mcmc2_probs, 2, mean)
percentile95_low <- apply(UK_2param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(UK_2param_mcmc2_probs,2,quantile,.975)
#     log_prior log_likelihood  log_posterior              m              v 
#    0.00000000  -641.16360797  -641.16360797    -4.36138821     0.02805271 
#    0.00000000  -643.95162152  -643.95162152    -4.56306240     0.02164408
#    0.00000000  -640.17013460  -640.17013460    -4.16166788     0.03482034
rm(UK_2param_mcmc2)

# 3-parameter model
UK_3param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2024_12_05/ModelComparison_UK/UK_PPxSero_ggCaller_PopPUNK_3param_det_pmcmc_run2.rds")
UK_3param_mcmc2_probs <- coda::as.mcmc(cbind(UK_3param_mcmc2$probabilities, UK_3param_mcmc2$pars))
#coda::effectiveSize(UK_3param_mcmc2_probs)
#summary(coda::as.mcmc(UK_3param_mcmc2_probs))
UK_likelihood_3param <- median(UK_3param_mcmc2_probs[,"log_likelihood"])
#mean(UK_3param_mcmc2_probs[,"log_likelihood"])
param_mean <- apply(UK_3param_mcmc2_probs, 2, mean)
percentile95_low <- apply(UK_3param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(UK_3param_mcmc2_probs,2,quantile,.975)
#     log_prior log_likelihood  log_posterior        sigma_f              m              v 
#    0.00000000  -643.07829722  -643.07829722   -40.87901034    -4.45696522     0.05151384 
#    0.00000000  -652.65879091  -652.65879091  -307.70593047    -4.75223182     0.02277319 
#     0.0000000   -640.2573559   -640.2573559     -4.2676913     -2.4375028      0.6825091
rm(UK_3param_mcmc2)

# 4-parameter model GPSC version
UK_4param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2025_01_13/UK_GPSCs/UK_PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
UK_4param_mcmc2_probs <- coda::as.mcmc(cbind(UK_4param_mcmc2$probabilities, UK_4param_mcmc2$pars))
#coda::effectiveSize(UK_4param_mcmc2_probs)
#summary(coda::as.mcmc(UK_4param_mcmc2_probs))
UK_likelihood_4param <- median(UK_4param_mcmc2_probs[,"log_likelihood"])
#mean(UK_4param_mcmc2_probs[,"log_likelihood"])
param_mean <- apply(UK_4param_mcmc2_probs, 2, mean)
percentile95_low <- apply(UK_4param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(UK_4param_mcmc2_probs,2,quantile,.975)
# old
#     log_prior log_likelihood  log_posterior        sigma_f         prop_f              m              v 
#     0.0000000   -562.2620742   -562.2620742     -3.1745767      0.1432364     -3.9757905      0.0910330
#    0.00000000  -566.34600127  -566.34600127    -3.62415671     0.10767921    -4.26552560     0.07626173
#     0.0000000   -560.4980507   -560.4980507     -2.5274892      0.1524845     -3.6791013      0.1069834 
rm(UK_4param_mcmc2)

#this one does not exist yet:
# 5-parameter model
UK_5param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2024_12_11/ModelComparison_UK/UK_PPxSero_ggCaller_PopPUNK_5param_det_pmcmc_run2.rds")
UK_5param_mcmc2_probs <- coda::as.mcmc(cbind(UK_5param_mcmc2$probabilities, UK_5param_mcmc2$pars))
#coda::effectiveSize(UK_5param_mcmc2_probs)
#summary(coda::as.mcmc(UK_5param_mcmc2_probs))
UK_likelihood_5param <- median(UK_5param_mcmc2_probs[,"log_likelihood"])
#mean(UK_5param_mcmc2_probs[,"log_likelihood"])
param_mean <- apply(UK_5param_mcmc2_probs, 2, mean)
percentile95_low <- apply(UK_5param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(UK_5param_mcmc2_probs,2,quantile,.975)
#     log_prior log_likelihood  log_posterior        sigma_f         prop_f        sigma_w              m              v 
#    0.00000000  -565.12227695  -565.12227695    -4.04693826     0.31370789  -484.23214193    -4.11417055     0.09340846
#    0.00000000  -569.20950094  -569.20950094    -4.51206733     0.11112096  -971.53041773    -4.42753182     0.07971375
#     0.0000000   -560.7370031   -560.7370031     -2.7685434      0.3894192    -41.8319850     -3.7885388      0.1080344 
rm(UK_5param_mcmc2)


# 11.08.2025

UK_4param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_04/UK_newPrior/UK_PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
UK_4param_mcmc2_probs <- coda::as.mcmc(cbind(UK_4param_mcmc2$probabilities, UK_4param_mcmc2$pars))
UK_param_mean4 <- apply(UK_4param_mcmc2_probs, 2, mean)
UK_percentile95_low <- apply(UK_4param_mcmc2_probs,2,quantile,.025)
UK_percentile95_up <- apply(UK_4param_mcmc2_probs,2,quantile,.975)
UK_likelihood_4param <- mean(UK_4param_mcmc2_probs[,"log_likelihood"])
rm(UK_4param_mcmc2)

UK_2param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_04/UK_newPrior/UK_PPxSero_ggCaller_PopPUNK_Null_det_pmcmc_run2.rds")
UK_2param_mcmc2_probs <- coda::as.mcmc(cbind(UK_2param_mcmc2$probabilities, UK_2param_mcmc2$pars))
UK_param_mean2 <- apply(UK_2param_mcmc2_probs, 2, mean)
UK_likelihood_2param <- mean(UK_2param_mcmc2_probs[,"log_likelihood"])
UK_percentile95_low_2param <- apply(UK_2param_mcmc2_probs,2,quantile,.025)
UK_percentile95_up_2param <- apply(UK_2param_mcmc2_probs,2,quantile,.975)
rm(UK_2param_mcmc2)

UK_3param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_04/UK_newPrior/UK_PPxSero_ggCaller_PopPUNK_3param_det_pmcmc_run2.rds")
UK_3param_mcmc2_probs <- coda::as.mcmc(cbind(UK_3param_mcmc2$probabilities, UK_3param_mcmc2$pars))
UK_likelihood_3param <- mean(UK_3param_mcmc2_probs[,"log_likelihood"])
UK_param_mean3 <- apply(UK_3param_mcmc2_probs, 2, mean)
UK_percentile95_low_3param <- apply(UK_3param_mcmc2_probs,2,quantile,.025)
UK_percentile95_up_3param <- apply(UK_3param_mcmc2_probs,2,quantile,.975)
rm(UK_3param_mcmc2)

UK_5param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_04/UK_newPrior/UK_PPxSero_ggCaller_PopPUNK_5param_det_pmcmc_run2.rds")
UK_5param_mcmc2_probs <- coda::as.mcmc(cbind(UK_5param_mcmc2$probabilities, UK_5param_mcmc2$pars))
UK_likelihood_5param <- mean(UK_5param_mcmc2_probs[,"log_likelihood"])
UK_param_mean5 <- apply(UK_5param_mcmc2_probs, 2, mean)
UK_percentile95_low_5param <- apply(UK_5param_mcmc2_probs,2,quantile,.025)
UK_percentile95_up_5param <- apply(UK_5param_mcmc2_probs,2,quantile,.975)
rm(UK_5param_mcmc2)


# Large Pop
UK_4param_mcmc2_largePop <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2026_03_25/UK_LargePop/UK_PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
UK_4param_mcmc2_probs_largePop <- coda::as.mcmc(cbind(UK_4param_mcmc2_largePop$probabilities, UK_4param_mcmc2_largePop$pars))
UK_param_mean4_largePop <- apply(UK_4param_mcmc2_probs_largePop, 2, mean)
UK_percentile95_low_largePop <- apply(UK_4param_mcmc2_probs_largePop,2,quantile,.025)
UK_percentile95_up_largePop <- apply(UK_4param_mcmc2_probs_largePop,2,quantile,.975)
UK_likelihood_4param_largePop <- mean(UK_4param_mcmc2_probs_largePop[,"log_likelihood"])
rm(UK_4param_mcmc2_largePop)

plot_pmcmc_posteriors(UK_4param_mcmc2, UK_4param_mcmc2_largePop, label1 = "15,000", label2 = "50,000", burnin = 10000)

```

# Likelihood ratio test
#  lambda = -2*(x_1-x_2).
# pvalue_testing_null=1-pchisq(lambda,difference_in_degrees_of_freedom) 
```{r}
lambda <- -2 * (UK_likelihood_2param - UK_likelihood_3param)
pchisq(lambda, 1,lower.tail = FALSE)
# 0.4402077

lambda <- -2 * (UK_likelihood_2param - UK_likelihood_4param)
pchisq(lambda, 2,lower.tail = FALSE)
# 1.323822e-48

lambda <- -2 * (UK_likelihood_3param - UK_likelihood_4param)
pchisq(lambda, 1,lower.tail = FALSE)
# 9.551598e-50

lambda <- -2 * (UK_likelihood_2param - UK_likelihood_5param)
pchisq(lambda, 3,lower.tail = FALSE)
# 1.595735e-47

lambda <- -2 * (UK_likelihood_3param - UK_likelihood_5param)
pchisq(lambda, 2,lower.tail = FALSE)
# 1.80618e-48

lambda <- -2 * (UK_likelihood_4param - UK_likelihood_5param)
pchisq(lambda, 1,lower.tail = FALSE)
# 1

UK_2_bic<- -2 * (UK_likelihood_2param) + log(672) * 2 # 1160.003
UK_3_bic<- -2 * (UK_likelihood_3param) + log(672) * 3 # 1165.918
UK_4_bic<- -2 * (UK_likelihood_4param) + log(672) * 4 # 952.5368
UK_5_bic<- -2 * (UK_likelihood_5param) + log(672) * 5 # 959.0727
```

# plot my Mass/UK fit vs Corander fit
```{r}
# copying Corander estimates from their table
Corander_Mass_point_null <- c(0.0375, 0.0073) # v m
Corander_Mass_point_3dof <- c(0.0733, 0.0056, 0.0075) # v m sigmaf
Corander_Mass_low_3dof <- c(0.043, 0.002, 0.0017)
Corander_Mass_up_3dof <- c(0.1207, 0.0131, 0.0234)
Corander_Mass_point_5dof <- c(0.0812, 0.0044, 0.1363, 0.2483, 0.0023) # v m sigmaf propf sigmaw
Corander_Mass_low_5dof <- c(0.0491, 0.0015, 0.0213, 0.1197, 0.001)
Corander_Mass_up_5dof <- c( 0.1254, 0.0165, 0.2113, 0.5448, 0.0514)

Corander_UK_point_null <- c(NA, NA) # v m
Corander_UK_point_3dof <- c(0.1175, 0.0032, 0.0028) # v m sigmaf
Corander_UK_low_3dof <- c(0.0667, 0.0011, 0.001)
Corander_UK_up_3dof <- c(0.2262, 0.0132, 0.0117)
Corander_UK_point_5dof <- c(0.2063, 0.0124, 0.1393, 0.4035, 0.0023)
Corander_UK_low_5dof <- c(0.0832, 0.0012, 0.0121, 0.1005, 0.001)
Corander_UK_up_5dof <- c( 0.3150, 0.0394, 0.2148, 0.5951, 0.0238)


# pre-process my fits


Mass_param_list <- list(Mass_param_mean2, Mass_percentile95_low_2param, Mass_percentile95_up_2param, Mass_param_mean3, Mass_percentile95_low_3param, Mass_percentile95_up_3param, Mass_param_mean5, Mass_percentile95_low_5param, Mass_percentile95_up_5param)

UK_param_list <- list(UK_param_mean2, UK_percentile95_low_2param, UK_percentile95_up_2param, UK_param_mean3, UK_percentile95_low_3param, UK_percentile95_up_3param, UK_param_mean5, UK_percentile95_low_5param, UK_percentile95_up_5param)

for (i in 1:length(Mass_param_list)) {
  Mass_param_list[[i]]["m"] <- exp(Mass_param_list[[i]]["m"])
  UK_param_list[[i]]["m"] <- exp(UK_param_list[[i]]["m"])
  if(i>3){
     Mass_param_list[[i]]["sigma_f"] <- exp(Mass_param_list[[i]]["sigma_f"])
     UK_param_list[[i]]["sigma_f"] <- exp(UK_param_list[[i]]["sigma_f"])
  }
  if(i>6){
    Mass_param_list[[i]]["sigma_w"] <- exp(Mass_param_list[[i]]["sigma_w"])
    UK_param_list[[i]]["sigma_w"] <- exp(UK_param_list[[i]]["sigma_w"])
  }
}



library(dplyr)
library(purrr)

make_df <- function(est, low, up, model, dataset, approach) {
  tibble(
    model = model,
    dataset = dataset,
    approach = approach,
    parameter = c("v", "m", "sigma_f", "prop_f", "sigma_w")[1:length(est)],
    estimate = est,
    lower = low,
    upper = up
  )
}

df <- bind_rows(
  make_df(Corander_Mass_point_null, Corander_Mass_point_null, Corander_Mass_point_null, "2", "Massachusetts", "Corander model"),
  make_df(Corander_Mass_point_3dof, Corander_Mass_low_3dof, Corander_Mass_up_3dof, "3", "Massachusetts", "Corander model"),
  make_df(Corander_Mass_point_5dof, Corander_Mass_low_5dof, Corander_Mass_up_5dof, "5", "Massachusetts", "Corander model"),
  # add the remaining dataset/approach combinations
  make_df(rev(Mass_param_list[[1]])[1:2], rev(Mass_param_list[[2]])[1:2], rev(Mass_param_list[[3]])[1:2], "2", "Massachusetts", "My model"),
  make_df(rev(Mass_param_list[[4]])[1:3], rev(Mass_param_list[[5]])[1:3], rev(Mass_param_list[[6]])[1:3], "3", "Massachusetts", "My model"),
  make_df(rev(Mass_param_list[[7]])[c(1,2, 5, 4, 3)], rev(Mass_param_list[[8]])[c(1,2, 5, 4, 3)], rev(Mass_param_list[[9]])[c(1,2, 5, 4, 3)], "5", "Massachusetts", "My model"),
  
  make_df(Corander_UK_point_null, Corander_UK_point_null, Corander_UK_point_null, "2", "Southampton", "Corander model"),
  make_df(Corander_UK_point_3dof, Corander_UK_low_3dof, Corander_UK_up_3dof, "3", "Southampton", "Corander model"),
  make_df(Corander_UK_point_5dof, Corander_UK_low_5dof, Corander_UK_up_5dof, "5", "Southampton", "Corander model"),
  # add the remaining dataset/approach combinations
  make_df(rev(UK_param_list[[1]])[1:2], rev(UK_param_list[[2]])[1:2], rev(UK_param_list[[3]])[1:2], "2", "Southampton", "My model"),
  make_df(rev(UK_param_list[[4]])[1:3], rev(UK_param_list[[5]])[1:3], rev(UK_param_list[[6]])[1:3], "3", "Southampton", "My model"),
  make_df(rev(UK_param_list[[7]])[c(1,2, 5, 4, 3)], rev(UK_param_list[[8]])[c(1,2, 5, 4, 3)], rev(UK_param_list[[9]])[c(1,2, 5, 4, 3)], "5", "Southampton", "My model")
)

df$parameter <- factor(
  df$parameter,
  levels = c("v", "m", "sigma_f","prop_f","sigma_w")
)
```

```{r}
library(ggplot2)

ggplot(df,
       aes(x = model,
           y = estimate,
           colour = approach)) +
  geom_point(
    position = position_dodge(width = 0.4),
    size = 2
  ) +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    position = position_dodge(width = 0.4),
    width = 0.15
  ) +
  facet_wrap(dataset ~ parameter,
             scales = "free", nrow = 2) +
  theme_minimal(base_size = 18) + labs(x = "Degrees of freedom", y = "Parameter estimate") + labs(color = "Model")
```

```{r}
### Navajo

# Null model
Navajo_2param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2024_12_05/ModelComparison_Navajo/Navajo_PPxSero2vacc_ggCaller_PopPUNK_Null_det_pmcmc_run2.rds")
Navajo_2param_mcmc2_probs <- coda::as.mcmc(cbind(Navajo_2param_mcmc2$probabilities, Navajo_2param_mcmc2$pars))
#coda::effectiveSize(Navajo_2param_mcmc2_probs)
#summary(coda::as.mcmc(Navajo_2param_mcmc2_probs))
Navajo_likelihood_2param <- median(Navajo_2param_mcmc2_probs[,"log_likelihood"])
#mean(Navajo_2param_mcmc2_probs[,"log_likelihood"])
param_mean <- apply(Navajo_2param_mcmc2_probs, 2, mean)
percentile95_low <- apply(Navajo_2param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(Navajo_2param_mcmc2_probs,2,quantile,.975)
#     log_prior log_likelihood  log_posterior              m              v 
#     0.0000000   -878.3942538   -878.3942538     -3.6796603      0.1242759 
#    0.00000000  -881.36090195  -881.36090195    -3.99475933     0.06562337
#     0.0000000   -877.3314368   -877.3314368     -3.2919001      0.2179553 
rm(Navajo_2param_mcmc2)

# 3-parameter model
Navajo_3param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2024_12_05/ModelComparison_Navajo/Navajo_PPxSero2vacc_ggCaller_PopPUNK_3param_det_pmcmc_run2.rds")
Navajo_3param_mcmc2_probs <- coda::as.mcmc(cbind(Navajo_3param_mcmc2$probabilities, Navajo_3param_mcmc2$pars))
#coda::effectiveSize(Navajo_3param_mcmc2_probs)
#summary(coda::as.mcmc(Navajo_3param_mcmc2_probs))
Navajo_likelihood_3param <- median(Navajo_3param_mcmc2_probs[,"log_likelihood"])
#mean(Navajo_3param_mcmc2_probs[,"log_likelihood"])
param_mean <- apply(Navajo_3param_mcmc2_probs, 2, mean)
percentile95_low <- apply(Navajo_3param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(Navajo_3param_mcmc2_probs,2,quantile,.975)
#     log_prior log_likelihood  log_posterior        sigma_f              m              v 
#     0.0000000   -775.6613259   -775.6613259     -4.5054937     -2.8647564      0.2628066
#     0.0000000   -779.0923917   -779.0923917     -5.0989938     -3.3458219      0.1111547
#     0.0000000   -774.0645465   -774.0645465     -3.8534879     -2.2976970      0.5704883
rm(Navajo_3param_mcmc2)

# 4-parameter model GPSC
# and now with winter assignment
Navajo_4param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2025_01_21/Navajo_winter/Navajo_PPxSero2vacc_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
Navajo_4param_mcmc2_probs <- coda::as.mcmc(cbind(Navajo_4param_mcmc2$probabilities, Navajo_4param_mcmc2$pars))
#coda::effectiveSize(Navajo_4param_mcmc2_probs)
#summary(coda::as.mcmc(Navajo_4param_mcmc2_probs))
Navajo_likelihood_4param <- median(Navajo_4param_mcmc2_probs[,"log_likelihood"])
#mean(Navajo_4param_mcmc2_probs[,"log_likelihood"])
param_mean <- apply(Navajo_4param_mcmc2_probs, 2, mean)
percentile95_low <- apply(Navajo_4param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(Navajo_4param_mcmc2_probs,2,quantile,.975)
#     log_prior log_likelihood  log_posterior        sigma_f         prop_f              m              v 
#     0.0000000   -862.1327737   -862.1327737     -4.9293875      0.9293230     -2.7939285      0.1896163
#     0.00000000  -866.25290628  -866.25290628    -5.48570053     0.72694992    -3.13584575     0.09774486
#     0.0000000   -860.0060398   -860.0060398     -4.3237886      0.9975610     -2.4308551      0.3520862
rm(Navajo_4param_mcmc2)
### this does not look good
# I had −611.0528 in TAC report, not -775.1046
# hmm, 21.10.2024 looks more like -775
# and in TAC presentation was also -778

#this one does not exist yet:
# 5-parameter model
Navajo_5param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2024_12_11/ModelComparison_Navajo/Navajo_PPxSero2vacc_ggCaller_PopPUNK_5param_det_pmcmc_run2.rds")
Navajo_5param_mcmc2_probs <- coda::as.mcmc(cbind(Navajo_5param_mcmc2$probabilities, Navajo_5param_mcmc2$pars))
#coda::effectiveSize(Navajo_5param_mcmc2_probs)
#summary(coda::as.mcmc(Navajo_5param_mcmc2_probs))
Navajo_likelihood_5param <- median(Navajo_5param_mcmc2_probs[,"log_likelihood"])
#mean(Navajo_5param_mcmc2_probs[,"log_likelihood"])
param_mean <- apply(Navajo_5param_mcmc2_probs, 2, mean)
percentile95_low <- apply(Navajo_5param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(Navajo_5param_mcmc2_probs,2,quantile,.975)
#     log_prior log_likelihood  log_posterior        sigma_f         prop_f        sigma_w              m              v 
#     0.0000000   -775.7187082   -775.7187082     -4.5390431      0.8471789   -433.9222723     -2.8335532      0.2792345
#     0.0000000   -779.8975459   -779.8975459     -6.2594906      0.2446372   -972.2649520     -3.3022625      0.1162148
#     0.0000000   -773.3938106   -773.3938106     -3.6396775      0.9935490     -4.2750278     -2.1766673      0.6036182
rm(Navajo_5param_mcmc2)

```

# Likelihood ratio test
#  lambda = -2*(x_1-x_2).
# pvalue_testing_null=1-pchisq(lambda,difference_in_degrees_of_freedom) 
```{r}
lambda <- -2 * (Navajo_likelihood_2param - Navajo_likelihood_3param)
pchisq(lambda, 1,lower.tail = FALSE)
#  1.341161e-46

lambda <- -2 * (Navajo_likelihood_2param - Navajo_likelihood_4param)
pchisq(lambda, 2,lower.tail = FALSE)
# 1.930802e-45

lambda <- -2 * (Navajo_likelihood_3param - Navajo_likelihood_4param)
pchisq(lambda, 1,lower.tail = FALSE)
# 0.5011485
```

```{r}
### Navajo 1vacc only (remove last two time points)

# Null model
Navajo_2param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2024_12_16/ModelComparison_Navajo1vacc/Navajo_PPxSero_ggCaller_PopPUNK_Null_det_pmcmc_run2.rds")
Navajo_2param_mcmc2_probs <- coda::as.mcmc(cbind(Navajo_2param_mcmc2$probabilities, Navajo_2param_mcmc2$pars))
#coda::effectiveSize(Navajo_2param_mcmc2_probs)
#summary(coda::as.mcmc(Navajo_2param_mcmc2_probs))
Navajo_likelihood_2param <- median(Navajo_2param_mcmc2_probs[,"log_likelihood"])
#mean(Navajo_2param_mcmc2_probs[,"log_likelihood"])
param_mean <- apply(Navajo_2param_mcmc2_probs, 2, mean)
percentile95_low <- apply(Navajo_2param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(Navajo_2param_mcmc2_probs,2,quantile,.975)
#     log_prior log_likelihood  log_posterior              m              v 
#     0.0000000   -701.8488109   -701.8488109     -3.6866418      0.2503319  
#    0.00000000  -705.01409216  -705.01409216    -4.02234594     0.08606398
#     0.0000000   -700.6613721   -700.6613721     -3.2655542      0.6474211 
rm(Navajo_2param_mcmc2)

# 3-parameter model
Navajo_3param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2024_12_16/ModelComparison_Navajo1vacc/Navajo_PPxSero_ggCaller_PopPUNK_3param_det_pmcmc_run2.rds")
Navajo_3param_mcmc2_probs <- coda::as.mcmc(cbind(Navajo_3param_mcmc2$probabilities, Navajo_3param_mcmc2$pars))
#coda::effectiveSize(Navajo_3param_mcmc2_probs)
#summary(coda::as.mcmc(Navajo_3param_mcmc2_probs))
Navajo_likelihood_3param <- median(Navajo_3param_mcmc2_probs[,"log_likelihood"])
#mean(Navajo_3param_mcmc2_probs[,"log_likelihood"])
param_mean <- apply(Navajo_3param_mcmc2_probs, 2, mean)
percentile95_low <- apply(Navajo_3param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(Navajo_3param_mcmc2_probs,2,quantile,.975)
#     log_prior log_likelihood  log_posterior        sigma_f              m              v 
#     0.0000000   -610.9405032   -610.9405032     -4.3314127     -2.8257480      0.5298071
#     0.0000000   -613.6806814   -613.6806814     -4.9125579     -3.2855823      0.1930156
#     0.0000000   -609.6755103   -609.6755103     -3.7622178     -2.3097844      0.9545226
rm(Navajo_3param_mcmc2)

# 4-parameter model
Navajo_4param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2024_12_16/ModelComparison_Navajo1vacc/Navajo_PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
Navajo_4param_mcmc2_probs <- coda::as.mcmc(cbind(Navajo_4param_mcmc2$probabilities, Navajo_4param_mcmc2$pars))
#coda::effectiveSize(Navajo_4param_mcmc2_probs)
#summary(coda::as.mcmc(Navajo_4param_mcmc2_probs))
Navajo_likelihood_4param <- median(Navajo_4param_mcmc2_probs[,"log_likelihood"])
#mean(Navajo_4param_mcmc2_probs[,"log_likelihood"])
param_mean <- apply(Navajo_4param_mcmc2_probs, 2, mean)
percentile95_low <- apply(Navajo_4param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(Navajo_4param_mcmc2_probs,2,quantile,.975)
#     log_prior log_likelihood  log_posterior        sigma_f         prop_f              m              v 
#     0.0000000   -610.7004392   -610.7004392     -4.1531442      0.8844782     -2.8148291      0.5380844
#     0.0000000   -613.8915897   -613.8915897     -4.7886661      0.7590421     -3.2743311      0.1927550
#     0.0000000   -608.9142628   -608.9142628     -3.4933885      0.9948709     -2.2823608      0.9599296
rm(Navajo_4param_mcmc2)
### this does not look good
# I had −611.0528 in TAC report, not -775.1046
# hmm, 21.10.2024 looks more like -775
# and in TAC presentation was also -778

#this one does not exist yet:
# 5-parameter model
Navajo_5param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2024_12_16/ModelComparison_Navajo1vacc/Navajo_PPxSero_ggCaller_PopPUNK_5param_det_pmcmc_run2.rds")
Navajo_5param_mcmc2_probs <- coda::as.mcmc(cbind(Navajo_5param_mcmc2$probabilities, Navajo_5param_mcmc2$pars))
#coda::effectiveSize(Navajo_5param_mcmc2_probs)
#summary(coda::as.mcmc(Navajo_5param_mcmc2_probs))
Navajo_likelihood_5param <- median(Navajo_5param_mcmc2_probs[,"log_likelihood"])
#mean(Navajo_5param_mcmc2_probs[,"log_likelihood"])
param_mean <- apply(Navajo_5param_mcmc2_probs, 2, mean)
percentile95_low <- apply(Navajo_5param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(Navajo_5param_mcmc2_probs,2,quantile,.975)
#     log_prior log_likelihood  log_posterior        sigma_f         prop_f        sigma_w              m              v 
#     0.0000000   -611.0122849   -611.0122849     -4.4521764      0.7516427   -390.2870444     -2.7957420      0.5501861
#     0.0000000   -614.4283639   -614.4283639     -8.2494761      0.1038850   -966.0429177     -3.2195127      0.2030093
#     0.0000000   -609.0143275   -609.0143275     -2.3695043      0.9933520     -4.0538421     -2.2925315      0.9379878
rm(Navajo_5param_mcmc2)

```

### Nepal
```{r}
# 4-parameter model GPSC
Nepal_4param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2025_01_13/Nepal_GPSCs/Nepal_PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
# and winter
Nepal_4param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2025_01_21/Nepal_winter/Nepal_PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
Nepal_4param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC/2025_01_29/Nepal_year/Nepal_PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
Nepal_4param_mcmc2_probs <- coda::as.mcmc(cbind(Nepal_4param_mcmc2$probabilities, Nepal_4param_mcmc2$pars))
#coda::effectiveSize(Nepal_4param_mcmc2_probs)
#summary(coda::as.mcmc(Nepal_4param_mcmc2_probs))
Nepal_likelihood_4param <- median(Nepal_4param_mcmc2_probs[,"log_likelihood"])
#mean(Nepal_4param_mcmc2_probs[,"log_likelihood"])
param_estimates <- map_estimate(Nepal_4param_mcmc2$pars)
param_mean <- apply(Nepal_4param_mcmc2_probs, 2, mean)
print(param_mean)
percentile95_low <- apply(Nepal_4param_mcmc2_probs,2,quantile,.025)
percentile95_up <- apply(Nepal_4param_mcmc2_probs,2,quantile,.975)
#     log_prior log_likelihood  log_posterior        sigma_f         prop_f              m              v 
#  0.000000e+00  -2.794504e+03  -2.794504e+03  -2.381981e+00   3.565618e-01  -2.525765e+00   2.693263e-02 
#  0.000000e+00  -2.801638e+03  -2.801638e+03  -2.922157e+00   3.495622e-01  -3.000324e+00   2.244008e-03
#  0.000000e+00  -2.791667e+03  -2.791667e+03  -1.639195e+00   4.343989e-01  -1.903038e+00   7.172079e-02

# check the covariance matrix
cov_matrix <- cov(Nepal_4param_mcmc2$pars)
saveRDS(cov_matrix, "Nepal_cov_matrix.rds")
# this is not the proposal matrix, this could be a future proposal matrix based on the covariances found in this run
# I think that might be a smart move

rm(Nepal_4param_mcmc2)

# 11.08.2025 
# 27.08.2025
# 17.11.2025

#Nepal_4param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_05/Nepal_newPrior/Nepal_PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
#Nepal_4param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_27/Nepal_newPrioRunWithGoodParams/Nepal_PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
Nepal_4param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_11_17/Nepal_NewGPSCorder_10k_fit2/Nepal_PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
Nepal_4param_mcmc2_probs <- coda::as.mcmc(cbind(Nepal_4param_mcmc2$probabilities, Nepal_4param_mcmc2$pars))
saveRDS(Nepal_4param_mcmc2_probs, "Nepal_4param_mcmc2_probs.rds")
Nepal_4param_mcmc2_probs <- Nepal_4param_mcmc2_probs[1:37000,]
Nepal_likelihood_4param <- mean(Nepal_4param_mcmc2_probs[,"log_likelihood"])
Nepal_param_mean4 <- apply(Nepal_4param_mcmc2_probs, 2, mean)
Nepal_percentile95_low <- apply(Nepal_4param_mcmc2_probs,2,quantile,.025)
Nepal_percentile95_up <- apply(Nepal_4param_mcmc2_probs,2,quantile,.975)
rm(Nepal_4param_mcmc2)

#Nepal_2param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_05/Nepal_newPrior/Nepal_PPxSero_ggCaller_PopPUNK_Null_det_pmcmc_run2.rds")
#Nepal_2param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_27/Nepal_newPrioRunWithGoodParams/Nepal_PPxSero_ggCaller_PopPUNK_Null_det_pmcmc_run2.rds")
Nepal_2param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_11_17/Nepal_NewGPSCorder_10k_fit2/Nepal_PPxSero_ggCaller_PopPUNK_Null_det_pmcmc_run2.rds")
Nepal_2param_mcmc2_probs <- coda::as.mcmc(cbind(Nepal_2param_mcmc2$probabilities, Nepal_2param_mcmc2$pars))
saveRDS(Nepal_2param_mcmc2_probs, "Nepal_2param_mcmc2_probs.rds")
Nepal_likelihood_2param <- mean(Nepal_2param_mcmc2_probs[,"log_likelihood"])
Nepal_param_mean2 <- apply(Nepal_2param_mcmc2_probs, 2, mean)
rm(Nepal_2param_mcmc2)

#Nepal_3param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_05/Nepal_newPrior/Nepal_PPxSero_ggCaller_PopPUNK_3param_det_pmcmc_run2.rds")
#Nepal_3param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_27/Nepal_newPrioRunWithGoodParams/Nepal_PPxSero_ggCaller_PopPUNK_3param_det_pmcmc_run2.rds")
Nepal_3param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_11_17/Nepal_NewGPSCorder_10k_fit2/Nepal_PPxSero_ggCaller_PopPUNK_3param_det_pmcmc_run2.rds")
Nepal_3param_mcmc2_probs <- coda::as.mcmc(cbind(Nepal_3param_mcmc2$probabilities, Nepal_3param_mcmc2$pars))
Nepal_likelihood_3param <- mean(Nepal_3param_mcmc2_probs[,"log_likelihood"])
Nepal_param_mean3 <- apply(Nepal_3param_mcmc2_probs, 2, mean)
rm(Nepal_3param_mcmc2)

#Nepal_5param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_05/Nepal_newPrior/Nepal_PPxSero_ggCaller_PopPUNK_5param_det_pmcmc_run2.rds")
#Nepal_5param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_27/Nepal_newPrioRunWithGoodParams/Nepal_PPxSero_ggCaller_PopPUNK_5param_det_pmcmc_run2.rds")
Nepal_5param_mcmc2 <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_11_17/Nepal_NewGPSCorder_10k_fit2/Nepal_PPxSero_ggCaller_PopPUNK_5param_det_pmcmc_run2.rds")
Nepal_5param_mcmc2_probs <- coda::as.mcmc(cbind(Nepal_5param_mcmc2$probabilities, Nepal_5param_mcmc2$pars))
Nepal_param_mean5 <- apply(Nepal_5param_mcmc2_probs, 2, mean)
Nepal_likelihood_5param <- mean(Nepal_5param_mcmc2_probs[,"log_likelihood"])
rm(Nepal_5param_mcmc2)

# Large Pop
Nepal_4param_mcmc2_largePop <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2026_03_25/Nepal_LargePop_fit2/Nepal_PPxSero_ggCaller_PopPUNK_4param_det_pmcmc_run2.rds")
Nepal_4param_mcmc2_probs_largePop <- coda::as.mcmc(cbind(Nepal_4param_mcmc2_largePop$probabilities, Nepal_4param_mcmc2_largePop$pars))
Nepal_param_mean4_largePop <- apply(Nepal_4param_mcmc2_probs_largePop, 2, mean)
Nepal_percentile95_low_largePop <- apply(Nepal_4param_mcmc2_probs_largePop,2,quantile,.025)
Nepal_percentile95_up_largePop <- apply(Nepal_4param_mcmc2_probs_largePop,2,quantile,.975)
Nepal_likelihood_4param_largePop <- mean(Nepal_4param_mcmc2_probs_largePop[,"log_likelihood"])
rm(Nepal_4param_mcmc2_largePop)

plot_pmcmc_posteriors(Nepal_4param_mcmc2, Nepal_4param_mcmc2_largePop, label1 = "15,000", label2 = "82,921", burnin = 5000)

```

```{r}
lambda <- -2 * (Nepal_likelihood_2param - Nepal_likelihood_3param)
pchisq(lambda, 1,lower.tail = FALSE)
# 3.416704e-34

lambda <- -2 * (Nepal_likelihood_2param - Nepal_likelihood_4param)
pchisq(lambda, 2,lower.tail = FALSE)
# 8.240171e-43

lambda <- -2 * (Nepal_likelihood_3param - Nepal_likelihood_4param)
pchisq(lambda, 1,lower.tail = FALSE)
# 1.822945e-11

lambda <- -2 * (Nepal_likelihood_2param - Nepal_likelihood_5param)
pchisq(lambda, 3,lower.tail = FALSE)
# 1.030786e-41

lambda <- -2 * (Nepal_likelihood_3param - Nepal_likelihood_5param)
pchisq(lambda, 2,lower.tail = FALSE)
# 1.757722e-10

lambda <- -2 * (Nepal_likelihood_4param - Nepal_likelihood_5param)
pchisq(lambda, 1,lower.tail = FALSE)
# 1

2 * 1 - 2 * (-Nepal_likelihood_2param)
# -2421.05
3 * 1 - 2 * (-Nepal_likelihood_3param)
# -2271.398
4 * 1 - 2 * (-Nepal_likelihood_4param)
# -2225.246
5 * 1 - 2 * (-Nepal_likelihood_5param)
# -2224.474

Nepal_4param_logLik <- rep(NA, 1)
Nepal_4param_logLik$logLik <- Nepal_likelihood_4param

Nepal_2_bic<- -2 * (Nepal_likelihood_2param) + log(2001) * 2 # 2438.253
Nepal_3_bic<- -2 * (Nepal_likelihood_3param) + log(2001) * 3 # 2297.202
Nepal_4_bic<- -2 * (Nepal_likelihood_4param) + log(2001) * 4 # 2259.651
Nepal_5_bic<- -2 * (Nepal_likelihood_5param) + log(2001) * 5 # 2267.481
```

```{r}
library(ggplot2)
# plot results
joined_mcmc_param_fits <- data.frame(params = rep(c("NFDS strength", "genes under NFDS", "immigration rate", "vaccination strength"), 3),
                                     population = c(rep("US",4), rep("UK",4), rep("Nepal",4)),
                                     mean = c(Mass_param_mean4[4:7], UK_param_mean4[4:7], Nepal_param_mean4[4:7]),
                                     lower = c(Mass_percentile95_low[4:7], UK_percentile95_low[4:7], Nepal_percentile95_low[4:7]),
                                     upper = c(Mass_percentile95_up[4:7], UK_percentile95_up[4:7], Nepal_percentile95_up[4:7]))
joined_mcmc_param_fits[c(1,3,5,7,9,11),c("mean", "lower", "upper")] <- exp(joined_mcmc_param_fits[c(1,3,5,7,9,11),c("mean", "lower", "upper")])

ggplot(joined_mcmc_param_fits,aes(x=params,y=mean, colour = population, group = interaction(population, params))) +
  geom_errorbar(aes(ymin=lower,ymax=upper,color=population,width=0.2),position = position_dodge(width=0.75),, linewidth = 1) +
  geom_point(position = position_dodge(width=0.75),size = 3) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20)) +
  labs(y= "Parameter estimate", x = "Parameter")

ggplot(joined_mcmc_param_fits,aes(x=population,y=mean, colour = population, group = interaction(population, params))) +
  geom_errorbar(aes(ymin=lower,ymax=upper,color=population,width=0.2), linewidth = 1) +
  geom_point(size = 3) +
  facet_wrap(~params, ncol = 2) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20)) +
  labs(y= "Parameter estimate", x = "Population")

joined_mcmc_param_fits$params <- factor(joined_mcmc_param_fits$params, levels = c("vaccination strength","NFDS strength","genes under NFDS","immigration rate"))
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

ggplot(joined_mcmc_param_fits,aes(x=population,y=mean, colour = population, group = interaction(population, params))) +
  geom_errorbar(aes(ymin=lower,ymax=upper,color=population,width=0.2), linewidth = 1.5) +
  geom_point(size = 4) +
  facet_wrap(~params, ncol = 4) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20), panel.spacing.x = unit(2, "lines")) +
  labs(y= "Parameter estimate", x = "Population") + scale_color_manual(values=c(cbbPalette[1], cbbPalette[3], cbbPalette[7]))


ggplot(joined_mcmc_param_fits,aes(x=population,y=mean, colour = population, group = interaction(population, params))) +
  geom_errorbar(aes(ymin=lower,ymax=upper,color=population,width=0.2), linewidth = 1.5) +
  geom_point(size = 4) +
  facet_wrap(~params, ncol = 4, scales = "free") +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20), panel.spacing.x = unit(2, "lines")) +
  labs(y= "Parameter estimate", x = "Population") + scale_color_manual(values=c(cbbPalette[1], cbbPalette[3], cbbPalette[7]))
```