---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---


```{r}
path_to_data <- "~/Documents/PhD_Project/Data/"
library(readxl)
```

#### read in meta data
```{r}
joined_metadata_list <- vector(mode = "list", length = 3) # meta data list for UK, US, Nepal
names(joined_metadata_list) <- c("UK", "US", "Nepal")

joined_PopPUNK_list <- vector(mode = "list", length = 3) # PopPUNK list for UK, US, Nepal
names(joined_metadata_list) <- c("UK", "US", "Nepal")
```

### Nepal
```{r}
# GPS1
Nepal_meta_data_gps1 <- as.data.frame(readxl::read_excel("~/Documents/PhD_Project/Data/StrepPneumo_Nepal_gps_v2/GPS1_and_GPS2_Nepal_dataset_3Dec2024.xlsx", sheet = "GPS1"))
Nepal_meta_data_gps1_selection <- Nepal_meta_data_gps1[,c(2:6,8:13, 18, 25, 75)]
Nepal_meta_data_gps1_selection$GPS <- 1

Nepal_meta_data_gps1_selection_v2 <- Nepal_meta_data_gps1[,c(2:4,8,9, 12, 13, 18, 25, 75)]
Nepal_meta_data_gps1_selection_v2$GPS <- 1

#GPS2 serotypes
Nepal_gps2_serotyping <- read.csv("~/Documents/PhD_Project/Data/StrepPneumo_Nepal_gps_v2/SerotypeSummary.csv", header = FALSE)
colnames(Nepal_gps2_serotyping) <- c("Sample","Serotype","Genetic_Variant","Contamination_Status")

spl_serotype_names <- function(serotype_str){
  split1 <- strsplit(serotype_str, "/data/SerotypingResults/", fixed = TRUE)[[1]][2]
  split2 <- strsplit(split1, "_RESULT", fixed = TRUE)[[1]][1]
  split2
}
Nepal_gps2_serotyping$Sample <- as.vector(sapply(Nepal_gps2_serotyping$Sample,spl_serotype_names))

gps2_serotype_dict <- Nepal_gps2_serotyping$Serotype
names(gps2_serotype_dict) <- Nepal_gps2_serotyping$Sample

# GPS2
Nepal_meta_data_gps2 <- as.data.frame(readxl::read_excel("~/Documents/PhD_Project/Data/StrepPneumo_Nepal_gps_v2/GPS1_and_GPS2_Nepal_dataset_3Dec2024.xlsx", sheet = "GPS2")) 
# length(which(Nepal_meta_data_gps2$QC == "FAIL")) 90 --> 90 of these sequences did not pass quality control--> I will filter these out now (Oct 2025)
Nepal_meta_data_gps2_selection <- Nepal_meta_data_gps2[which(Nepal_meta_data_gps2$QC != "FAIL"),c(1,2,4:11, 16,21,22)]
Nepal_meta_data_gps2_selection$Clinical_manifestation[which(Nepal_meta_data_gps2_selection$Clinical_manifestation == "OTHER")] <- "CARRIAGE"
Nepal_meta_data_gps2_selection$GPS <- 2
colnames(Nepal_meta_data_gps2_selection)[2] <- "Public_name"
Nepal_meta_data_gps2_selection$"In_silico_serotype" <- rep(NA, nrow(Nepal_meta_data_gps2_selection))
Nepal_meta_data_gps2_selection$"In_silico_serotype" <- gps2_serotype_dict[Nepal_meta_data_gps2_selection$"Lane_id"]
#note: there are 66 entries in here that do not have genome data available 

comm_carriage_dict <- (Nepal_meta_data_gps2$Study_name == "COMMUNITY CARRIAGE")
names(comm_carriage_dict) <- Nepal_meta_data_gps2$Lane_id

# new version of GPSC metadata
library(readr)
Nepal_meta_data_gps2_v2 <- read_csv("~/Documents/PhD_Project/Data/StrepPneumo_Nepal_gps_v2/GPS_Nepal_data_19Sep2025.csv")
filter_for_carriage <- (Nepal_meta_data_gps2_v2$Manifestation == "CARRIAGE")
names(filter_for_carriage) <- Nepal_meta_data_gps2_v2$Lane_id
for (i in names(filter_for_carriage)) {
  if(!is.na(comm_carriage_dict[i]) & comm_carriage_dict[i]){
    filter_for_carriage[i] <- TRUE
  }
}
Nepal_meta_data_gps2_v2_selection <- Nepal_meta_data_gps2_v2[filter_for_carriage,c(2:4, 7:9, 11, 22,23)]
#Nepal_meta_data_gps2_v2_selection$Month <- rep(NA, nrow(Nepal_meta_data_gps2_v2_selection))
#Nepal_meta_data_gps2_v2_selection$Phenotypic_serotype <- rep(NA, nrow(Nepal_meta_data_gps2_v2_selection))
#Nepal_meta_data_gps2_v2_selection$Clinical_manifestation[which(Nepal_meta_data_gps2_selection$Clinical_manifestation == "OTHER")] <- "CARRIAGE"
#Nepal_meta_data_gps2_v2_selection$GPS <- 2
for (i in unique(Nepal_meta_data_gps2_v2_selection$Year)) {
  print(paste(i, length(which(Nepal_meta_data_gps2_v2_selection$Year == i)), sep = " "))
}
# total 1881 nrow(Nepal_meta_data_gps2_v2_selection) 
# will use this going forward (Oct 2025)

#install.packages("xtable")
#library(xtable)
#print(xtable(data.frame(Nepal_meta_data_gps2_v2_selection$Lane_id[1:100]), type = "latex"), file = "/Users/llorenz/Documents/PhD_Project/Papers/NFDS_Model/Data/NepalAccNos.tex")
#Nepall_accNos <- data.frame(Nepal_meta_data_gps2_v2_selection$Lane_id)
#colnames(Nepall_accNos) <- "Lane_id"
#write.csv(Nepall_accNos, file = "/Users/llorenz/Documents/PhD_Project/Papers/NFDS_Model/Data/NepalAccNos.csv")

# use this updated metadata sheet for serotyping
#gps2_serotype_dict_v2 <- Nepal_meta_data_gps2_v2$In_silico_serotype
#names(gps2_serotype_dict_v2) <- Nepal_meta_data_gps2_v2$Lane_id

#Nepal_meta_data_gps2_v2_selection <- Nepal_meta_data_gps2_selection
#Nepal_meta_data_gps2_v2_selection$"In_silico_serotype" <- rep(NA, nrow(Nepal_meta_data_gps2_v2_selection))
#Nepal_meta_data_gps2_v2_selection$"In_silico_serotype" <- gps2_serotype_dict_v2[Nepal_meta_data_gps2_v2_selection$"Lane_id"]

# check how much these two serotyping methods really differ
same <- 0
diff <- 0
diff_names <- c()
#diff_sero1 <- rep(NA, 449)
not_in_set <- 0
not_in_set_names <- c()
for (sero_i in names(gps2_serotype_dict)) {
  if(!is.na(gps2_serotype_dict_v2[sero_i]) & gps2_serotype_dict[sero_i] == gps2_serotype_dict_v2[sero_i]){
    same <- same + 1
  } else if(!is.na(gps2_serotype_dict_v2[sero_i])){
    diff <- diff + 1
    diff_names <- c(diff_names, sero_i)
    print(paste(gps2_serotype_dict[sero_i], " and ", gps2_serotype_dict_v2[sero_i], sep = ""))
  } else{
    not_in_set <- not_in_set + 1
    not_in_set_names <- c(not_in_set_names, sero_i)
  }
}
#actually, not really any differences. Just slightly different names, e.g. 06A and 6A.
# not_in_set_names seem to be the ones that did not pass quality control and do not appear in new Nepal dataset

Nepal_meta_data_gps_joined_old <- rbind(Nepal_meta_data_gps1_selection, Nepal_meta_data_gps2_selection)
#Nepal_meta_data_gps_joined_v2 <- rbind(Nepal_meta_data_gps1_selection, Nepal_meta_data_gps2_v2_selection)

# 2610 samples
# now filter by those that are carriage/"other"

Nepal_meta_data_gps_joined_old <- Nepal_meta_data_gps_joined_old[which(Nepal_meta_data_gps_joined_old$Clinical_manifestation == "CARRIAGE"),]
#Nepal_meta_data_gps_joined_v2 <- Nepal_meta_data_gps_joined_v2[which(Nepal_meta_data_gps_joined_v2$Clinical_manifestation == "CARRIAGE"),]
# 2001 samples

# add phenotypic serotype to dataset
pheno_sero_dict <- Nepal_meta_data_gps_joined_old$Phenotypic_serotype
names(pheno_sero_dict) <- Nepal_meta_data_gps_joined_old$Lane_id
Nepal_meta_data_gps2_v2_selection$Phenotypic_serotype <- pheno_sero_dict[Nepal_meta_data_gps2_v2_selection$Lane_id]

Nepal_meta_data_gps_joined <- Nepal_meta_data_gps2_v2_selection

Nepal_time_points <- sort(unique(Nepal_meta_data_gps_joined$Year))

Nepal_SeqYear_dict <- Nepal_meta_data_gps_joined$Year
names(Nepal_SeqYear_dict) <- Nepal_meta_data_gps_joined$Lane_id

Nepal_meta_IDtoInd <- 1:nrow(Nepal_meta_data_gps_joined)
names(Nepal_meta_IDtoInd) <- Nepal_meta_data_gps_joined$Lane_id

Nepal_MetaDataGPSC_dict <- Nepal_meta_data_gps_joined$GPSC
names(Nepal_MetaDataGPSC_dict) <- Nepal_meta_data_gps_joined$Lane_id
```

# create serotype dictionary
```{r}
# unique(Nepal_meta_data_gps_joined$In_silico_serotype)
# collate "NCC1_PSPK_NON_ENCAPSULATED", "ALTERNATIVE_ALIB_NT", "SWISS_NT", "UNTYPABLE"  to "NT"
# try to substitute "COVERAGE TOO LOW" with phenotypic serotype

Nepal_Serotype_dict <- rep(NA, nrow(Nepal_meta_data_gps_joined))
names(Nepal_Serotype_dict) <- Nepal_meta_data_gps_joined$Lane_id

Nepal_Serotype_dict[Nepal_meta_data_gps_joined$Lane_id[which(Nepal_meta_data_gps_joined$In_silico_serotype != "COVERAGE TOO LOW")]] <- Nepal_meta_data_gps_joined$In_silico_serotype[which(Nepal_meta_data_gps_joined$In_silico_serotype != "COVERAGE TOO LOW")]

Nepal_Serotype_dict[Nepal_meta_data_gps_joined$Lane_id[which(Nepal_meta_data_gps_joined$In_silico_serotype == "COVERAGE TOO LOW")]] <- Nepal_meta_data_gps_joined$Phenotypic_serotype[which(Nepal_meta_data_gps_joined$In_silico_serotype == "COVERAGE TOO LOW")]

Nepal_Serotype_dict[which(Nepal_Serotype_dict == "UNTYPABLE" | Nepal_Serotype_dict == "SWISS_NT" | Nepal_Serotype_dict == "ALTERNATIVE_ALIB_NT" | Nepal_Serotype_dict == "untypable" | Nepal_Serotype_dict == "NCC1_PSPK_NON_ENCAPSULATED" | Nepal_Serotype_dict == "NCC3_aliD_non_encapsulated" | Nepal_Serotype_dict == "NCC2_S_mitis_aliC_aliD_non_encapsulated" | Nepal_Serotype_dict == "NCC2_aliC_aliD_non_encapsulated")] <- "NT"

# unique(Nepal_Serotype_dict)
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "6E(6B)")] <- "6B"
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "6A(6A-III)")] <- "6A"
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "15B")] <- "15B/15C"
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "15C")] <- "15B/15C"
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "19A(19A-I/19A-II)")] <- "19A"
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "23B1")] <- "23B"
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "24B/24C/24F")] <- "SEROGROUP 24"
Nepal_Serotype_dict[which(Nepal_Serotype_dict == "24F")] <- "SEROGROUP 24"

# length(unique(Nepal_Serotype_dict)) 68

Nepal_has_serotype <- rep(TRUE, nrow(Nepal_meta_data_gps_joined))
names(Nepal_has_serotype) <- Nepal_meta_data_gps_joined$Lane_id
```

# read in PopPUNK results
```{r}
PP_all <- read.csv(paste(path_to_data, "StrepPneumo_Nepal_gps_v2/poppunk_clusters_GPSCs_v9/poppunk_clusters_external_clusters.csv",sep = ""))
# since Nepal was in the database, I had to run PopPUNK with --write-references
# so now I need to filter this dataset to only have the real Nepal sequences

# remove appendix "_contigs_velvet"
#remove_contigs_velvet <- function(str_with_app){
#  str_without_app <- strsplit(str_with_app, "_contigs_velvet", fixed = TRUE)[[1]][1]
#  str_without_app
#}

#PP_all$Taxon <- sapply(PP_all$Taxon, remove_contigs_velvet)
PP_all <- PP_all[!duplicated(PP_all$sample),]

PP_taxon_ind_dict <- 1:nrow(PP_all)
names(PP_taxon_ind_dict) <- PP_all$sample

PP_is_Nepal_dict <- rep(FALSE,nrow(PP_all))
names(PP_is_Nepal_dict) <- PP_all$sample

for (i in 1:length(Nepal_meta_data_gps_joined$Lane_id)) {
  if(Nepal_has_serotype[i]){ # checks if we have serotype for this sequence and only includes it in PP dataset if yes
    PP_is_Nepal_dict[Nepal_meta_data_gps_joined$Lane_id[i]] <- TRUE
  }
}

Nepal_PP <- PP_all[which(PP_is_Nepal_dict[PP_all$sample]),]
Nepal_PP <- Nepal_PP[which(!(is.na(Nepal_PP$GPSC))),]
no_Nepal_PP <- length(unique(Nepal_PP$GPSC)) 
# now: 170 GPSCs

Nepal_seq_clusters_dict <- Nepal_PP$GPSC
names(Nepal_seq_clusters_dict) <- Nepal_PP$sample

# based on metadata
Nepal_seq_clusters_dict_metadata <- Nepal_meta_data_gps2_v2_selection$GPSC
names(Nepal_seq_clusters_dict_metadata) <- Nepal_meta_data_gps2_v2_selection$Lane_id

for (sample in names(Nepal_seq_clusters_dict)) {
  if(!(is.na(Nepal_seq_clusters_dict_metadata[sample])) & Nepal_seq_clusters_dict[sample] == Nepal_seq_clusters_dict_metadata[sample]){
    #print("same!")
  }
  else{
    print(paste(sample, Nepal_seq_clusters_dict[sample], Nepal_seq_clusters_dict_metadata[sample]))
  }
}
# main disagreement: 10;209 vs 10 or 209
unique(Nepal_meta_data_gps2_v2_selection$In_silico_serotype[which(Nepal_meta_data_gps2_v2_selection$GPSC == "10")])
# "7B"           "10A"          "19F"          "17F"          "6A"           "19A"          "22F"          "SEROGROUP 24" "23F"          "6C"
table(Nepal_meta_data_gps2_v2_selection$Year[which(Nepal_meta_data_gps2_v2_selection$GPSC == "10")])
unique(Nepal_meta_data_gps2_v2_selection$In_silico_serotype[which(Nepal_meta_data_gps2_v2_selection$GPSC == "209")])
# "35B"
table(Nepal_meta_data_gps2_v2_selection$Year[which(Nepal_meta_data_gps2_v2_selection$GPSC == "209")])
```

# filter meta data by PopPUNK sequences (PopPUNK qc removed some of the sequences)
```{r}
Nepal_meta_dataPP_metaGPSC <- Nepal_meta_data_gps_joined[sort(Nepal_meta_IDtoInd[Nepal_PP$sample]),c("Lane_id", "Year")]
Nepal_meta_dataPP_metaGPSC$Serotype <- Nepal_Serotype_dict[Nepal_meta_dataPP_metaGPSC$Lane_id]
Nepal_meta_dataPP_metaGPSC$GPSC <- Nepal_seq_clusters_dict_metadata[Nepal_meta_dataPP_metaGPSC$Lane_id] ### changed - this is now based on meta data GPSCs
colnames(Nepal_meta_dataPP_metaGPSC) <- c("Lane_id", "Year", "Serotype", "GPSC")

#rownames(Nepal_meta_dataPP_metaGPSC) <- 1:nrow(Nepal_meta_dataPP_metaGPSC)

Nepal_meta_IDtoIndPP <- 1:nrow(Nepal_meta_dataPP_metaGPSC)
names(Nepal_meta_IDtoIndPP) <- Nepal_meta_dataPP_metaGPSC$Lane_id
```

```{r}
# save Nepal metadata in metadata list
joined_metadata_list[["Nepal"]] <- Nepal_meta_dataPP_metaGPSC

saveRDS(Nepal_meta_dataPP_metaGPSC, "Nepal_meta_dataPP_metaGPSC.rds")
```

```{r}
# Add Serotype Information to PP
Nepal_PP$Serotype <- Nepal_Serotype_dict[Nepal_PP$sample]

# create dictionary for VTs of the PCV10
# 1, 4, 5, 6b, 7f, 9V, 14, 18c, 19f and 23f serotypes (source: NCBI)
PCV10_VTsPP <- rep("NVT",length(unique(Nepal_PP$Serotype)))
names(PCV10_VTsPP) <- unique(Nepal_PP$Serotype)
PCV10_VTsPP[c("1","4","5","6A","6B", "7F", "9V", "14", "18C", "19F", "23F")] <- "VT" # also added 6A

Nepal_PP$VT <- PCV10_VTsPP[Nepal_PP$Serotype]

### changed
# will substitute my GPSCs by the ones from the metadata
Nepal_PP_metaGPSC <- Nepal_PP
for (i in 1:nrow(Nepal_PP_metaGPSC)) {
  if(! is.na(Nepal_seq_clusters_dict_metadata[Nepal_PP_metaGPSC$sample[i]])){ # some samples have no GPSC in the meta data :-/
    Nepal_PP_metaGPSC$GPSC[i] <- Nepal_seq_clusters_dict_metadata[Nepal_PP_metaGPSC$sample[i]]
  } else{
    Nepal_seq_clusters_dict_metadata[Nepal_PP_metaGPSC$sample[i]] <- Nepal_PP$GPSC[i]
  }
  
}



#joined_PopPUNK_list[["Nepal"]] <- Nepal_PP

saveRDS(Nepal_PP_metaGPSC, "Nepal_PP_metaGPSC.rds")

no_Nepal_PP <- length(unique(Nepal_PP_metaGPSC$GPSC)) 
#Nepal_seq_clusters_dict <- Nepal_PP_metaGPSC$GPSC
#names(Nepal_seq_clusters_dict) <- Nepal_PP_metaGPSC$sample
```



### ggCaller 

# get Nepal part of joined ggCaller
```{r}

joined_ggCaller_bool <- readRDS("joined_ggCaller_bool.rds")

# filter ggCaller results by Nepal 
Nepal_ggCallerPP_bool <- data.frame(matrix(0, nrow = nrow(joined_ggCaller_bool), ncol = length(Nepal_PP_metaGPSC$sample)+1))
Nepal_ggCallerPP_bool[-1,-1] <- joined_ggCaller_bool[-1,c(Nepal_PP_metaGPSC$sample)]
colnames(Nepal_ggCallerPP_bool) <- c("Gene",Nepal_PP_metaGPSC$sample)
Nepal_ggCallerPP_bool[1,] <- c("Gene",Nepal_PP_metaGPSC$sample)
Nepal_ggCallerPP_bool[,1] <- joined_ggCaller_bool[,1]
```

# process Nepal ggCaller
# split ggCaller gene presence absence matrix into different time points
```{r}
### continue working here
Nepal_time_points_selected <- min(Nepal_time_points):max(Nepal_time_points)
Nepal_ggCaller_byYear <- vector(mode = "list", length = length(Nepal_time_points_selected))

for (i in 1:length(Nepal_time_points_selected)) {
  Nepal_ggCaller_year <- data.frame(matrix(0, nrow = nrow(Nepal_ggCallerPP_bool), ncol = length(which(Nepal_meta_dataPP_metaGPSC$Year==Nepal_time_points_selected[i]))+1)) # initialize data frame
  Nepal_ggCaller_year[1,-1] <- Nepal_ggCallerPP_bool[1,c(FALSE,Nepal_SeqYear_dict[unlist(Nepal_ggCallerPP_bool[1,-1])]==Nepal_time_points_selected[i])] # fill in first row (seq names)
  Nepal_ggCaller_year[-1,1] <- Nepal_ggCallerPP_bool[-1,1] # fill in first column (gene cluster names)
  Nepal_ggCaller_year[-1,-1] <- Nepal_ggCallerPP_bool[-1,c(FALSE,Nepal_SeqYear_dict[unlist(Nepal_ggCallerPP_bool[1,-1])]==Nepal_time_points_selected[i])] # fill in presence absence information
  Nepal_ggCaller_byYear[[i]] <- Nepal_ggCaller_year
}
```

# find intermediate frequency genes
# I will combine all pre-vaccine data points for this
```{r}
sum_as_int <- function(x){
  sum(as.integer(x))
}

Nepal_gene_freq_preVac <- rep(0, nrow(Nepal_ggCaller_byYear[[1]])-1)
#Nepal_time_points_preVac <- c(5,8,10) # updated this now too (July 2025) to use all pre-vac samples
Nepal_time_points_preVac <- c(1,4,6) # Oct 2025, correction after just using carriage data
preVac_datapoints <- 0
for (i in Nepal_time_points_preVac) {
  Nepal_gene_freq_preVac <- Nepal_gene_freq_preVac + apply(Nepal_ggCaller_byYear[[i]][-1,-1],1, sum_as_int)
  preVac_datapoints <- preVac_datapoints + length(Nepal_ggCaller_byYear[[i]][1,-1])
}
Nepal_gene_freq_preVac <- Nepal_gene_freq_preVac / preVac_datapoints

Nepal_gene_filter <- as.integer(Nepal_gene_freq_preVac<=0.95 & Nepal_gene_freq_preVac>=0.05)
# now 2340 (after filtering for carriage only)
```

# create intermediate frequency dataframes
```{r}
Nepal_ggCaller_byYear_intermed <- vector(mode = "list", length = length(Nepal_time_points_selected))

for (i in 1:length(Nepal_time_points_selected)) {
  Nepal_ggCaller_year <- data.frame(matrix(0, nrow = sum(Nepal_gene_filter)+1,ncol = length(which(Nepal_meta_dataPP_metaGPSC$Year==Nepal_time_points_selected[i]))+1))
  if(ncol(Nepal_ggCaller_byYear[[i]])>1){
    Nepal_ggCaller_year <-  Nepal_ggCaller_byYear[[i]][c(1,which(Nepal_gene_filter==1)+1),]
  }
  else{
    Nepal_ggCaller_year <-  data.frame(matrix(Nepal_ggCaller_byYear[[2]][c(1,which(Nepal_gene_filter==1)+1),1], nrow = sum(Nepal_gene_filter)+1,ncol = 1))
  }
  Nepal_ggCaller_byYear_intermed[[i]] <- Nepal_ggCaller_year 
}

Nepal_ggCaller_intermed <- data.frame(matrix(0, nrow = sum(Nepal_gene_filter)+1,ncol = ncol(Nepal_ggCallerPP_bool)))
Nepal_ggCaller_intermed <-  Nepal_ggCallerPP_bool[c(1,which(Nepal_gene_filter==1)+1),]

# use 2009 as preVac reference
#Nepal_ggCaller_intermed_preVac <- Nepal_ggCaller_byYear_intermed[[5]]
#Nepal_ggCaller_intermed_preVac <- Nepal_ggCaller_byYear_intermed[[1]]
# no, use all pre-vac years
#Nepal_ggCaller_intermed_preVac <- cbind(Nepal_ggCaller_byYear_intermed[[5]],Nepal_ggCaller_byYear_intermed[[8]][,-1],Nepal_ggCaller_byYear_intermed[[10]][,-1])
Nepal_ggCaller_intermed_preVac <- cbind(Nepal_ggCaller_byYear_intermed[[1]],Nepal_ggCaller_byYear_intermed[[4]][,-1],Nepal_ggCaller_byYear_intermed[[6]][,-1])

```

# create consensus gene presence absence matrices for the clusters
```{r}
Nepal_clusters <- readRDS("Nepal_PP_metaGPSC.rds")

cons_genomes <- function(x){
  as.double(ceiling(median(as.integer(x))))
}

Nepal_PPclustToIndex_dict <- 1:length(unique(Nepal_PP_metaGPSC$GPSC))
names(Nepal_PPclustToIndex_dict) <- sort(unique(Nepal_PP_metaGPSC$GPSC))

# overall
Nepal_ggCaller_intermed_consensus <- data.frame(matrix(0, nrow = sum(Nepal_gene_filter)+1, ncol = no_Nepal_PP+1))
Nepal_ggCaller_intermed_consensus[1,-1] <- paste("SeqCl_",names(Nepal_PPclustToIndex_dict),sep = "")

for (i in unique(Nepal_PP_metaGPSC$GPSC)) {
  Nepal_ggCaller_intermed_consensus[-1,(Nepal_PPclustToIndex_dict[i])+1] <- apply(as.matrix(Nepal_ggCaller_intermed[-1,c(FALSE,Nepal_seq_clusters_dict_metadata[unlist(Nepal_ggCaller_intermed[1,-1])]==i)]), 1, cons_genomes)

}
Nepal_ggCaller_intermed_consensus[,1] <- Nepal_ggCaller_intermed[,1]

#by year
Nepal_ggCaller_byYear_intermed_consensus <- vector(mode = "list", length = length(Nepal_time_points_selected))

for (i in 1:length(Nepal_time_points_selected)) {
  Nepal_ggCaller_year <- data.frame(matrix(0, nrow = sum(Nepal_gene_filter)+1,ncol = no_Nepal_PP+1))
  Nepal_ggCaller_year[1,-1] <- paste("SeqCl_",names(Nepal_PPclustToIndex_dict),sep = "")
  for (j in unique(Nepal_PP_metaGPSC$GPSC)) {
    Nepal_ggCaller_year[-1,(Nepal_PPclustToIndex_dict[j])+1] <- apply(as.matrix(Nepal_ggCaller_byYear_intermed[[i]][-1,c(FALSE,Nepal_seq_clusters_dict_metadata[unlist(Nepal_ggCaller_byYear_intermed[[i]][1,-1])]==j)]), 1, cons_genomes)
  }
  Nepal_ggCaller_year[,1] <- Nepal_ggCaller_byYear_intermed[[i]][,1]
  Nepal_ggCaller_byYear_intermed_consensus[[i]] <- Nepal_ggCaller_year
}

saveRDS(Nepal_ggCaller_intermed_consensus,"Nepal_ggCaller_intermed_consensus_metaGPSC.rds")
```

# calculate cluster frequencies for different years
```{r}
Nepal_cluster_freqs <- vector(mode = "list", length = length(Nepal_time_points_selected))

for(j in 1:length(Nepal_time_points_selected)){
  Nepal_cluster_freqs[[j]] <- rep(0, no_Nepal_PP)
  for (i in unique(Nepal_PP_metaGPSC$GPSC)){
    #Nepal_cluster_freqs[[j]][which(unique(Nepal_PP$GPSC) == i)] <- length(which(Nepal_meta_dataPP_metaGPSC[Nepal_meta_IDtoIndPP[Nepal_PP[Nepal_PP$GPSC == i,"sample"]],]$Year==Nepal_time_points_selected[j]))
    Nepal_cluster_freqs[[j]][Nepal_PPclustToIndex_dict[i]] <- length(which(Nepal_meta_dataPP_metaGPSC[Nepal_meta_IDtoIndPP[Nepal_PP_metaGPSC[Nepal_PP_metaGPSC$GPSC == i,"sample"]],]$Year==Nepal_time_points_selected[j]))
  }
}

# I should use a combination of 2009, 2012 and 2014 (see next code bit)
#Nepal_time_points_preVac <- c(5,8,10)
Nepal_time_points_preVac <- c(1,4,6)


Nepal_cluster_freqs_preVac <- rep(0, no_Nepal_PP)
for (i in Nepal_time_points_preVac) {
  Nepal_cluster_freqs_preVac <- Nepal_cluster_freqs_preVac + replace(Nepal_cluster_freqs[[i]],which(is.na(Nepal_cluster_freqs[[i]])),0)
}

for (i in 1:length(Nepal_time_points_selected)) {
  file_name <- paste("Nepal_cluster_freqs_metaGPSC", as.character(i),".rds", sep = "")
  saveRDS(Nepal_cluster_freqs[[i]], file_name)
}
```

#create initial population
```{r}
### create initial population that is based on the first time point data set
# sample from it with an Poisson process
expand_factor <- 15000 / sum(Nepal_cluster_freqs_preVac)
exp_noise <- 10
Nepal_model_start_pop <- (sapply((Nepal_cluster_freqs_preVac + rexp(n = length(Nepal_cluster_freqs_preVac), rate = exp_noise)) * expand_factor, rpois, n=1))

# visual check for the "sampling"
plot(Nepal_cluster_freqs_preVac/sum(Nepal_cluster_freqs_preVac))
points(Nepal_model_start_pop/sum(Nepal_model_start_pop), col = "red")

# aha, I compared the data but 90% of the data are actually coming from 2009 (of the preVac era)
# so it would be much better to only use that time point I think

# create start population based on all three pre-vac years
Nepal_model_start_pop <- (sapply((Nepal_cluster_freqs_preVac + rexp(n = length(Nepal_cluster_freqs_preVac), rate = exp_noise)) * expand_factor, rpois, n=1))

# visual check for the "sampling"
plot(Nepal_cluster_freqs_preVac/sum(Nepal_cluster_freqs_preVac))
points(Nepal_model_start_pop/sum(Nepal_model_start_pop), col = "red")

saveRDS(Nepal_model_start_pop, "Nepal_model_start_pop_metaGPSC.rds")
```


# calculate delta ranking
# I will do 2019 - (2009+2012+2014)
```{r}
# For strong and weak selection in the model, I need to calculate the beta statistics.

# calculate gene frequencies first, separate for three time points
#Nepal_gene_freq_preVac <- apply(Nepal_ggCaller_intermed_preVac[-1,-1], 1, sum_as_int)
#Nepal_gene_freq_postVac <- apply(Nepal_ggCaller_byYear_intermed[[length(Nepal_ggCaller_byYear_intermed)]][-1,-1], 1, sum_as_int)
Nepal_gene_freq_preVac <- apply(Nepal_ggCaller_intermed_preVac[-1,-1], 1, sum_as_int) # this now includes all three pre-vac samples
#Nepal_gene_freq_postVac <- apply(Nepal_ggCaller_byYear_intermed[[15]][-1,-1], 1, sum_as_int)
Nepal_gene_freq_postVac <- apply(Nepal_ggCaller_byYear_intermed[[11]][-1,-1], 1, sum_as_int)


# first, calculate pre/peri and post vacc frequencies of genes:
#Nepal_gene_freq_preVac_rel <- Nepal_gene_freq_preVac/sum(Nepal_gene_freq_preVac)
#Nepal_gene_freq_postVac_rel <- Nepal_gene_freq_postVac/sum(Nepal_gene_freq_postVac)
# no, I don't want to divide by the sum of the number of genes, I want to divide by number of seqs
Nepal_gene_freq_preVac_rel <- Nepal_gene_freq_preVac/(ncol(Nepal_ggCaller_intermed_preVac[-1,-1]))
Nepal_gene_freq_postVac_rel <- Nepal_gene_freq_postVac/(ncol(Nepal_ggCaller_byYear_intermed[[11]][-1,-1]))

# calculate delta statistic (refer to Corander et al. for more info)
Nepal_delta_data <- (Nepal_gene_freq_postVac_rel - Nepal_gene_freq_preVac_rel) ^ 2 / (1 - Nepal_gene_freq_preVac_rel * (1 - Nepal_gene_freq_preVac_rel))
Nepal_delta_ranking <- rank(Nepal_delta_data)

saveRDS(Nepal_delta_ranking,"Nepal_delta_ranking.rds")
saveRDS(Nepal_delta_data,"Nepal_delta_data.rds")
```

### Data processing for PPxSero model
```{r}
Nepal_clusters <- readRDS("Nepal_PP_metaGPSC.rds")
Nepal_model_start_pop <- readRDS("Nepal_model_start_pop_metaGPSC.rds")
Nepal_clusters_no <- length(unique(Nepal_clusters$GPSC))
Nepal_avg_cluster_freq <- rep(1/Nepal_clusters_no, Nepal_clusters_no)


PPxSero_matrix <- matrix(0, nrow = length(unique(Nepal_clusters$GPSC)), ncol = length(unique(Nepal_clusters$Serotype)))

SerotypeToIndex_dict <- 1:length(unique(Nepal_clusters$Serotype))
names(SerotypeToIndex_dict) <- sort(unique(Nepal_clusters$Serotype))

IndexToSerotype_dict <- sort(unique(Nepal_clusters$Serotype))
names(IndexToSerotype_dict) <- 1:length(unique(Nepal_clusters$Serotype))

Nepal_PPclustToIndex_dict <- 1:length(unique(Nepal_clusters$GPSC))
names(Nepal_PPclustToIndex_dict) <- sort(unique(Nepal_clusters$GPSC))

for (i in 1:nrow(Nepal_clusters)) {
  PPxSero_matrix[Nepal_PPclustToIndex_dict[as.character(Nepal_clusters$GPSC[i])],SerotypeToIndex_dict[Nepal_clusters$Serotype[i]]] <- 1
}


PPxSero_matrix_prob2 <-  matrix(0, nrow = length(unique(Nepal_clusters$GPSC)), ncol = length(unique(Nepal_clusters$Serotype)))
for (i in (unique(Nepal_clusters$GPSC))) {
  PPxSero_matrix_prob2[Nepal_PPclustToIndex_dict[i],] <- ((PPxSero_matrix[Nepal_PPclustToIndex_dict[i],] * Nepal_avg_cluster_freq[1])/sum(PPxSero_matrix[Nepal_PPclustToIndex_dict[i],]))
}
saveRDS(PPxSero_matrix_prob2, "Nepal_PPsero_mig_metaGPSC.rds")
# new migration matrix that is closer to original one

# start population:
# Create PP clusters, split by serotype
#calculate the frequency of the gene clusters and year and serotype
Nepal_mass_cluster_freq_1_sero <- matrix(0,nrow = length(unique(Nepal_clusters$GPSC)), ncol = length(unique(Nepal_clusters$Serotype)))
for (i in (unique(Nepal_clusters$GPSC))){
  for (k in unique(Nepal_clusters$Serotype)){
  Nepal_mass_cluster_freq_1_sero[Nepal_PPclustToIndex_dict[i],SerotypeToIndex_dict[k]] <- length(which(Nepal_SeqYear_dict[Nepal_clusters[which(Nepal_clusters$GPSC==i),][Nepal_clusters[which(Nepal_clusters$GPSC==i),]$Serotype ==k, ]$"sample"] <= 2014))
  }
}

#save sequence cluster frequencies
saveRDS(Nepal_mass_cluster_freq_1_sero, file = "Nepal_mass_cluster_freq_1_sero_metaGPSC.rds")


PPsero_startpop2 <-  matrix(0, nrow = length(unique(Nepal_clusters$GPSC)), ncol = length(unique(Nepal_clusters$Serotype)))
PP_expand_factor <- 15000 / sum(Nepal_cluster_freqs_preVac)
exp_noise <- 1000
for (i in (unique(Nepal_clusters$GPSC))) {
  #PPsero_startpop2[PPclustToIndex_dict[i],] <- round((PPxSero_matrix[i,] * Nepal_model_start_pop[i])/sum(PPxSero_matrix[i,]))
  PPsero_startpop2[Nepal_PPclustToIndex_dict[i],] <- (sapply((Nepal_mass_cluster_freq_1_sero[Nepal_PPclustToIndex_dict[i],] + rexp(n = length(Nepal_mass_cluster_freq_1_sero[Nepal_PPclustToIndex_dict[i],]), rate = exp_noise)) * PP_expand_factor, rpois, n=1)) + ceiling(PPxSero_matrix_prob2[Nepal_PPclustToIndex_dict[i],])
}
saveRDS(PPsero_startpop2, "Nepal_PPsero_startpop_metaGPSC.rds")


# serotype VT vector
SeroVT <- rep(0, length(unique(Nepal_clusters$Serotype)))
for (i in 1:nrow(Nepal_clusters)) {
  SeroVT[SerotypeToIndex_dict[Nepal_clusters$Serotype[i]]] <- as.integer(Nepal_clusters$VT[i]=="VT")
}
names(SeroVT) <- names(SerotypeToIndex_dict)
saveRDS(SeroVT, "Nepal_SeroVT_metaGPSC.rds")

### vt vector that includes 6A (which appears to have cross immunity with 6B and is therefore affected by the vaccine, e.g. https://pmc.ncbi.nlm.nih.gov/articles/PMC5848315/)
SeroVT_6A <- SeroVT
SeroVT_6A["6A"] <- 1
saveRDS(SeroVT_6A, "Nepal_SeroVT_6A_metaGPSC.rds")
```

