---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: console
---
```{r}
library(ggplot2)
```


# calculate model and data credible intervals
```{r}
calc_model_cred_int <- function(sim_model, data_size){
  sample_reps <- 1000
  credible_sample <- array(NA, dim = c(nrow(sim_model) * sample_reps, ncol(sim_model)))
  for (i in 1:nrow(sim_model)) {
    credible_sample[(sample_reps *(i-1)+1):(sample_reps *(i)),] <- t(rmultinom(n = sample_reps, size = data_size,  prob = sim_model[i,]))
  }
  mean_credible_sample <- apply(credible_sample,2,mean)/data_size
  percentile95_up_credible_sample <- apply(credible_sample,2,quantile,.975)/data_size
  percentile95_low_credible_sample <- apply(credible_sample,2,quantile,.025)/data_size
  return(list(mean_credible_sample, percentile95_up_credible_sample, percentile95_low_credible_sample))
}

calc_data_cred_int <- function(post_data, data_size){
  sample_reps <- 1000
  credible_sample <- t(rmultinom(n = sample_reps, size = data_size,  prob = post_data))
  mean_credible_sample <- apply(credible_sample,2,mean)/data_size
  percentile95_up_credible_sample <- apply(credible_sample,2,quantile,.975)/data_size
  percentile95_low_credible_sample <- apply(credible_sample,2,quantile,.025)/data_size
  return(list(mean_credible_sample, percentile95_up_credible_sample, percentile95_low_credible_sample))
}
```

# calculate model and data credible intervals
```{r}
calc_data_cred_int_preVac <- function(post_data, data_size){
  sample_reps <- 1000
  credible_sample <- t(rmultinom(n = sample_reps, size = data_size,  prob = post_data))
  return(credible_sample)
}


calc_model_cred_int_minusPre <- function(sim_model, data_size, credible_sample_pre_data){
  sample_reps <- 1000
  credible_sample <- array(NA, dim = c(nrow(sim_model) * sample_reps, ncol(sim_model)))
  for (i in 1:nrow(sim_model)) {
    credible_sample[(sample_reps *(i-1)+1):(sample_reps *(i)),] <- t(rmultinom(n = sample_reps, size = data_size,  prob = sim_model[i,]))
  }
  mean_credible_sample <- apply(credible_sample,2,mean)/data_size
  
  mean_credible_sample_pre_data <- apply(credible_sample_pre_data,2,mean)/data_size
  
  mean_credible_sample <- mean_credible_sample - mean_credible_sample_pre_data
  percentile95_up_credible_sample <- apply((credible_sample - rep(credible_sample_pre_data, each = nrow(sim_model))), 2,quantile,.975)/data_size
  percentile95_low_credible_sample <- apply((credible_sample - rep(credible_sample_pre_data, each = nrow(sim_model))),2,quantile,.025)/data_size
  return(list(mean_credible_sample, percentile95_up_credible_sample, percentile95_low_credible_sample))
}

calc_data_cred_int_minusPre <- function(post_data, data_size, credible_sample_pre_data){
  sample_reps <- 1000
  credible_sample <- t(rmultinom(n = sample_reps, size = data_size,  prob = post_data))
  mean_credible_sample <- apply(credible_sample,2,mean)/data_size
  
  mean_credible_sample_pre_data <- apply(credible_sample_pre_data,2,mean)/data_size
  
  mean_credible_sample <- mean_credible_sample - mean_credible_sample_pre_data
  percentile95_up_credible_sample <- apply(credible_sample - credible_sample_pre_data,2,quantile,.975)/data_size
  percentile95_low_credible_sample <- apply(credible_sample - credible_sample_pre_data,2,quantile,.025)/data_size
  return(list(mean_credible_sample, percentile95_up_credible_sample, percentile95_low_credible_sample))
}
```

# calculate VT, NVT, and mixed information for GPSCs
```{r}
calc_VTNVT_cluster_info <- function(seq_cluster_data){
  cluster_VTs <- rep("NVT", length(unique(seq_cluster_data$GPSC)))
  names(cluster_VTs) <- unique(seq_cluster_data$GPSC)
  for (i in 1:length(unique(seq_cluster_data$GPSC))) {
    count_vts <- 0
    count_in_cluster <- length(which(seq_cluster_data$GPSC == unique(seq_cluster_data$GPSC)[i]))
    for (j in which(seq_cluster_data$GPSC == unique(seq_cluster_data$GPSC)[i])) {
      if(seq_cluster_data$VT[j] == "VT"){
        count_vts <- count_vts + 1
      }
    }
    if(count_vts == count_in_cluster){
      cluster_VTs[i] <- "VT"
    } else if(count_vts == 0){
      cluster_VTs[i] <- "NVT"
    } else{
      cluster_VTs[i] <- "mixed"
    }
  }
  cluster_VTs
}
```

# create plot with data, NFDS model, null model, including credible intervals
```{r}
create_cred_int_plot <- function(post_VT_data, NFDS_model_out, Null_model_out, seq_cluster_data, output_location, preVac_data){
  # calculate credible intervals for data, NFDS model output, and null model output
  data_cred_int_pre <- calc_data_cred_int(preVac_data, sum(preVac_data))
  data_cred_int <- calc_data_cred_int(post_VT_data, sum(post_VT_data))
  NFDS_cred_int <- calc_model_cred_int(NFDS_model_out, sum(post_VT_data))
  null_cred_int <- calc_model_cred_int(Null_model_out, sum(post_VT_data))
  
  # calculate which models are VT, NVT, or mixed
  cluster_VTs_info <- calc_VTNVT_cluster_info(seq_cluster_data)
  
  # choose to display only clusters that are at least at 1% frequency at the beginning
  no_small_clusters <- which(preVac_data/sum(preVac_data) > 0.01)
  
  # create df for ggplot
  data_and_model_with_errors_and_null_ModelSample <- data.frame(
  cluster=rep((unique(seq_cluster_data$GPSC)[no_small_clusters]),3),
  xvals = rep(1:length(unique(seq_cluster_data$GPSC)[no_small_clusters]),3),
  Group = c(rep("Data",length(unique(seq_cluster_data$GPSC)[no_small_clusters])),rep("NFDS Model",length(unique(seq_cluster_data$GPSC)[no_small_clusters])),rep("Null Model",length(unique(seq_cluster_data$GPSC)[no_small_clusters]))),
  VT = rep(cluster_VTs_info[no_small_clusters],3),
  values=c(data_cred_int[[1]][no_small_clusters], NFDS_cred_int[[1]][no_small_clusters], null_cred_int[[1]][no_small_clusters]),
  upper=c(data_cred_int[[2]][no_small_clusters], NFDS_cred_int[[2]][no_small_clusters], null_cred_int[[2]][no_small_clusters]),
  lower = c(data_cred_int[[3]][no_small_clusters], NFDS_cred_int[[3]][no_small_clusters], null_cred_int[[3]][no_small_clusters]))

  # VTs and NVTs separately
  data_and_model_with_errors_and_null_ModelSample_NVT <- data_and_model_with_errors_and_null_ModelSample[which(data_and_model_with_errors_and_null_ModelSample$VT=="NVT"),]
  data_and_model_with_errors_and_null_ModelSample_NVT$xvals <- rep(1:(nrow(data_and_model_with_errors_and_null_ModelSample_NVT)/3),3)
  data_and_model_with_errors_and_null_ModelSample_VT <- data_and_model_with_errors_and_null_ModelSample[which(data_and_model_with_errors_and_null_ModelSample$VT=="VT"),]
  data_and_model_with_errors_and_null_ModelSample_VT$xvals <- rep(1:(nrow(data_and_model_with_errors_and_null_ModelSample_VT)/3),3)
  data_and_model_with_errors_and_null_ModelSample_mixed <- data_and_model_with_errors_and_null_ModelSample[which(data_and_model_with_errors_and_null_ModelSample$VT=="mixed"),]
  data_and_model_with_errors_and_null_ModelSample_mixed$xvals <- rep(1:(nrow(data_and_model_with_errors_and_null_ModelSample_mixed)/3),3)
 
  # make NVT plot

  plot_NVT <- ggplot(data_and_model_with_errors_and_null_ModelSample_NVT, aes(y=values, x=xvals,group = Group, color = Group)) + 
    geom_point(position = position_dodge(.3), stat="identity",size = 1.5) +
  geom_errorbar( aes(ymin=lower, ymax=upper), width=0.5, alpha=0.9, linewidth=1,position=position_dodge(.3)) +
  scale_x_continuous("Strains", breaks = 1:(length(which(cluster_VTs_info[no_small_clusters]=="NVT"))), labels = names(cluster_VTs_info[no_small_clusters])[which(cluster_VTs_info[no_small_clusters]=="NVT")])+
  scale_color_manual(values = c("black", "#0072B2", "#CC79A7"),labels = c("Data", "NFDS Model","Neutral Model")) +
  labs(x = "Strains", y = "Frequency") +
  ggtitle("NVT Clusters Data vs. NFDS Model vs. Neutral Model\n with 95% multinomial confidence intervals") +  # Add the plot title
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20)) +
  ylim(0, 0.25)
  
    ggsave(plot_NVT, 
      filename = paste(output_location,"NVT_DatavsNFDSvsNull.pdf",sep = ""),   # The directory and file name you want to save the file in
      device = "pdf",
      width = 24, # The width of the plot in inches
      height = 12)
  
  # make VT plot
  if(nrow(data_and_model_with_errors_and_null_ModelSample_VT)>0){

    plot_VT <- ggplot(data_and_model_with_errors_and_null_ModelSample_VT, aes(y=values, x=xvals,group = Group, color = Group)) + 
    geom_point(position = position_dodge(.3), stat="identity",size = 1.5) +
    geom_errorbar( aes(ymin=lower, ymax=upper), width=0.5, alpha=0.9, linewidth=1,position=position_dodge(.3)) +
    scale_x_continuous("Strains", breaks = 1:(length(which(cluster_VTs_info[no_small_clusters]=="VT"))), labels = names(cluster_VTs_info[no_small_clusters])[which(cluster_VTs_info[no_small_clusters]=="VT")])+
    scale_color_manual(values = c("black", "#0072B2", "#CC79A7"),labels = c("Data", "NFDS Model","Neutral Model")) +
    labs(x = "Strains", y = "Frequency") +
    ggtitle("VT Clusters Data vs. NFDS Model vs. Neutral Model\n with 95% multinomial confidence intervals") +  # Add the plot title
    theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20)) +
    ylim(0, 0.25)

    ggsave(plot_VT, 
      filename = paste(output_location,"VT_DatavsNFDSvsNull.pdf",sep = ""),   # The directory and file name you want to save the file in
      device = "pdf",
      width = 24, # The width of the plot in inches
      height = 12)
  }
  
  # make mixed plot

  plot_mixed <- ggplot(data_and_model_with_errors_and_null_ModelSample_mixed, aes(y=values, x=xvals,group = Group, color = Group)) + 
    geom_point(position = position_dodge(.3), stat="identity",size = 1.5) +
  geom_errorbar( aes(ymin=lower, ymax=upper), width=0.5, alpha=0.9, linewidth=1,position=position_dodge(.3)) +
  scale_x_continuous("Strains", breaks = 1:(length(which(cluster_VTs_info[no_small_clusters]=="mixed"))), labels = names(cluster_VTs_info[no_small_clusters])[which(cluster_VTs_info[no_small_clusters]=="mixed")])+
  scale_color_manual(values = c("black", "#0072B2", "#CC79A7"),labels = c("Data", "NFDS Model","Neutral Model")) +
  labs(x = "Strains", y = "Frequency") +
  ggtitle("Mixed Clusters Data vs. NFDS Model vs. Neutral Model\n with 95% multinomial confidence intervals") +  # Add the plot title
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20)) +
  ylim(0, 0.25)
  
  ggsave(plot_mixed, 
    filename = paste(output_location,"Mixed_DatavsNFDSvsNull.pdf",sep = ""),   # The directory and file name you want to save the file in
    device = "pdf",
    width = 24, # The width of the plot in inches
    height = 12)
}
```

```{r}
create_cred_int_plot_change <- function(post_VT_data, NFDS_model_out, Null_model_out, seq_cluster_data, output_location, preVac_data, minmax){
  # calculate credible intervals for data, NFDS model output, and null model output
  data_cred_int_pre <- calc_data_cred_int_preVac(preVac_data, sum(post_VT_data))
  data_cred_int <- calc_data_cred_int_minusPre(post_VT_data, sum(post_VT_data), data_cred_int_pre)
  NFDS_cred_int <- calc_model_cred_int_minusPre(NFDS_model_out, sum(post_VT_data), data_cred_int_pre)
  null_cred_int <- calc_model_cred_int_minusPre(Null_model_out, sum(post_VT_data), data_cred_int_pre)
  
  # calculate which models are VT, NVT, or mixed
  cluster_VTs_info <- calc_VTNVT_cluster_info(seq_cluster_data)
  
  # choose to display only clusters that are at least at 1% frequency at the beginning
  no_small_clusters <- which(preVac_data/sum(preVac_data) > 0.0)
  
  # create df for ggplot
  data_and_model_with_errors_and_null_ModelSample <- data.frame(
  cluster=rep((unique(seq_cluster_data$GPSC)[no_small_clusters]),3),
  xvals = rep(1:length(unique(seq_cluster_data$GPSC)[no_small_clusters]),3),
  Group = c(rep("Data",length(unique(seq_cluster_data$GPSC)[no_small_clusters])),rep("NFDS Model",length(unique(seq_cluster_data$GPSC)[no_small_clusters])),rep("Null Model",length(unique(seq_cluster_data$GPSC)[no_small_clusters]))),
  VT = rep(cluster_VTs_info[no_small_clusters],3),
  values=c(data_cred_int[[1]][no_small_clusters], NFDS_cred_int[[1]][no_small_clusters], null_cred_int[[1]][no_small_clusters]),
  upper=c(data_cred_int[[2]][no_small_clusters], NFDS_cred_int[[2]][no_small_clusters], null_cred_int[[2]][no_small_clusters]),
  lower = c(data_cred_int[[3]][no_small_clusters], NFDS_cred_int[[3]][no_small_clusters], null_cred_int[[3]][no_small_clusters]))
  
  # VTs and NVTs separately
  data_and_model_with_errors_and_null_ModelSample_NVT <- data_and_model_with_errors_and_null_ModelSample[which(data_and_model_with_errors_and_null_ModelSample$VT=="NVT"),]
  data_and_model_with_errors_and_null_ModelSample_NVT$xvals <- rep(1:(nrow(data_and_model_with_errors_and_null_ModelSample_NVT)/3),3)
  data_and_model_with_errors_and_null_ModelSample_VT <- data_and_model_with_errors_and_null_ModelSample[which(data_and_model_with_errors_and_null_ModelSample$VT=="VT"),]
  data_and_model_with_errors_and_null_ModelSample_VT$xvals <- rep(1:(nrow(data_and_model_with_errors_and_null_ModelSample_VT)/3),3)
  data_and_model_with_errors_and_null_ModelSample_mixed <- data_and_model_with_errors_and_null_ModelSample[which(data_and_model_with_errors_and_null_ModelSample$VT=="mixed"),]
  data_and_model_with_errors_and_null_ModelSample_mixed$xvals <- rep(1:(nrow(data_and_model_with_errors_and_null_ModelSample_mixed)/3),3)
 
  # make NVT plot

  plot_NVT <- ggplot(data_and_model_with_errors_and_null_ModelSample_NVT, aes(y=values, x=xvals,group = Group, color = Group)) + 
    geom_hline(yintercept = 0, linetype="dashed") +
    geom_point(position = position_dodge(.3), stat="identity",size = 1.5) +
  geom_errorbar( aes(ymin=lower, ymax=upper), width=0.5, alpha=0.9, linewidth=1,position=position_dodge(.3)) +
  scale_x_continuous("Strains", breaks = 1:(length(which(cluster_VTs_info[no_small_clusters]=="NVT"))), labels = names(cluster_VTs_info[no_small_clusters])[which(cluster_VTs_info[no_small_clusters]=="NVT")])+
  scale_color_manual(values = c("black", "#0072B2", "#CC79A7"),labels = c("Data", "NFDS Model","Neutral Model")) +
  labs(x = "Strains", y = "Frequency") +
  ggtitle("NVT Clusters Data vs. NFDS Model vs. Neutral Model\n with 95% multinomial confidence intervals") +  # Add the plot title
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20)) +
  ylim(-minmax, minmax)
  
    ggsave(plot_NVT, 
      filename = paste(output_location,"Change_NVT_DatavsNFDSvsNull.pdf",sep = ""),   # The directory and file name you want to save the file in
      device = "pdf",
      width = 24, # The width of the plot in inches
      height = 12)
  
  # make VT plot
  if(nrow(data_and_model_with_errors_and_null_ModelSample_VT)>0){

    plot_VT <- ggplot(data_and_model_with_errors_and_null_ModelSample_VT, aes(y=values, x=xvals,group = Group, color = Group)) +  
    geom_hline(yintercept = 0, linetype="dashed") +
    geom_point(position = position_dodge(.3), stat="identity",size = 1.5) +
    geom_errorbar( aes(ymin=lower, ymax=upper), width=0.5, alpha=0.9, linewidth=1,position=position_dodge(.3)) +
    scale_x_continuous("Strains", breaks = 1:(length(which(cluster_VTs_info[no_small_clusters]=="VT"))), labels = names(cluster_VTs_info[no_small_clusters])[which(cluster_VTs_info[no_small_clusters]=="VT")])+
    scale_color_manual(values = c("black", "#0072B2", "#CC79A7"),labels = c("Data", "NFDS Model","Neutral Model")) +
    labs(x = "Strains", y = "Frequency") +
    ggtitle("VT Clusters Data vs. NFDS Model vs. Neutral Model\n with 95% multinomial confidence intervals") +  # Add the plot title
    theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20)) +
    ylim(-minmax, minmax)

    ggsave(plot_VT, 
      filename = paste(output_location,"Change_VT_DatavsNFDSvsNull.pdf",sep = ""),   # The directory and file name you want to save the file in
      device = "pdf",
      width = 24, # The width of the plot in inches
      height = 12)
  }
  
  # make mixed plot

  plot_mixed <- ggplot(data_and_model_with_errors_and_null_ModelSample_mixed, aes(y=values, x=xvals,group = Group, color = Group)) +  
    geom_hline(yintercept = 0, linetype="dashed") +
    geom_point(position = position_dodge(.3), stat="identity",size = 1.5) +
  geom_errorbar( aes(ymin=lower, ymax=upper), width=0.5, alpha=0.9, linewidth=1,position=position_dodge(.3)) +
  scale_x_continuous("Strains", breaks = 1:(length(which(cluster_VTs_info[no_small_clusters]=="mixed"))), labels = names(cluster_VTs_info[no_small_clusters])[which(cluster_VTs_info[no_small_clusters]=="mixed")])+
  scale_color_manual(values = c("black", "#0072B2", "#CC79A7"),labels = c("Data", "NFDS Model","Neutral Model")) +
  labs(x = "Strains", y = "Frequency") +
  ggtitle("Mixed Clusters Data vs. NFDS Model vs. Neutral Model\n with 95% multinomial confidence intervals") +  # Add the plot title
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20)) +
  ylim(-minmax, minmax)
  
  ggsave(plot_mixed, 
    filename = paste(output_location,"Change_Mixed_DatavsNFDSvsNull.pdf",sep = ""),   # The directory and file name you want to save the file in
    device = "pdf",
    width = 24, # The width of the plot in inches
    height = 12)
}
```

```{r}
create_cred_int_plot_change_sorted <- function(post_VT_data, NFDS_model_out, Null_model_out, seq_cluster_data, output_location, preVac_data, minmax){
  # calculate credible intervals for data, NFDS model output, and null model output
  data_cred_int_pre <- calc_data_cred_int_preVac(preVac_data, sum(post_VT_data))
  data_cred_int <- calc_data_cred_int_minusPre(post_VT_data, sum(post_VT_data), data_cred_int_pre)
  NFDS_cred_int <- calc_model_cred_int_minusPre(NFDS_model_out, sum(post_VT_data), data_cred_int_pre)
  null_cred_int <- calc_model_cred_int_minusPre(Null_model_out, sum(post_VT_data), data_cred_int_pre)
  
  # calculate which models are VT, NVT, or mixed
  cluster_VTs_info <- calc_VTNVT_cluster_info(seq_cluster_data)
  
  # choose to display only clusters that are at least at 1% frequency at the beginning
  no_small_clusters <- which(preVac_data/sum(preVac_data) > 0.0)
  
  # create df for ggplot
  data_and_model_with_errors_and_null_ModelSample <- data.frame(
  cluster=rep((unique(seq_cluster_data$GPSC)[no_small_clusters]),3),
  xvals = rep(1:length(unique(seq_cluster_data$GPSC)[no_small_clusters]),3),
  Group = c(rep("Data",length(unique(seq_cluster_data$GPSC)[no_small_clusters])),rep("NFDS Model",length(unique(seq_cluster_data$GPSC)[no_small_clusters])),rep("Null Model",length(unique(seq_cluster_data$GPSC)[no_small_clusters]))),
  VT = rep(cluster_VTs_info[no_small_clusters],3),
  values=c(data_cred_int[[1]][no_small_clusters], NFDS_cred_int[[1]][no_small_clusters], null_cred_int[[1]][no_small_clusters]),
  upper=c(data_cred_int[[2]][no_small_clusters], NFDS_cred_int[[2]][no_small_clusters], null_cred_int[[2]][no_small_clusters]),
  lower = c(data_cred_int[[3]][no_small_clusters], NFDS_cred_int[[3]][no_small_clusters], null_cred_int[[3]][no_small_clusters]))
  
  # compute order by frequency changes
  # VTs and NVTs separately
  data_and_model_with_errors_and_null_ModelSample_NVT <- data_and_model_with_errors_and_null_ModelSample[which(data_and_model_with_errors_and_null_ModelSample$VT=="NVT"),]
  data_and_model_with_errors_and_null_ModelSample_NVT$xvals <- rep(1:(nrow(data_and_model_with_errors_and_null_ModelSample_NVT)/3),3)
  #data_and_model_with_errors_and_null_ModelSample_NVT$xvals <- rep(order(data_and_model_with_errors_and_null_ModelSample_NVT$values[which(data_and_model_with_errors_and_null_ModelSample_NVT$Group == "Data")]),3)
  data_and_model_with_errors_and_null_ModelSample_VT <- data_and_model_with_errors_and_null_ModelSample[which(data_and_model_with_errors_and_null_ModelSample$VT=="VT"),]
  data_and_model_with_errors_and_null_ModelSample_VT$xvals <- rep(1:(nrow(data_and_model_with_errors_and_null_ModelSample_VT)/3),3)
  #data_and_model_with_errors_and_null_ModelSample_VT$xvals <- rep(order(data_and_model_with_errors_and_null_ModelSample_VT$values[which(data_and_model_with_errors_and_null_ModelSample_VT$Group == "Data")]),3)
  data_and_model_with_errors_and_null_ModelSample_mixed <- data_and_model_with_errors_and_null_ModelSample[which(data_and_model_with_errors_and_null_ModelSample$VT=="mixed"),]
  data_and_model_with_errors_and_null_ModelSample_mixed$xvals <- rep(1:(nrow(data_and_model_with_errors_and_null_ModelSample_mixed)/3),3)
  #data_and_model_with_errors_and_null_ModelSample_mixed$xvals <- rep(order(data_and_model_with_errors_and_null_ModelSample_mixed$values[which(data_and_model_with_errors_and_null_ModelSample_mixed$Group == "Data")]),3)
  
  level_order_NVT <- order(data_and_model_with_errors_and_null_ModelSample_NVT$values[which(data_and_model_with_errors_and_null_ModelSample_NVT$Group == "Data")], decreasing = TRUE)
  level_order_VT <- order(data_and_model_with_errors_and_null_ModelSample_VT$values[which(data_and_model_with_errors_and_null_ModelSample_VT$Group == "Data")], decreasing = TRUE)
  level_order_mixed <- order(data_and_model_with_errors_and_null_ModelSample_mixed$values[which(data_and_model_with_errors_and_null_ModelSample_mixed$Group == "Data")], decreasing = TRUE)
 
  # make NVT plot

  plot_NVT <- ggplot(data_and_model_with_errors_and_null_ModelSample_NVT, aes(y=values, x=factor(xvals,level = level_order_NVT), group = Group, color = Group)) + 
    geom_hline(yintercept = 0, linetype="dashed") +
    geom_point(position = position_dodge(.5), stat="identity",size = 2.5) +
  geom_errorbar( aes(ymin=lower, ymax=upper), width=0.5, alpha=0.9, linewidth=1,position=position_dodge(.5)) +
  scale_x_discrete("Strains", breaks = 1:(length(which(cluster_VTs_info[no_small_clusters]=="NVT"))), labels = names(cluster_VTs_info[no_small_clusters])[which(cluster_VTs_info[no_small_clusters]=="NVT")])+
  scale_color_manual(values = c("black", "#0072B2", "#CC79A7"),labels = c("Data", "NFDS Model","Neutral Model")) +
  labs(x = "Strains", y = "Frequency") +
  ggtitle("NVT Clusters Data vs. NFDS Model vs. Neutral Model\n with 95% multinomial confidence intervals") +  # Add the plot title
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20)) +
  ylim(-minmax, minmax)
  
    ggsave(plot_NVT, 
      filename = paste(output_location,"Change_NVT_DatavsNFDSvsNull.pdf",sep = ""),   # The directory and file name you want to save the file in
      device = "pdf",
      width = length(level_order_NVT), # The width of the plot in inches
      height = 6)
  
  # make VT plot
  if(nrow(data_and_model_with_errors_and_null_ModelSample_VT)>0){

    plot_VT <- ggplot(data_and_model_with_errors_and_null_ModelSample_VT, aes(y=values, x=factor(xvals,level = level_order_VT), group = Group, color = Group)) +  
    geom_hline(yintercept = 0, linetype="dashed") +
    geom_point(position = position_dodge(.5), stat="identity",size = 2.5) +
  geom_errorbar( aes(ymin=lower, ymax=upper), width=0.5, alpha=0.9, linewidth=1,position=position_dodge(.5)) +
    scale_x_discrete("Strains", breaks = 1:(length(which(cluster_VTs_info[no_small_clusters]=="VT"))), labels = names(cluster_VTs_info[no_small_clusters])[which(cluster_VTs_info[no_small_clusters]=="VT")])+
    scale_color_manual(values = c("black", "#0072B2", "#CC79A7"),labels = c("Data", "NFDS Model","Neutral Model")) +
    labs(x = "Strains", y = "Frequency") +
    ggtitle("VT Clusters Data vs. NFDS Model vs. Neutral Model\n with 95% multinomial confidence intervals") +  # Add the plot title
    theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20)) +
    ylim(-minmax, minmax)

    ggsave(plot_VT, 
      filename = paste(output_location,"Change_VT_DatavsNFDSvsNull.pdf",sep = ""),   # The directory and file name you want to save the file in
      device = "pdf",
      width = length(level_order_VT), # The width of the plot in inches
      height = 6)
  }
  
  # make mixed plot

  plot_mixed <- ggplot(data_and_model_with_errors_and_null_ModelSample_mixed, aes(y=values, x=factor(xvals,level = level_order_mixed), group = Group, color = Group)) +  
    geom_hline(yintercept = 0, linetype="dashed") +
    geom_point(position = position_dodge(.5), stat="identity",size = 2.5) +
  geom_errorbar( aes(ymin=lower, ymax=upper), width=0.5, alpha=0.9, linewidth=1,position=position_dodge(.5)) +
  scale_x_discrete("Strains", breaks = 1:(length(which(cluster_VTs_info[no_small_clusters]=="mixed"))), labels = names(cluster_VTs_info[no_small_clusters])[which(cluster_VTs_info[no_small_clusters]=="mixed")])+
  scale_color_manual(values = c("black", "#0072B2", "#CC79A7"),labels = c("Data", "NFDS Model","Neutral Model")) +
  labs(x = "Strains", y = "Frequency") +
  ggtitle("Mixed Clusters Data vs. NFDS Model vs. Neutral Model\n with 95% multinomial confidence intervals") +  # Add the plot title
  theme(panel.background = element_rect(fill = "white",),
        panel.grid = element_line(color = "grey"),text = element_text(size = 20)) +
  ylim(-minmax, minmax)
  
  ggsave(plot_mixed, 
    filename = paste(output_location,"Change_Mixed_DatavsNFDSvsNull.pdf",sep = ""),   # The directory and file name you want to save the file in
    device = "pdf",
    width = length(level_order_mixed), # The width of the plot in inches
    height = 6)
}
```


# read in forward simulation of model and data and cluster information (from PopPUNK) for US
```{r}
US_params_4_model_data_rel <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_10_23/US_SimulateModelForPlots/params_4_model_data_rel.rds")
US_params_null_model_data_rel <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_10_23/US_SimulateModelForPlots/params_2_model_data_rel.rds")

US_preVac <- readRDS(file = "PP_mass_cluster_freq_1.rds")

US_2007 <- readRDS(file = "PP_mass_cluster_freq_3.rds")
US_seq_clusters <- readRDS("PopPUNK_clusters.rds")
colnames(US_seq_clusters)[2] <- "GPSC"

location_for_plots <- "/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_10_23/US_SimulateModelForPlots/"
minmax_US <- 0.15

#unique(US_seq_clusters$GPSC)
```

```{r}
create_cred_int_plot_change_sorted(US_2007, US_params_4_model_data_rel, US_params_null_model_data_rel, US_seq_clusters, location_for_plots, US_preVac, minmax_US)

create_cred_int_plot_change(US_2007, US_params_4_model_data_rel, US_params_null_model_data_rel, US_seq_clusters, location_for_plots, US_preVac, minmax_US)

create_cred_int_plot_withPre2(US_2007, US_params_4_model_data_rel, US_params_null_model_data_rel, US_seq_clusters, location_for_plots, US_preVac)
create_cred_int_plot_withPre(US_2007, US_params_4_model_data_rel, US_params_null_model_data_rel, US_seq_clusters, location_for_plots, US_preVac)
create_cred_int_plot_change(US_2007, US_params_4_model_data_rel, US_params_null_model_data_rel, US_seq_clusters, location_for_plots, US_preVac, minmax_US)
create_cred_int_plot(US_2007, US_params_4_model_data_rel, US_params_null_model_data_rel, US_seq_clusters, location_for_plots, US_preVac)
```

# read in forward simulation of model and data and cluster information (from PopPUNK) for UK
```{r}
UK_params_4_model_data_rel <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_10_23/UK_SimulateModelForPlots/params_4_model_data_rel.rds")
UK_params_null_model_data_rel <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_10_23/UK_SimulateModelForPlots/params_2_model_data_rel.rds")

UK_preVac <- readRDS(file = "UK_cluster_freqs_1.rds")
#UK_no_small_clusters <- which(UK_preVac/sum(UK_preVac) > 0.01) # 17 out of 54

UK_2012 <- readRDS(file = "UK_cluster_freqs_7.rds")
UK_seq_clusters <- readRDS("UK_PP.rds")
colnames(UK_seq_clusters)[2] <- "GPSC"
colnames(UK_seq_clusters)[6] <- "VT" # use PVC13 for VT definition here

location_for_plots_UK <- "/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_10_23/UK_SimulateModelForPlots/"
minmax_UK <- 0.35
```

```{r}
create_cred_int_plot_change_sorted(UK_2012, UK_params_4_model_data_rel, UK_params_null_model_data_rel, UK_seq_clusters, location_for_plots_UK, UK_preVac, minmax_UK)

create_cred_int_plot_change(UK_2012, UK_params_4_model_data_rel, UK_params_null_model_data_rel, UK_seq_clusters, location_for_plots_UK, UK_preVac, minmax_UK)


create_cred_int_plot_withPre(UK_2012, UK_params_4_model_data_rel, UK_params_null_model_data_rel, UK_seq_clusters, location_for_plots_UK, UK_preVac)
create_cred_int_plot(UK_2012, UK_params_4_model_data_rel, UK_params_null_model_data_rel, UK_seq_clusters, location_for_plots_UK, UK_preVac)
```

# read in forward simulation of model and data and cluster information (from PopPUNK)
```{r}
### waiting for new Nepal fits for this one
Nepal_params_4_model_data_rel <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_15/Nepal_simModelForPlot_year5/Nepal_params_4_model_data_rel.rds")
Nepal_params_null_model_data_rel <- readRDS("/Users/llorenz/Documents/PhD_Project/Code/1st_project/WF_plots_postTAC2/2025_08_15/Nepal_simModelForPlot_year5/Nepal_params_null_model_data_rel.rds")

Nepal_cluster_freq_2019 <- readRDS(file = "Nepal_cluster_freqs_11.rds")
Nepal_seq_clusters <- readRDS("Nepal_PP.rds")
```

```{r}
create_cred_int_plot(Nepal_cluster_freq_2019, Nepal_params_4_model_data_rel, Nepal_params_null_model_data_rel, Nepal_seq_clusters)
```

